### 2015-05-27 3dpas_intro.tex
initial files

% cpedoc.tex V2.0, 13 May 2010

\documentclass[times]{cpeauth}

\usepackage{moreverb}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2012} %\documentclass[acmjacm,acmnow]{acmtrans2m}
%\documentclass[conference,final]{IEEEtran}
%\documentclass[10pt,letterpaper]{article}
%\documentclass[preprint,12pt]{article}
%\documentclass[11pt,letterpaper]{article}

\usepackage{graphicx} \usepackage{url} \usepackage{color} \usepackage{enumerate}
\usepackage{longtable} %\usepackage{authblk} % Is used internally within the CPE
Class \usepackage[firstpage]{draftwatermark}

%\SetWatermarkLightness{0.7} %\SetWatermarkScale{4}

%\usepackage{pdfsync} % \acmVolume{} % \acmNumber{} % \acmYear{} % \acmMonth{}

%\usepackage{sectsty} %\usepackage{setspace}

% \usepackage{ifpdf} % \usepackage{wrapfig} % %\usepackage[cm]{fullpage} %
\usepackage{textcomp} % \usepackage{srcltx} % \usepackage{fancyhdr} %
\usepackage{setspace} \usepackage{lscape} % \usepackage{longtable} %
\usepackage{paralist}

%\usepackage{hyperref} \usepackage[novbox]{pdfsync}

% \usepackage{ams}

% \ifpdf % \usepackage[pdftex, % colorlinks=true, % linkcolor=red, %
citecolor=red, % filecolor=red, % ]{hyperref} % \else %
\usepackage[hypertex]{hyperref} % \fi

% Space saving %\usepackage[small,compact]{titlesec}
%\usepackage[small,it]{caption}

% \renewcommand\floatpagefraction{.9} % \renewcommand\topfraction{.9} %
\renewcommand\bottomfraction{.9} %\renewcommand\textfrare{.1}

\setcounter{totalnumber}{50} \setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

%\textwidth = 6in %\textheight = 8.5 in %\oddsidemargin = 0.0 in
%\evensidemargin = 0.0 in

% \topmargin = 0.0 in % \headheight = 0.0 in % \headsep = 0.0 in
%\parskip = 0.0in
%\parindent = 0.5cm
\parskip = 0.04in
% \parindent = 0.0cm % \textfloatsep = 0.1in

% \usepackage[draft,pdftex,colorlinks=true, linkcolor=blue,citecolor=blue, %
urlcolor=blue]{hyperref}

% \ifpdf % \pdfinfo{ /Author (Shantenu Jha et al.)  /Title (Programming
Abstractions for Large-Scale Distributed Applications) } % \fi

%\newcommand{\projectnamefull}{\textit{Programming Abstractions for %
Large-Scale Distributed Applications} }

\newcommand{\upup}{\vspace*{-0.5em}} \newcommand{\up}{\vspace*{-0.25em}}
\newcommand{\I}[1]{\textit{#1}} \newcommand{\B}[1]{\textbf{#1}}
\newcommand{\T}[1]{\texttt{#1}} \newcommand{\BI}[1]{\B{\I{#1}}}

 \newenvironment{shortlist}{ \vspace*{-0.8em}
  \begin{itemize} \setlength{\itemsep}{-0.3em} }{
  \end{itemize} \vspace*{-0.6em} }

 \newenvironment{shortenum}{ \vspace*{-0.8em}
  \begin{enumerate} \setlength{\itemsep}{-0.3em} }{
  \end{enumerate} \vspace*{-0.6em} }

% \ifpdf % \DeclareGraphicsExtensions{.pdf, .png, .jpg} % \else %
\DeclareGraphicsExtensions{.ps, .eps} % \fi

\definecolor{orange}{rgb}{1.0,0.3,0.0} \definecolor{violet}{rgb}{0.75,0,1}
\definecolor{darkgreen}{rgb}{0,0.6,0} \definecolor{cyan}{rgb}{0.2,0.7,0.7}
\definecolor{blueish}{rgb}{0.2,0.2,0.8}

\long\def\comment#1{{\bf \textcolor{magenta}{\bf #1}}} \long\def\ccomment#1{{\bf
\textcolor{blue}{\bf #1}}} \newcommand{\C}{\comment} \newcommand{\CC}{\ccomment}

\newcommand{\yes}{$\bullet$}

\long\def\finalcomment#1{{\bf \textcolor{red}{\bf #1}}}
\newcommand{\CF}{\finalcomment}

\newif\ifdraft \drafttrue \ifdraft \newcommand{\katznote}[1]{
{\textcolor{magenta} { ***Dan: #1 }}} \newcommand{\rananote}[1]{
{\textcolor{cyan} { ***Omer: #1 }}} \newcommand{\omernote}[1]{ {\textcolor{cyan}
{ ***Omer: #1 }}} \newcommand{\note}[1]{ {\textcolor{red} { #1 }}}
\newcommand{\jhanote}[1]{ {\textcolor{blue} { ***Shantenu: #1 }}}
\newcommand{\jonnote}[1]{ {\textcolor{red} { ***Jon: #1 }}}
\newcommand{\alnote}[1]{ {\textcolor{darkgreen} { ***Andre: #1 }}}
\newcommand{\nchnote}[1]{ {\textcolor{orange} { ***Neil: #1 }}}
\newcommand{\ysnote}[1]{ {\textcolor{violet} { ***Yogesh: #1 }}}
\newcommand{\sdnote}[1]{ {\textcolor{blueish} { ***Simon: #1 }}} %
\usepackage[pdftex,colorlinks=true, linkcolor=blue,citecolor=blue, %
urlcolor=blue]{hyperref} \else \newcommand{\katznote}[1]{}
\newcommand{\rananote}[1]{} \newcommand{\note}[1]{} \newcommand{\jhanote}[1]{}
\newcommand{\jonnote}[1]{} \newcommand{\alnote}[1]{} \newcommand{\nchnote}[1]{}
\newcommand{\ysnote}[1]{} \fi

\begin{document} \runningheads{S.~Jha et al.}{Understanding Applications and
Infrastructure for D3 Science}

%\title{3DPAS: Distributed Dynamic Data-intensive Programming % Abstractions and
Systems for Large-Scale Science and Engineering}

\title{Introducing Distributed Dynamic Data-intensive (D3) Science:
Understanding Applications and Infrastructure}

% An Applications-Driven Analysis}

%\title{Towards Distributed Dynamic Data-intensive (D3) Science: A % Survey and
Analysis of D3 Science Applications and % Cyberinfrastructure}

\author{Shantenu Jha\affil{1,2}\corrauth\, Neil Chue Hong\affil{2}, Simon
Dobson\affil{3}, Daniel S. Katz\affil{4,5}, Andre~Luckow\affil{1}, Omer
Rana\affil{6}, Yogesh Simmhan\affil{7}}

\address{\affilnum{1}Rutgers University\break \affilnum{2}University of
Edinburgh\break \affilnum{3}University of St.~Andrews\break
\affilnum{4}University of Chicago\break \affilnum{5}Argonne National
Laboratory\break \affilnum{6}Cardiff University\break \affilnum{7}Indian
Institute of Science\break }

%\affilnum{9}University of Minnesota\break

\corraddr{Rutgers University, NJ, USA. E-mail: shantenu.jha@rutgers.edu}

% \author{Shantenu Jha} % \address{Rutgers University and University of
Edinburgh; \texttt{shantenu.jha@rutgers.edu}} % \author{Neil Chue Hong} %
\address{University of Edinburgh; \texttt{N.ChueHong@epcc.ed.ac.uk}} %
\author{Simon Dobson} % \address{University of St.~Andrews;
\texttt{simon.dobson@st-andrews.ac.uk}} % \author{Daniel S. Katz} %
\address{University of Chicago and Argonne National Laboratory;
\texttt{d.katz@ieee.org}} % \author{Andre Luckow} % \address{Louisiana State
University; \texttt{luckow@cs.uni-potsdam.de}} % \author{Omer Rana} %
\address{Cardiff University; \texttt{O.F.Rana@cs.cardiff.ac.uk}} %
\author{Yogesh Simmhan} % \address{University of Southern California;
\texttt{simmhan@usc.edu}} % \author{Jon Weissman} % \address{University of
Minnesota; \texttt{jon@cs.umn.edu}}

\keywords{dynamic, distributed, data intensive, scientific applications}

\begin{abstract} % Many problems at the forefront of science, engineering,
medicine, % and the social sciences are increasingly complex and %
interdisciplinary due to the plethora of data sources and % computational
methods available today.  A common feature across many % of these problems is
the amount and diversity of data and % computation that must be integrated to
yield insights. Such data is % increasingly large-scale and distributed: the
datasets, their % locations, and associated scheduling decisions are %
time-dependent.

% Applications need access % to these disparate data-sets and to the methods
that operate on % them.  Data-intensive applications need first-class support
for % distributed heterogeneous platform execution.

The 3DPAS Research Theme (funded by the eSI, Edinburgh and US NSF) investigated
applications that have these characteristics; in other words, it operated at the
triple point of {\it distributed} and {\it dynamic} and {\it data-intensive}
(D3) attributes.  This document represents the conclusions of the 3DPAS Theme.
\end{abstract}
\maketitle

\ifdraft
\newpage
\tableofcontents
%\newpage %\listoftables
\newpage \fi

%\input{notes}

\section{Introduction: Context, Scope and Outline}

\jhanote{NCH points out that we are inconsistent in the order of the ``D''s. SJ
to fix globally. Good catch Neil!}

The landscape of scientific and enterprise computing is being fundamentally
altered by the current surge in data. The resulting upheaval, grandiosely named
BigData science by the Lilliputians in the data science community, represents a
new reality where the volumes of data produced by scientific applications are
increasing rapidly, driven by advancing technologies (e.g., increasing compute
capacity and higher resolution sensors) and decreasing costs for data
acquisition and storage~\cite{tony-propaganda}. With the increase in data
volumes, types, and sources of data, an important concomitant development has
been an increase in the requirement for and complexity in the processing of
data. Data are often naturally distributed and have to be processed in situ, and
occasionally data that are localized have to be distributed in order to be
processed, either for reasons of performance or for collaborative analytics.
Furthermore when processing large volumes of data, data source/destination, data
production and transfer rates, and data processing capacity can all vary over
time. Thus properties such as resource scheduling, data placement, and transfer
decisions, that were statically determinable at small scales, emerge as dynamic
and important considerations at large scales of data and distribution.  The
BigData science problem is therefore only partly about large volumes of data.

Our focus is on a subspace of the BigData problem space, namely that which is
characterized by the distributed and dynamic aspects of data-intensive
science. We introduce the concept of Dynamic Distributed Data-intensive science
and refer to this subspace of BigData science as D3 science.  In order to
appreciate the D3 subspace we will briefly discuss the importance and role of
distribution and the pervasiveness of dynamical aspects.

% \begin{itemize}
% \item Centralized: data is at one location
% \item Decentralized: data is stored at multiple locations
% \end{itemize}

% \jhanote{I intend to convert these bullet points into a paragraph % about the
importance of treating distributed data}

Although the capacity to store data centrally is growing faster than Moore's
Law, it is not sufficient in all cases to be able to ``dump the data'' in one
place (or in the cloud.)  Oftentimes the costs (data transfer, storage,
consequences of catastrophic failure) prevent the centralization of data, and
decentralization---storing or place data in multiple locations---become prudent.

Conversely, not all production of data is centralized; in some cases there are
multiple producers (and consumers) of data, not to mention one-of-a-kind data
that must be shared across multiple consumers.  Furthermore in collaborative
science, data output from one analytical engine are often the input for the next
stage of the analytical pipeline.  Last but not least, distribution via
replication is a tested and effective approach to reliability and
fault-tolerance; we will argue that it is also potentially a great performance
enhancer and a means to ensure data consistency too.

Whether data are a priori distributed or localized, a critical challenge is to
find an efficient and effective means to either distribute the initially
localized data, or manage the distribution (or localization) of a priori
distributed data.

% \begin{itemize}
% \item Costs (data transfer, storage costs) for centralized data too % high
% \item data locally required for analysis (store data where it is used)
% \item volume too large to store centrally or move around
% \item required for scaling the system (load balancing, local reads are %
cheaper)
% \item reliability and fault tolerance (no single point of failure, %
redundancy)
% \item data consistency
% \item data security
% \end{itemize}

% \ysnote{distributed to ensure security also? Privacy % controls on moving data
out of a location or country.}

The challenges of distributed data-intensive applications span beyond the
management and curation of large volumes of data. A critical requirement is to
make sense of this data, i.e., transform the data into information en route to
assembling knowledge. Among other things, the transformation from data to
information requires extensive amounts of computational processing of the data;
thus the data-intensive nature of the BigData science problem is strongly
coupled with compute-intensive requirements.  As the data are often distributed,
a consequence is that many attributes of distributed systems permeate to
data. Thus almost naturally, distributed data have attributes of dynamism, both
at the data level as well as the infrastructure level.

That is not to say all data-intensive applications have elements of both data
and compute dynamism and are always distributed, but it is a reflection that at
extreme scales, dynamism and and distribution become increasingly important and
often correlated issues.  Thus, analyzing the applications along an analytical
space spanned by the three ``D''s of data-intensive, distributed, and dynamic
provides a general and effective framework.  The diverse set of requirements to
support D3 scenarios points to the need for a careful understanding of D3
characteristics as well as sophisticated, extensible, and skillfully architected
infrastructure and programming systems.

With recent attention on ``Big Data'' problems, several classification schemes
have been proposed; one that has garnered attention is the
3Vs~\cite{gartner3v}. These three dimensions (Volume, Velocity and Variety) just
describe the characteristics of the data. %, rather than describing the
interplay between Big %Data and applications that operate over them.
\jhanote{Does someone have the authentic reference for the 4Vs?} \katznote{see
http://www.villanovau.com/university-online-programs/what-is-big-data/ but it's
not `authoritative'} In contrast, D3 emphasizes the nature of scientific
applications as characterized by their data attributes. Furthermore the D3
attributes are considered as a collective property, as opposed to different
disjoint attributes.  Thus we believe there is a fundamental distinction between
the D3 and 3Vs classification schemes.

%\ysnote{Since we're clearly positioning around BigData, % one question that is
going to come up is how are the 3D's related to % the 3V's -- volume, velocity
and variety -- of Big Data. We may want % to explicitly address that early
on. For e.g. ``While the three % dimensions of Volume, Velocity and Variety of %
BigData~\cite{gartner3v}, or 3V's, are well recognized in industry, % these
narrowly consider the characteristics of the data rather than % the interplay
between Big Data and applications that operate over % them as we do through the
3D's. Furthermore, our emphasis is also on % the role of Big Data in science,
which is less studied (and arguably % more diverse and complex) than business
applications. That said, the % volume and velocity dimensions of the 3V's
partially capture the % data-intensive and dynamic aspects of our study.''}
\jhanote{we do not discuss explicitly the 4th D, i.e. diversity (courtesy AL) it
is assumed but not discussed}

%\subsubsection*{Terminology}

%\alnote{Should we introduce a degree of distribution? machine, % cluster,
data-center, WAN, Wireless} \jhanote{A figure akin to our % the figure in our
pilot-data paper maybe Ok to help establish the % vocabulary. Not sure this
would be a conflict or self-plagiarism?  % Either way, this should be in the
sub-section on ``Scope''}

Distribution can occur on many levels.  %, as shown in
%\figurename~\ref{fig:figures_terminology}.  In this paper, we use {\em locally
distributed} to indicate that data/compute resides on multiple nodes within one
data center.  {\em Geographically distributed} describes the distribution of
data and/or compute across multiple data centers. Further, we define the degree
as the number of data sources and sites that are used for storing and processing
the data. A distributed application is an application that is composed of
entities (i.e., either storage/compute resource or a data source) from at least
two physical locations. In this work, our focus is almost exclusively on
geographically distributed data and applications.

%\jhanote{terminology defined elsewhere in the paper isn't defined % here.. so a
bit misleading to call this part terminology. AL points % out ``dynamic'' as an
example.}

%\katznote{AL thinking about changes to the figure, according to SJ - % now SJ
to remove this figure completely.}

%\katznote{where are the data sources %in this figure?}\alnote{added}

%\begin{figure}[htbp] % \centering
% \includegraphics[width=0.7\textwidth]{figures/terminology.pdf}
% \caption{Levels of Distribution}
% \label{fig:figures_terminology}
%\end{figure}

% (\S\ref{sec:prog_infra}). \katznote{removed this ref, as it seemed % out of
place to have a ref to a subsection in the intro, as the only such ref}

% \subsection*{Distributed Data} % One specific attribute that we believe is
pervasive but not % necessarily obvious is ``dynamic data.''  Additionally the
requirement % of many applications impose a requirement that data is dynamic.

% \jhanote{I do not know why floating sensors was a subsection? As an % examplar
for D3 Science? If so, made it a subsubsection..}

% \katznote{as I wrote in the title, it got moved from apps and won't be there %
any more, but there was some thought that it might be useful as an example %
application here.  Feel free to delete it if you don't think it adds anything
here.}

% \subsubsection{Floating Sensors -- material than can be used in 1 as % an
example application that is not going to be discussed in %
\S\ref{sec:scenarios}}

% This {\em traditional application} constructs real-time hydrodynamics models,
and forecasts the behavior of a complicated water system: the Sacramento % - San
Joaquin River delta.  It begins after data generation and includes data
transformation \& decisions, storage, and processing.

% \subsubsection*{Application description}

% The application integrates both static and dynamic sensors streaming %
real-time data with a high-resolution hydrodynamics model running on % the NERSC
Magellan cloud testbed. The model uses sensor data to % initialize and shape the
run, and the model output is in turn used to % drive the real-time positioning
of the dynamic sensors. The output of % this process is a map of the current and
future behavior of the % modeled watershed.

% Custom built dynamic sensors are deployed in the watershed along with %
traditional statically placed sensors that provide data. These sensors %
communicate over a variety of different network infrastructures, e.g., % GSM,
CDMA, 802.11, wired ethernet, etc.  and the application acquires % the data over
the internet from the sensor platforms.  The application % processes the sensor
data in real-time as it is streamed.

% \subsubsection*{Big data aspects}

% Not known.

% \subsubsection*{Distributed aspects}

% Not known.

% \subsubsection*{Dynamic aspects}

% The sensor are all streaming real-time data to the application. The position
of the dynamic sensors % themselves change over time.

% \subsubsection*{Other important issues}

% The first full, end-to-end, field test of the system is scheduled for % May
2011.

%\jhanote{If some one knows a bit more about Bayesian %
approaches/inference/statistics and its 'implication' on dynamic % data via
dymamic models, i'd love to hear / learn more}

\jhanote{refine next two paragraph...} Analogous to the distinction between
programming in the large versus system level programming, there is a distinction
between the ``cyberinfrastructure/systems considerations/design at the
large-scale'' versus ``CI considerations at the small-scale''. This paper is
primarily about CI/systems considerations at the large-scale.  Searching for
common solutions and patterns at the large-scale is likely to be more fruitful
than at small-scales where customized tools and optimization of low-level
systems features is likely to return greater yield.

Thus, subjecting a range of otherwise distinct applications to a common
analytical framework provides advantages to the infrastructure scientists and
cyberinfrastructure experts, namely suggestions for where effort is both needed
and likely to yield maximum returns. In other words, as D3 applications evolve,
our analysis can provide guidance on where to focus effort on infrastructure.
Last but not least, the advantages of providing a common but extensible
vocabulary that can be used by the entire community, application end-users,
developers and infrastructure scientists alike, in an emerging field cannot be
overstated. One may argue that such vocabulary and terminology has been missing
from the practice of distributed cyberinfrastructure in general, with less than
desirable consequences.

%\katznote{some of this is repeating ideas from the para ``An important %
motivation...''  Perhaps these two paras should be % combined?}\jhanote{DSK,
this should not be an issue now, given that % this has now moved upstream}

\subsection{Scope}

% Before a discussion of the applications, infrastructure and research %
methodology, a brief mention of the motivation and thereby scope of % this work
is in order.

% Without a doubt, the era of {\it BigData} is here. However, BigData % is not
just an issue of large volumes of data, it is also about the

%One motivation of this work is to understand the complexity and
%characteristics of data that emerge as a consequence of scale, namely %the
dynamic properties and the distribution of applications, and the %complex
interplay between them that arises.

One motivation of this work is to understand the dynamic properties and the
distribution of applications, and the complex interplay between them that arises
as a consequence of scale.  Additionally, we are motivated by an attempt to
understand infrastructure---existing capabilities, trends, and
limitations---available to the complexity, challenges and characteristics of
dynamic and distributed data at scale.

We would like the reader to keep in mind three primary context and scope-setting
points:

%\katznote{a lot of this paragraph seems too repetitive of the previous %
section} \jhanote{Dan, I fixed the previous paragraph. OK?}

\begin{itemize}

\item This {\it survey} is not meant to be complete, either in scope of domains
covered or types of infrastructure and application surveyed.  They by definition
represent the interests and experience of the authors.

\item The focus is on applications and infrastructure that arise from science
and engineering projects in academia; this is not exclusive but is more a
guiding principle.  While we are cognizant of enterprise applications and
infrastructure, and recognize the impressive advances made therein, we also
acknowledge that the factors such as scale, distribution, open-source, and
interoperability influence the solutions employed and infrastructures used; it
is important to therefore focus on problems that emerge as in the space defined
by these ``academic constraints''.

\item Almost by definition, any attempt to capture the state-of-the-art of such
a dynamic and emerging domain is bound to be fraught with limitations and issues
of scoping.  A complete and rigorous survey of the landscape of BigData is bound
to be obsolete by the time it emerges; so given the rate of change, we asked
what could be done. What we present is a partial analysis motivated by a set of
existing and active D3 applications, their trials and tribulations, successes
and solutions. We believe there is merit in capturing the state-of-play as is,
deriving a research agenda and set of specific yet crosscutting issues, and
presenting these to the community.
\end{itemize}

%\jhanote{I think these two points can be ignored. To be discussed. But % either
way this should go in the scope subsection} % \katznote{Dan and Shantenu request
that Neil and Omer think % about the next two points, and either flesh them out
with text % or remove them.}
%
%\begin{itemize}
%\item How does this differ from Scientific Databases? How does it % relate to
Scientific Data Management and Scientific data flow
%\item Reference to EU 2030 Objectives (part of the Project Europe 2030 %
Report) -- underpins ``Data Intensive Research'' from an EU % perspective
(strategy) + relate this to NSF ``CIF21'' vision
%\end{itemize}

%\jhanote{some points to ensure we've addressed. I cut and paste some % comments
we received from the DPA survey paper that we should be % ensure at addressed --
implicitly or explicitly}
%
%\begin{verbatim}
%
%(i) It is difficult to argue about the completeness of the presented %material.
%
%(ii) Picking a handful of distributed applications to study does not form a
%basis of a taxonomy.
%
%(iii) Furthermore, it is unclear if the applications %are indeed representative
of any urgent need to execute across %distributed systems as the authors claim.
%
%(iv) At the same time, there is not enough description of the computer %science
behind the applications.
%
%(v) Several applications seem similar to each other and, %as far as I can tell,
do not add anything new.
%
%(v) The section is more a %laundry list of applications - I would have expected
to see a %synthesis of different application and, for example, a classification
%into different classes. No general requirements of applications are %derived. A
smaller number of more representative sample applications %(not covering nine
pages) would have been better.
%
%(vi) My main problem with the paper is the research methodology.  The paper
%uses 6 hand-picked applications to derive a vector of attributes for %each of
them, derive some patterns from observing these vectors, and %uses these
patterns to perform a distributed computing tools gap %analysis.
%
%(vii) It is unclear how important the applications used in this %study are.
%
%(viii) Your lack of depth in explaining the applications leaves the %reader
empathetic.  It's important that the reader feels that they can %learn something
from reading this paper.  How is what you are saying %applicable to me (a
computer scientist or computational domain scientist)?
%
%(ix) Have you considered adopting a different approach, like conducting a
%survey to glean the likes/dislikes of distributed application %developers?  In
essence focus on the development process, as opposed %to applications.  With a
large enough pool of responses, you may come %up with statistically significant
conclusions that will be of interest %to the wider community.
%
%\end{verbatim}

\subsection{Methodology}

This paper originated in work at the UK e-Science Institute (eSI) at the
University of Edinburgh, in a research theme examining Dynamic Distributed
Data-intensive Programming Abstractions and Systems, called
3DPAS~\cite{3dpas-theme}.  The description of this theme's work was:

\begin{quote} Many problems at the forefront of science, engineering, medicine,
and the social sciences, are increasingly complex and interdisciplinary due to
the plethora of data sources and computational methods available today. A common
feature across many of these problem domains is the amount and diversity of data
and computation that must be integrated to yield insights.

For many complex data-intensive applications, moving the data may have
restrictions. Increasingly important type of data-intensive applications are
data-driven applications. For example, data is increasingly large-scale,
distributed arising from sensors, scientific instruments \& simulations. Such
data-driven applications will involve computational activities triggered as a
consequence of independent data creation; thus it is imperative for an
application to be able to respond to unplanned changes in data load or
content. Understanding how to support dynamic computations is a fundamental, but
currently a critical missing element in data-intensive computing.

The 3DPAS theme seeks to understand the landscape of dynamic, distributed,
data-intensive computing: the programming models and abstractions, the run-time
and middleware services, and the computational infrastructure. It will analyze
existing tools and services, identify missing pieces and new abstractions, and
propose practical solutions and best practices.
\end{quote}

The theme held three workshops, and in each, application scientists and
computing technologists were invited to give presentations on their work, and
then to be involved in discussions to organize and understand the presented
materials.  Between the workshops, the theme organizers also drafted outlines of
a report that was a predecessor of this paper, for discussion in the next
workshop.  Overall, there were about 50 workshop attendees.

In order to select applications and gather information about them, during the
various workshop discussions, when an application that ``felt'' new was
mentioned, we asked the person who mentioned it to provide a written set of
answers to the following questions about the application:

\begin{enumerate}
\item What is the purpose of the application?
\item How is the application used to do this?
\item What infrastructure is used? (including compute, data, network,
instruments, etc.)
\item What dynamic data is used in the application?
\begin{enumerate}
\item What are the types of data,
\item What is the size of the data set(s)?
\end{enumerate}
\item How does the application get the data?
\item What are the time (or quality) constraints on the application?
\item How much diverse data integration is involved?
\item How diverse is the data?
\item Please feel free to also talk about the current state of the application,
if it exists today, and any specific gaps that you know need to be overcome.
\end{enumerate}

We then used the text of these answers to describe each of the applications (in
\S\ref{sec:scenarios}) in terms of:

\begin{itemize}
\item What does the application do? What problem does it solve?
\item How does it do it? How does the application run?  What infrastructure does
it need/use?
\item What are the big data aspects of the application?
\item What are the distributed aspects of the application?
\item What are the dynamic aspects of the application?
\item What else is important about the application?  What does it do well?
Poorly? How was the app written? What tools does it use?
\end{itemize}

After the workshops, the theme organizers and a set of attendees who were
interested in participating started discussing how to organize and analyze the
applications and the technologies, including infrastructure, programming systems
and abstractions.  These participants are the current authors of this paper, and
this paper is the result of that process.

As part of the theme, two 3DPAS workshops~\cite{3dapas,D3-escience} were
organized at major conferences: HPDC 2011 and IEEE eScience 2011.  These
workshops had both peer-reviewed and invited papers.

A combination of invitation-only focused discussion meetings and broader
community engagement events were used to ensure that the applications surveyed
and ideas captured in this paper are representative of current community
thinking as well as of intellectual relevance.

\subsection{Outline of this paper}

The rest of this paper consists of the application scenarios
(\S\ref{sec:scenarios}), a discussion of the aspects of distribution and
dynamism in the applications (\S\ref{sec:distdyndata}), a description of the
software infrastructure required for D3 applications % Yogesh: This section's
description seemed more verbose than others, hence commenting out details % ,
such as % software frameworks, tools, and middleware and platform services
(\S\ref{sec:infrastructure}), analysis and characterization of the applications
and conclusions (\S\ref{sec:conclusions}).

\section{D3 Application Scenarios \label{sec:scenarios}}

Starting with the experience of the attendees of the 3DPAS workshops, we have
assembled a set of D3 ``applications'' that we describe in this section.  We
realize that we have overloaded the term application here.  In this document,
some of our applications are really {\em traditional applications}, meaning a
program that is developed by a user or project to try to answer a science
question, usually with some input data needed and some output data produced.
Applications in this category are: NGS Analytics (\S\ref{bioSilvia}), CMB
(\S\ref{astroJulian}), Fusion (\S\ref{fusionScott}), Industrial Incident
Notification and Response (\S\ref{Kees}), MODIS Data Processing
(\S\ref{modisKeith}), and Distributed Network Intrusion Detection
(\S\ref{detectionJon}).

Another group of applications are {\em infrastructural applications}.
Applications of this type are collaborative, in that they depend on a context
that is often agreed upon by a community, most often in terms of how data is
stored by the community and how it is accessed, as well as, more often than not,
a weak consensus on the infrastructure to be used. %\jhanote{Dan: Please confirm
you are OK with this modified % defintion of infrastructural application} This
allows different sub-applications (that we again refer to as `applications') to
focus on different stages, such as generating data, or processing data.
Answering a science question now may involve a set of applications that need to
be run in series, perhaps in different stages, that may be run by different
groups that do not frequently interact.  This type of application includes:
ATLAS/WLCG (\S\ref{WLCGSteve}), LSST (\S\ref{astroAdam}), SOA Astronomy
(\S\ref{astroSOAAdam}), Sensor Network Application (\S\ref{sensorSimon}),
Climate (\S\ref{climateDan}), Interactive Exploration of Environmental Data
(\S\ref{envJon}), and Power Grids (\S\ref{power}).

In other words, multiple infrastructural applications are used together in
particular science domains.  Furthermore, infrastructural applications involve
more than just the end user (scientist), and typically require exerting control
at layers below, i.e., ``programming the system (infrastructure)'', as opposed
to a traditional applications, where end user effort and control is often
confined to a well-defined application kernel.

% \jhanote{Need better distinction between infrastructure and % traditional
applications. They are the same at the scale-up and % scale-out level, but
different in scale-across and people. Another % issue is that about ``control
and configure'', i.e traditional % applications only exert control and configure
the application layer, % whilst infrastructure applications need to configure
the lower % ie. infrastructural layer too}

% \katznote{I don't agree in general that more is needed, but if you % think so,
please give it a try.  I also don't agree specifically % that infrastructure
applications `configure' the lower layer.  % Rather, I would say they assume a
particular, community-defined % (perhaps de facto) layer.  Finally, while I
understand scale-up % and scale-out, and I don't know what `scale-across' is.}

Some of the applications we describe are really groups of applications, meaning
that we are grouping together a number of independent programs, written by
different authors, that are trying to answer the same type of science question
and thus, where a single user will only run one of them at a time.
%\katznote{Shantenu, please check %the preceding and see if this is ok, in the
context of the whole %paragraph and section.}  These codes and their developers
may either be competing or collaborating, and sometimes both, informally known
as collabetition. Specifically, this is the case for NGS Analytics
(\S\ref{bioSilvia}), SOA Astronomy (\S\ref{astroSOAAdam}), and Interactive
Exploration of Environmental Data (\S\ref{envJon}).  (Note that this issue is
orthogonal to traditional vs. infrastructural applications: NGS Analytics is
traditional, where SOA Astronomy and Interactive Exploration of Environmental
Data are infrastructural.)  In each of these examples, we have chosen one
specific application that is meant to represent a possibly large number of
comparable applications.

%\jhanote{Dan: This paragraph seems to suggest some % similarity with
infrastructural applications, but doesn't quite do % so. Can we make a bit more
explicit the intent. For example, If we % break into ``collaborating'' vs
``competing'' and put NGS into a % ``competing'' category the remainder of the
paragraph above would be % a useful and fair elaboration on infrastructural
applications?  % Reading forward, Section 5.1, it appears to be consistent with
the % ``observations'' section. For example, `` Some of the applications % we
describe are really groups of applications, meaning that we are % grouping
together a number of independent programs, written by % different authors.
These independent programs might be competing % means to answer the same type of
science question and thus are not % intended to run together, or they could be a
set of smaller % applications collaborating to answer the same overall science %
question.''}  \katznote{some changes made above.  I'm not sure why % NGS is
competition and the others are collaboration - to me they are % all a bit of
both.} \jhanote{NGS is classified as a traditional % application; others as
infrastructural and as having multiple % distinct applications. Thus it is
natural to have collaborative.  % But as NGS is said to be a traditional,
different applications can % be perceived to be competitive. I agree that the
term competitive is % not perfect} \katznote{I've added a note to say that this
(applications % selected to represent groups) is orthogonal to traditional vs %
infrastructural.  (This is actually why this is a different paragraph.)  % Does
this help?}

%\katznote{not sure which category is right for: Power %Grids (\S\ref{power})
Yogesh, what do you think?}  %\ysnote{The infra category is appropriate. i.e. no
change required}

%\alnote{After reading the definition a couple of times, I still have a
%difficult time to distinguish between infrastructural and archetypical
%applications. The definitions sound very similar.}  %\katznote{this section has
been re-written.  I've removed archetypical applications, and added a paragraph
at the end of this section (2.0) to mention this idea without naming it.  Please
see if this is now better.}  %\alnote{reads very well!}

The discussions in the workshops and our analysis of the applications led us to
examine the use of data in different parts of the applications, where we found
some common patterns across applications.  For example, applications that
involve streaming data collect data from sensors, transform and filter the data,
store the results, and then later process (analyze) the stored results.  In
general, we think that all of the applications we have studied can be mapped to
a set of stages or phases.  This is illustrated in
\figurename~\ref{fig:figures_application-stages}.

The question of where the application, as opposed to the entire system, %(where
the system is defined as all of \figurename~\ref{fig:figures_terminology}),
actually starts is somewhat difficult.  In this paper, we have decided not to
treat hardware data sources (e.g., sensors, telescopes) as part of the
applications.  Therefore, these applications generally start with the data
movement that follows the actual data generation.

Unlike the sensor applications, another set of applications generate data
computationally, for example, from a model or simulation.  These applications
then do start with data generation, and they often have just three stages, where
data is generated in the first stage, stored in a second stage, and then
processed in a third stage (i.e., the transformation \& decisions stage in
\figurename~\ref{fig:figures_application-stages} does not occur).

Other applications may operate on stored data (i.e., they only contain the
processing stage and its associate data movement).  %\alnote{We should refer to
these application stages from the % application descriptions. Currently, this
classification is not % used.}  \katznote{have now done, but this was harder to
do than I % thought it would be, particularly to define where the application %
starts - for example, LSST as an app does not include data % generation, but
starts after this (data transformation \& % decisions), and SOA Astronomy does
not include data storage, but % does include the movement from the storage.
Perhaps I need to go % back and make the fact that many apps start with data
movement more % clear.}\alnote{ok}

\begin{figure}[htbp] \centering
        \includegraphics[width=1.0\textwidth]{figures/application-stages-new.pdf}
        \caption{Application Stages: The work in data-intensive applications can
be put into four common application stages: data generation, transformation \&
decisions, storage, and processing, though not all applications have all stages.
% \alnote{We have some applications (e.g. sensor networks) that give feedback to
the % data generation stage. Do we want to refer to this possibility here?}  %
\alnote{data transformation and decision making sounds to me similar to
processing} % \katznote{what do you suggest?}\alnote{The proposal would be to
just call it Data Processing as before. Since % there are many forms of
processing (transformation being just a specific form of processing)}
\katznote{I want to distinguish between the processing that happens before
storage % and the processing that happens after storage, so I am thinking of not
changing this.  % Would this really bother you?}
                \label{fig:figures_application-stages}}
\end{figure}

%\katznote{Dan to clean up the description of app phases in the intro and the
apps, %explain where apps start, with data movement, data generation, data
storage, etc. -- %then ask Shantenu to read and comment on this.}

%\katznote{think about linking this with 4.1}

%Application Scenario Characteristics for Dynamic Data (Various)
%
%Identify: (1) what is the case scenario within the application; (2) what data
is dynamic; (3) how the data is being used in the context of each of these
scenarios.
%
%\katznote{proposed structure for rewriting these entries:\\ %1. what does the
application do - what problem does it solve (question 1 below) \\ %2. how does
it do it (questions 2,3,5,6 below, also address cyberinfrastructure (where does
the app run)) -- perhaps have a diagram here of what happens in the app \\
%3. what are the big data aspects (question 4b below) \\ %4. what are the
distributed aspects (question 3 below)\\ %5. what are the dynamic aspects
(question 4a below)\\ %6. what else is important (questions 7,8 below, also
address programming systems (how was the app written; what tools does it use))\\
%}
%
%
%Currently trying to do this by asking the following questions:
%\begin{enumerate}
%\item[1] What is the purpose of the application?
%\item[2] How is the application used to do this?
%\item[3] What infrastructure is used? (including compute, data, network,
instruments, etc.)
%\item[4] What dynamic data is used in the application?
%\begin{enumerate}
%\item[4a] What are the types of data,
%\item[4b] What is the size of the data set(s)?
%\end{enumerate}
%\item[5] How does the application get the data?
%\item[6] What are the time (or quality) constraints on the application?
%\item[7] How much diverse data integration is involved?
%\item[8] How diverse is the data?
%\end{enumerate}
%
%Please feel free to also talk about the current state of the %application, if
it exists today, and any specific gaps that you know %need to be overcome.

%\subsection{Silvia's Biosciences app: NGS, medical imaging
(Silvia)\label{bioSilvia}}

\subsection{Next Generation Sequencing (NGS) Analytics \label{bioSilvia}}

The (grand) challenge in bioinformatics is the need to provide support for the
analysis of genome sequence data that is becoming available due to the abundance
and high-throughput capability of the next-generation sequencing devices. The
rate of growth of data from the NGS machines humbles Moore's law; it has
increased by more than three orders of magnitude in barely a decade.

There are many different types and components of analyses of NGS data.  However,
fundamentally important to all types is the ability to efficiently and
effectively map/align the short reads to a reference genome.

Alignment underpins multiple scientific questions such as comparing metagenomic
(DNA sequence) data sets against each other, which could be used to distinguish
human disease biomarkers among different individuals, determine different soil
(or water or air or ...)  microbial composition from different samples, etc.

This subsection describes a set of {\em traditional applications} that are used
in the processing of next generation sequencing data.  The applications have two
of the possible stages: data storage and data processing.
%\katznote{Shantenu/Yogesh, can you confirm this or fix it?  % I'm sure data
generation is not part of this app, but am not sure % about transformation \&
decisions. - see % Figure~\ref{fig:figures_application-stages}.} \jhanote{Dan:
I'm % confused. Not sure I understand what you mean by data-generation is % not
part of this app?}  % \katznote{the sequencer is not part of the application -
the application % starts once the data has been generated, thus it starts on a
data movement % arrow.  I'm just not sure which one.}\alnote{I think the phases
storage and % processing are fine for NGS}

\subsubsection*{Application description}

There are multiple tools (programs) that can perform the mapping/alignment
phase, e.g., BWA, BFAST, Bowtie, MAQ, etc. Each has its own strength and its own
native data support/capability; they also implement different alignment
algorithms. Thus, based upon different design objectives and trade-offs, they
typically have different constraints and efficiencies.

Several of these tools are also composed together as a linear pipeline for
performing sequencing and analysis. The pipeline may go through alignment, SNP
calling, haplotyping, and so on, with different tools used in each stage. These
pipelines can often be performed independently on each chromosome to provide
simple parallelism.

All algorithms, architectures, storage, and networks have broken or will break
soon.  Some NGS analytics are not good fits for core-rich and memory-poor modern
architectures. Some problems (e.g., comparing k-mers across metagenomic data
sets) need terabyte-scale memory, not thousands of cores.~\cite{ngs-gap}

Individual data sets, which will be terabyte-scale, are generated by NGS
machines at a variety of locations. There could be many thousands of such sets
to compare. Having them all local may not be feasible (the transfer time from
where they are generated may be prohibitive).

\subsubsection*{Big data aspects}

Individual data sets will be terabyte-scale. For example, for a standard genomic
sequence, there are billions of short (35-100 bases) reads of DNA.  And there
could be many thousands of such sets to compare for a given problem.  Getting
more accuracy for sequencing may require multiple coverage (e.g. 10x), that can
cause sequence sizes to increase linearly.  This makes NGS data management
fundamentally a big data problem.  NGS analytics, in addition to being a big
data problem, is also a computationally demanding and distributed computing
problem.  The computational demands arise from the often complex and intensive
analysis that has to be performed on data, which in turn arise from algorithms
that are designed to account for repetitions, errors and incomplete information.

\subsubsection*{Distributed aspects}

The type of NGS research being conducted can distinguish the distributed
properties of the application. Researcher groups that own sequencing machines
need to deal with large dataset sizes that need to be resequenced and reduced
for further analysis. Other researchers may analyze large catalogs of sequenced
genome databases are maintained, and perform pattern matching. Further, clinical
may be interested in in-depth analysis of an individual sequence for disease
diagnosis or visualization.

The distributed computing aspects arise at multiple levels: for example, the
simple act of having to move data from source (generation) to the destination
where computing (analysis) will occur is a challenge due to the large volumes
involved. Trade-offs exists between the cost/challenges in distributing data
versus I/O saturation or memory bottlenecks.

When coupled with the compute intensive nature of the problem, it soon emerges
that a fundamental challenge is not only whether to distribute, but where to
distribute, what to distribute (should the computing move to the data, or the
data move to the compute), and how to distribute (what tools and infrastructure
to use).

\subsubsection*{Dynamic aspects}

The dynamic challenges of NGS data are more subtle; the data themselves are not
dynamic -- the sequences are acquired once and analyzed many times. However the
execution of these applications on distributed infrastructure have dynamic
aspects when optimized resource usage is considered. For example, workload
decomposition and distribution must be determined dynamically in order to
optimally use resources, e.g., compute resource selection can be based on data
location, processing profile and/or network capacity.

\subsubsection*{Other important issues}

There are several toolkits that have emerged for NGS providing incrementally
better algorithms for sequencing and analysis \cite{SAMtools, BioJava, BioPerl,
.NETBio}.  However, as a consequence of the broad range of infrastructural
requirements, there isn't currently a common ``standard'' cyberinfrastructure
used for NGS analytics; researchers use what is easily available to them, or try
to use COTS infrastructure. Historically, leaders in the field have developed
infrastructure for the larger community to build-upon, e.g., QCDOC or
MD-GRAPE. But given the broad range of requirements, a hardware-only solution is
unlikely for a broad range of problem instances, and a combination of hardware
and software approaches will be employed. This reinforces the need for flexible
programming systems and associated runtime environments that support collective
(hardware-software) optimization.

Nowadays, NGS sequencers are even being shipped with servers or small clusters
for performing the alignment and resequencing as a pipeline right after the
reads have been generated~\cite{ilm}.  %\alnote{What kind of codes are used %in
these clusters? What are the characteristics of these codes?}  %But any
additional analysis is not supported.  %\alnote{What would be an additional
analysis}
%
Scientific workflow systems like Taverna~\cite{Taverna:2006},
BioKepler~\cite{biokepler:2011}, and Galaxy~\cite{galaxy} (though Galaxy does
not support distributed systems currently) have been successful in integrating
genome analysis tools for bioinformatics. Script based pipelines that combine
executables are also used often. Very few workflow systems have explicit support
for distributed data or dynamic data handling.  Therefore reusable approaches at
extending existing tools to support distributed and dynamic data are likely to
be useful.  %\katznote{Shantenu/Yogesh/Andre, please read this paragraph and
tell me if changes %should be made to this section at this point}\alnote{ok with
me}

%\alnote{Some details on how distributed/dynamic data is handled in % these
workflows would be good.}

%\katznote{from Silvia Delgado Olabarriaga, Tom Slezak, and Shantenu}
%
%\katznote{summary below assigned to Shantenu}

%\subsubsection{new summary}

%\katznote{Please make a new title for this section}

% \katznote{I was assuming you would leave the questions here, so that % we
could pull some of this material into a table later...}

% 1. {\bf What does the application do? What problem does it solve?}

% \katznote{from answer to question 1}

%2. {\bf How does it do it?}

% \katznote{from answers to questions 2,3,5,6 % below, also address
cyberinfrastructure (where does the app run) -- % perhaps have a diagram here of
what happens in the app}

%\ysnote{The first para of this section reads more like ``what does the
application do'' rather than % ``how''. Should it move above? Also, the
description of cores/memory above can move here as it % relates to CI. }
%\ysnote{This section will benefit from describing the (often) linear pipeline
of % sequencing, with individual chromosomes that can be processed
independently.}  %\katznote{Yogesh or Shantenu, can you do address Andre's
comments in this section?}  %\ysnote{I've taken a pass. Shantenu, do you care to
make changes?}  % 3. {\bf What are the big data aspects?}  \katznote{from answer
to % question 4b below} 4. {\bf What are the distributed aspects?}  %
\katznote{from answer to question 3 below} \ysnote{distinguish % between
individual research project where compute can move to % data; clinical testlabs
that sequence and transmit sequenced data % to end user for post processing;
collaboratory projects like % 1000-genome where data has to be aggregated
centrally. Sequencing % done remotely but visualization done locally.}  %5. {\bf
What are the dynamic aspects?}  %\katznote{from answer to question 4a below}
%6. {\bf What else is important?}

%\katznote{from answers to questions % 7,8 below, also address programming
systems (how was the app % written; what tools does it use)}

%\ysnote{See biohpc.net, bioperl, biojava, bio.net} %\katznote{I'm not sure what
action goes with this comment.  Yogesh?}  % It is unlikely that researchers or
NGS teams will build integrated % customised infrastructure for no
cyberinfrastructure to solve this % problem. The solution to

% \katznote{do we want to retain the figure? - currently commented out}

% In general this is a large (for biology -- don't think in terms of %
exascale!) data problem with large (all-vs-all comparison) % computing. There
are strategies to make this more scalable that will % reduce the computing
volume.

% i) large (exascale) data and large (exascale) computing (e.g., data % mining,
analytics, bioinformatics)

% ii) large (exascale) data and modest computing (e.g., search)

% iii) modest data and large (exascale) computing (e.g., % simulation/modeling)

% Potentially peta-bytes or exa-bytes of data. Not sure how to quantify % the
computing. (Huge memories will be potentially more valuable than % huge numbers
of cores -- although this is heresy to the physics % crowd.)

% Computing might then have to be % distributed. Biology is wrestling with this
now.

% Although only the alignment step is computation-intensive, it is % more
practical to perform all the others on the grid to minimize data % transport
to/from the user's workstation, which is normally on low % speed networks. That
is, the computation is statically and manually % taken to the data in this
case. This can lead to sub-optimal resource % usage.

% Data is not really `dynamic' in this case -- the sequences are % acquired once
and analyzed many times. However the execution of this % application on a
distributed infrastructure has dynamic aspects when % optimized resource usage
is considered:

%\begin{itemize}
%\item Split/merge the data according to available resources
%\item Compute resource selection based on data location, processing profile
and/or network capacity
%\item Resource selection based on security constraints (access control, trust)
%\item Collection and processing of monitoring information to steer grid
enactment.
%\end{itemize}
%
%
%\subsubsection{old material follows}
%
%\begin{itemize}
%\item Infrastructure monitoring for the execution of workflows to % manage
failure
%\item the application is sequence alignment (split data (this is where there is
dynamic decision making) to execute the alignment in parallel)
%\item DTI Imaging: two cases are interesting: DTI atlas (split, run, merge,
split, merge - reduction from 10000 tasks to 1 output); PCA for classification
of patients/control
%\end{itemize}
%
%\subsubsection{Silvia's contribution}
%
%NGS sequence alignment on the AMC e-bioscience infrastructure
%
%1. What is the purpose of the application?
%
%The application performs analysis of DNA sequencing data for various %goals in
life science research, e.g., Mutation screening, Virus %discovery, Genome-wide
Associations (GWA), Linkage analysis, %Comparison of bacterial genomes, microRNA
expression and Exome %sequencing. The data analysis is composed of various
methods that are %implemented by a bioinformatician, whereas the data is
typically owned %by the life scientist.
%
%2. How is the application used to do this?
%
%Note: Here we focus on data analysis from the perspective of the
%bioinformatician, excluding steps related to data acquisition %(transformation
of raw images into sequences of amino acids) and %posterior biostatistics and
epidemiological analysis.
%
%The data analysis is implemented as a pipeline of generic and %customized
components, mostly third-party public software tools. The %pipeline may differ
for each specific study, depending on the data %acquisition and the research
question. Typically it includes data %preparation (file conversion, reformatting
and filtering); alignment %(comparison) with some reference database as the
human genome; %post-processing (output conversion, reformatting, statistical
%analysis) and visualization. Sequence alignment is the most
%computing-intensive step, which became prohibitive for sequential %execution in
next generation sequence experiments. All data is stored %in files that are
often completely read in memory for fast %processing. Compressed formats are
usually adopted. Although the files %are large, the data can be easily
decomposed because the processing is %done on each sequence (`string') at a
time.
%
%3. What infrastructure is used? (including compute, data, network,
%instruments, etc.)
%
%\begin{figure*}[htbp]
% \begin{center}
% \includegraphics[width=\textwidth]{figures/silvia.pdf}
% \caption{Figure from Silvia}
% \label{Fig:silvia}
% \end{center}
%\end{figure*}
%
%see Figure~\ref{Fig:silvia}.
%
%Data:
%
%Data acquisition: Roche, Solid and Illumina next generation sequencers
%
%Local storage: dedicated server at the AMC network (ftp)
%
%Grid storage: Storage Elements of the Dutch e-Science Grid %(www.biggrid.nl)
%
%Data transport:
%
%From sequencer to server: AMC network (XXX) or off-line media
%
%Between data server and grid storage: public network (XXX); lightpath %(near
future)
%
%Between grid storage and worker nodes: public (XXX) or site (XXX) %network
%
%Computing:
%
%Local Computing: clusters and servers at the AMC and workstations %(laptops,
desktops) anywhere Grid Computing: Dutch Grid %(www.biggrid.nl), part of EGI
%
%Originally all steps were performed in some local server or on the %user's
workstation. The visualization is still performed in the user's %(windows)
workstation, but all the other steps are currently performed %on the Dutch grid
infrastructure. The pipeline is described as a grid %workflow using the GWENDIA
language %(\url{http://dx.doi.org/10.1145/1645164.1645171}) and enacted on the
%Dutch Grid using the MOTEUR2 workflow engine
%(\url{http://modalis.polytech.unice.fr/softwares/moteur/start}).
%
%PS: although only the alignment step is computation-intensive, it is %more
practical to perform all the others on the grid to minimize data %transport
to/from the user's workstation, which is normally on low %speed networks. That
is, the computation is statically and manually %taken to the data in this
case. This can lead to sub-optimal resource %usage.
%
%4. What dynamic data is used in the application?
%
%Note: data is not really `dynamic' in this case -- the sequences are %acquired
once and analyzed many times. However the execution of this %application on a
distributed infrastructure has dynamic aspects when %optimized resource usage is
considered:
%
%\begin{itemize}
%\item Split/merge the data according to available resources
%\item Compute resource selection based on data location, processing profile
and/or network capacity
%\item Resource selection based on security constraints (access control, trust)
%\item Collection and processing of monitoring information to steer grid
enactment.
%\end{itemize}
%
%(a) What are the types of data,
%
%DNA sequences (strings) stored in open formats (FASTA, SFF, SAM, BAM) %Bundles
of results from alignment tools (custom format, binary?)
%
%(b) What is the size of the data set(s)?
%
%one example for BWA alignment tool (short sequences) follows:
%
%input data: Sequencing Data in the csFasta format, normally 25-35~GB; %Quality
files in the .qual format, normally 50-80~GB; Reference DB in %the Fasta BS
format: 3.2~GB (human genome) 140~MB (one chromosome)
%
%Intermediate files: Reference BWA index, normally 4.5GB (human genome) %240~MB
(one chromosome); Sequencing Data in the FastQ format %(fastq.gz), normally
20-30~GB
%
%Output: Results in .sai format (direct output of BWA), normally %2-3~GB;
Results in .sam format, normally 55-75~GB; Results in .bam %format, normally
20-30~GB
%
%5. How does the application get the data?
%
%From a local file. The legacy code is wrapped to stage the inputs and %outputs
in/out the worker node.
%
%All data is stored in files that are often completely read in memory %for fast
processing. Compressed formats are usually adopted. Although %the files are
large, the data can be easily decomposed because the %processing is done on each
sequence (`string') at a time
%
%6. What are the time (or quality) constraints on the application?
%
%In research there are no strict constraints on the time to complete %the
analysis.  In patient care, time could determine life, death or %money, but we
have no concrete case yet.
%
%Quality in this case refers to the accuracy of the results %obtained. This is
essential for research and patient care.

\subsection{ATLAS (an Example of Use of the WLCG) \label{WLCGSteve}}

%\katznote{from Steve Fisher}
%
%\katznote{summary below assigned to Dan}
%\subsubsection{new summary}

%1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

The Large Hadron Collider (LHC)~\cite{lhc} at CERN in Geneva is producing a
large amount of data across a number of experiments, including
ATLAS~\cite{atlas}. The application discussed in this section is an {\em
infrastructural application} (meaning that it consists of multiple stages that
are run by different people, in this case data acquisition, storage,
distribution, and analysis). It has the goal of allowing a physics group or a
user to analyze data to understand a specific physics channel (underlying
physical process) as recorded by ATLAS.  The set of available data grows
steadily over time as more data are collected (by the experiment) and
reconstructed. This application thus contains two stages: data storage and data
processing.  %\alnote{prev. also acquisition was %part of the application}
\katznote{I don't see that it should be.  Doesn't the app %start after the data
is in the Tier 1 archive?}

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}

Both physics groups and individuals run jobs on the WLCG (Worldwide LHC
Computing Grid)~\cite{lhcb} as they see fit, either reading the reconstructed
data or reading intermediate datasets created by themselves or their
colleagues. If, in the future, resources should prove inadequate then
adjustments to working practices to coordinate processing may be made to
increase physics output.

The WLCG has a hierarchical architecture. CERN (the `Tier~0' site) is connected
to eleven `Tier~1' (national) sites by a dedicated optical network. This may be
extended in the future.  All data is stored at CERN, and as needed, various
parts of the dataset are copied (and cached, i.e., replicated) to the Tier~1
sites, and from there to `Tier 2' (regional) sites and to `Tier~3' (university
or group or individual) sites.

Before data is analyzed, the file to be analyzed is copied from the closest site
to the SE (Storage Element) at the WLCG site where the job will run. From there
it is either copied to a local disk or read from the SE.

There is no particular time constraint on the processing, although physics
groups are eager to understand the ATLAS experiments and thus want to process
data as quickly as possible.  In the event some data are missed during
processing, the only effect is a reduction of statistical precision in the final
analysis. Accurate bookkeeping is therefore important in order to know which
data have been processed.

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below}

Roughly 20~TB of data are collected every day. Multiple generations of processed
data are kept. Many data files are replicated on multiple sites; this is done
both by copying it to where it is likely to be useful and partly dynamically as
the need arises.  The bulk of the data is individual event data at different
stages of refinement. This is held in an in-house representation of C++
serialized objects. Though this format is not standard it is used by all LHC
experiments.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

%\katznote{from answer to question 3 below}

The WLCG has 250,000 cores distributed over 140 sites and with 100~PB of
disk. It uses a mixture of gLite software, experiment specific software and
physics group specific software.  The data is distributed and replicated as
described previously.

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

The data being processed can be considered as dynamic as it grows steadily as
new raw data are collected and the basic processing program (called
reconstruction) is run to produce information about specific particles involved
in a collision. This reconstruction is rerun periodically (two or three times
per year is expected) as calibrations of the detector are
improved. Reconstruction algorithms are also improved as are the physics groups'
selection codes.  %\alnote{Is the reconstruction code the fundamental entity
that is required to deal with new data? Sounds rather like an auxiliary tool.}
%\katznote{reconstruction is a second type of processing.}\alnote{should we
explicitly state that?}

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

Data analysis is really a pleasingly parallel task, where various events are
analyzed in parallel and then a summary is created of all the analyses.

The key challenge here is really in the infrastructure -- building a system that
can store massive amounts of data and move them to where the processing is to be
done by a large diverse group of users.

%\subsubsection{old material follows}
%
%
%\begin{itemize}
%\item Most of the dynamism is in infrastructure, not in the application
(e.g. use of PilotJob)
%\item Application dynamism: location of file (e.g. if file turns out to be not
local, look up in registry-- but not do this again)
%\item No specific QoS issue concerned with data transfer times
%\end{itemize}
%
%\subsubsection{New answers from Steve}
%
%3DPAS - Particle Physics - the example of Atlas physics channel selection.
%
% 1.  What is the purpose of the application?
%
%A physics group is interested in a specific physics channel (underlying
physical process) as recorded by the Atlas experiment at the LHC (Large Hadron
Collider) at CERN in Geneva. The set of available data grows steadily over time
as more data are collected and reconstructed.
%
% 2.  How is the application used to do this?
%
% Both physics groups and individuals run jobs as they see fit, either reading
the reconstructed data or reading intermediate datasets created by themselves or
their colleagues. If, in the future, resources should prove inadequate then
adjustments to working practices to coordinate processing may be made to
increase physics output.
%
% 3.  What infrastructure is used? (including compute, data, network,
instruments, etc.)
%
%The WLCG (Worldwide LHC (Large Hadron Collider) Computing Grid) is used with a
mixture of gLite software, experiment specific software and physics group
specific software. WLCG has 250,000 cores distributed over 140 sites and with
100~PB of disk. CERN is connected to the eleven `Tier 1' national sites by a
dedicated optical network. This may be extended in the future.
%
% 4.  What dynamic data is used in the application?
%
%The data being processed can be considered as dynamic as it grows steadily as
new raw data are collected and the basic reconstruction program is run to
produce information about specific particles involved in a collision. This
reconstruction is rerun periodically (two or three times per year is expected)
as calibrations of the detector are improved. Reconstruction algorithms are also
improved as are the physics groups' selection codes. One output of the
reconstruction is `tag' data - summary information written to a database to
facilitate event selection. This is not yet used very much. Calibration data are
held on an RDBMS with Frontier/Squid front ends.
%
% a.  What are the types of data,
%
% The bulk of the data is individual event data at different stages of
refinement. This is held in a in-house representation of C++ serialized
objects. Though this format is not standard it is used by all LHC experiments.
%
% b.  What is the size of the data set(s)?
%
% Roughly 20~TB of data are collected every day. Multiple generations of
processed data are kept. Many data files are replicated on multiple sites; this
is done both by copying it to where it is likely to be useful and partly
dynamically as the need arises.
%
% 5.  How does the application get the data?
%
%A copy of the file is made to the SE (Storage Element) at the site where the
job will run. From there it is either copied to a local disk or read from the
SE.
%
% 6.  What are the time (or quality) constraints on the application?
%
%If some data are missed during processing the only effect is a reduction of
statistical precision in the final analysis. Accurate book keeping is important
to know which data have been processed. If the selection code has been improved
then naturally the physics groups would like to see the results as soon as
possible.
%
%
%\subsubsection{Old answers from Steve}
%
% 1.  What is the purpose of the application?
%
%A physics group is interested in a specific physics channel (underlying
physical process) as recorded by the Atlas experiment at the LHC (Large Hadron
Collider) at CERN in Geneva. The set of available data grows steadily over time
as more data are collected and reconstructed. It would not be efficient to allow
individual physics groups to process this large volume of data in an
uncoordinated manner as it requires reading all available data.
%
% 2.  How is the application used to do this?
%
%Physics groups register their analysis code to be run on the next scheduled
analysis ``train''. The train leaves every two weeks and reads through all the
data, runs the physics selection code of each group and writes out the selected
events and output from the computation to produce relatively small data sets for
each physics group. The groups then use these selections to perform more ad-hoc
analysis.
%
% 3.  What infrastructure is used? (including compute, data, network,
instruments, etc.)
%
%The WLCG (Worldwide LHC (Large Hadron Collider) Computing Grid) is used with a
mixture of gLite software, experiment specific software and physics group
specific software. WLCG has 250,000 cores distributed over 140 sites and with
100~PB of disk. There is no wide area dedicated network infrastructure.
%
% 4.  What dynamic data is used in the application?
%
%The data being processed can be considered as dynamic as it grows steadily as
new raw data are collected and the basic reconstruction program is run to
produce information about specific particles involved in a collision. This
reconstruction is rerun periodically (two or three times per year is expected)
as calibrations of the detector are improved. Reconstruction algorithms are also
improved as are the physics groups' selection code. Calibration data are held on
an RDBMS with Frontier/Squid front ends.
%
% a What are the types of data,
%
% The bulk of the data is individual event data at different stages of
refinement. The is held in a non-standard representation of C++ serialized
objects.
%
% b.  What is the size of the data set(s)?
%
% Currently the reconstructed data (excluding copies) is about m TB and is
growing at about n TB per month. Many data files are replicated on multiple
sites. This is done both by copying it to where it is likely to be useful and
partly dynamically as the need arises.
%
% 5.  How does the application get the data?
%
%This is mostly by copying the file to where the job will run - though some
remote file access is also used.
%
% 6.  What are the time (or quality) constraints on the application?
%
%If some data are missed during processing the only effect is a reduction of
statistical precision in the final analysis. If the selection code has been
improved then naturally the physics groups would like to see the results as soon
as possible.

%\subsection{Astrophysics -- Virtual Observatory (Bob) \label{astroBob}}
%
%\katznote{this will be merged into the LSST and SOA entries}
%
%
%\begin{itemize}
%\item Analysis of data to generate event streams $\rightarrow$ lead to the %
positioning of telescopes
%\end{itemize}
%
%\subsubsection{Answers from Bob}
%
%I'll assume that the ``application'' in question is the sum of the two
%activities, despite the fact that they are different in many respects, %and
I'll try and distinguish between them where appropriate.
%
%Let me know if you want clarification or further information - in some cases I
wasn't quite sure what you were after.
%
%1. What is the purpose of the application?
%
%An increasingly important feature of astronomical research is the existence of
%systematic sky surveys covering the whole sky (or large fractions thereof) in
%many different wavebands. The principal outcome from these surveys are
%astronomical source catalogues, which contain values for attributes
%characterizing all the celestial sources detected in the sky survey dataset.
%These are now typically implemented in relational databases, with SQL
interfaces %implemented through web pages, and the principal goal of the Virtual
Observatory %(VO) is to standardize access to these, and other astronomical data
resources, %so that astronomers can readily perform multi-wavelength analyses,
combining all %the extant data on particular sources.
%
% 2. How is the application used to do this?
%
%The VO includes a registry that contains metadata describing all the data
%resources published using VO data access standards, enabling the discovery of
%datasets relevant to a particular analysis. Web service implementations of data
%access protocols enable users to access data in these repositories
%programmatically, and distributed scratch space is provided for the storage of
%intermediate result sets, so the user does not need to route large data flows
%through his/her own workstation. In the future, more and more data analysis
%software will be made available through web services compliant with VO
%standards, enabling more of the data integration and analysis process to be
%combined in workflow.
%
% 3. What infrastructure is used? (including compute, data, network, %
instruments, etc.)
%
%The data resources published to the VO exist in a number of data centers
%distributed internationally. Major datasets---e.g. those from large sky
%surveys---are typically implemented in relational databases running on
high-spec servers %or clusters thereof. The compute requirements of the VO are
currently modest, %but are likely to increase over time. Astronomers typically
use normal academic %networks, but more specialized networking is required in
some cases: e.g., sky %survey data is moved from Cambridge to Edinburgh using
the UKLight project's %dedicated optical network, and radio interferometers
require high-bandwidth %fibre connections from each antenna to the correlator.
%
% 4. What dynamic data is used in the application?  %(a) What are the types of
data, (b) What is the size of the data set(s)?
%
%The VO currently includes little dynamic data. The one exception is VOEvent
%messages, which are notifications of transient events which can be used to
%trigger follow-up observations on other telescopes. These are small XML
messages %containing basic information about the position of a transient source
and a %small amount of other metadata which the operators of a given telescope
can use %to define triggers for follow-up observations of interest to them or
their %community.
%
% 5. How does the application get the data?
%
%Assuming this refers solely to the dynamic data, then a number of telescopes
%undertaking real-time detection of transient events emit VOEvent messages
%through feeds to which other parties can subscribe.
%
%6. What are the time (or quality) constraints on the application?
%
%The time constraints differ for different types of transient source, but in
some %cases notification must be made within a few minutes if it is to be
useful; the %overheads associated with setting up a follow-up observation mean
that much %shorter time scales are practically irrelevant, although they are
still %scientifically interesting. The quality constraints centre on providing a
%telescope with good enough information about a potential transient source that
%it can make a reliable decision about whether to interrupt its existing
%observing program to follow-up the event. In most cases, the candidate
%transient will be some sort of noise event, so fairly sophisticated filtering
is %required to bring the false positive rate down to something acceptable,
given %the cost of observing time on cutting edge telescopes.

\subsection{LSST \label{astroAdam}}

%\katznote{from Adam Barker and Bob Mann}
%
%\katznote{summary below assigned to Dan}
%\subsubsection{new summary}

% 1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

This application, which similarly to the previous application is an {\em
infrastructural application}, wants to find and study variable objects and
moving objects, using a telescope that takes images each night. This is a
scenario that may be used by Large Synoptic Survey Telescope (LSST)~\cite{lsst},
and is somewhat similar to that done by the Palomar Transient Factory
(PTF)~\cite{ptf,ptf2,ptf3} and by Pan-STARRS~\cite{ps1,ps2}.  Overall, this
application is a leading-edge example of the work needed to build a set of data
(catalog and images) that then can be used by others for analysis (an example of
which is the next application, in \S\ref{astroSOAAdam}.)  This application
contains three stages: data transformation \& decisions, data storage.
%\alnote{Is the telescope not %part of the application?} \katznote{I don't know
- what do you think?  This %is a general issue I'm struggling with in this and
other applications.} \alnote{I struggle as well - it is just confusing that the
term telescope is part of the name} (Data generation comes from the telescope
itself, and processing is done outside of this application, e.g. using the next
application.)

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app} %\ysnote{This section feels like it has too much
details/steps. Maybe we just need to stress the dynamic and distributed parts
e.g. \#1-6 can be collapsed into a couple of bullets, and \#7 into another.}

Processing of data (data analysis and detection of new sources) from each image
should be done while the next image is being taken by the telescope.  This means
that if anything unusual is detected, normal observation can be
interrupted. Other observing resources can also be notified instantly, so they
can observe the same event. As data is collected, it is added to all the data
previously detected from the same location of sky to create a very deep master
image. LSST will also build up a database of all known moving objects.

Every time a new image of the sky is obtained, the master image will be
subtracted from it. The result is a subtracted image that contains the
difference between the current sky image and its average state.

This subtracted image is then processed by a cluster of computers with 3 main
steps:
\begin{enumerate}[(i)]

\item Using the existing object catalog (containing the known orbits of all
known objects), find objects which are expected to appear in subtracted image,
given the area of sky and time of day.  Cross-match expected objects with
sources in subtracted image, resulting in the unmatched-source catalog, which
contains sources that cannot be matched with a previously known object.  If
these sources can be tracked over a few images, an orbit can be determined for
them and they can be added to the object catalog.  Also, the orbit catalog can
be updated with redetections of known objects; each rediscovery provides
information that improves the known orbits.

\item Attempt to classify all entries in the orbit catalog (i.e., all the
objects which now have known orbits). If any Near Earth Objects are detected to
be passing close to the earth, alerts are generated to astronomers so that
follow up observations can be scheduled.

\item If an unknown object is found that cannot be classified locally, at least
two possible options exist. One is a coordinated effort of comparison with other
observatories, where the local observatory issues a call for participation to a
set of additional possible observatories, each of which can accept or reject the
call.  The local observatory then chooses one of the accepting observatories
from which to obtain data, and that observatory takes data and returns it. The
local observatory then decides what to do: classify the object, contact a human,
or call for more observations (automatically to other observatories that decide
whether or not to accept, and when)

A second option is less coordinated, where this observatory would send VOEvent
messages, notifications of transient events which can be used to trigger
follow-up observations on other telescopes. These are small XML messages
containing basic information about the position of a transient source and a
small amount of other metadata which the operators of a given telescope can use
to define triggers for follow-up observations of interest to them or their
community.

\end{enumerate}

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below} %\ysnote{The FITS images themselves
may be less interesting since they may only be used ocassionally % (e.g. for
re-processing images using a new algorithm) after the initial object extraction
into % databases. The object databases sizes may be more relevant, and they can
be 3-orders of magnitude smaller % than the images.}

LSST will generate 36~GB of data every 30 seconds, and over a 10-hour winter
night, will collect up to 30~TB.  Data will be stored as FITS image, and in
large databases containing analyses of the discoveries.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

%\katznote{from answer to question 3 below}

The overall system will contain a central telescope with computing data and
resources, linked to a wide area network of observatories and data and computing
resources.  %\alnote{are these data and computing resources still part %of the
application LSST?} \katznote{I'm not sure - how do you think we should handle
this?} \alnote{not sure, but since we say that processing and storage is part of
the app, the resource for this need to be at some place}

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

The time constraints differ for different types of transient sources, but in
some cases notification of remote telescopes must be made within a few minutes
if it is to be useful; the overheads associated with setting up a follow-up
observation mean that much shorter time scales are practically irrelevant,
although they are still scientifically interesting. The quality constraints
center on providing a remote telescope with good enough information about a
potential transient source that it can make a reliable decision about whether to
interrupt its existing observing program to follow-up the event. In most cases,
the candidate transient will be some sort of noise event, so fairly
sophisticated filtering is required to bring the false positive rate down to
something acceptable, given the cost of observing time on cutting edge
telescopes.

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

%\alnote{Is the workflow part still part of the app as defined before with the
stages: data transformation \& decision and storage?} \katznote{I think this is
data transformation.}\alnote{ok}

Data will initially be analyzed at the telescope; discoveries will be sent to
other centers.

Pan-STARRS, a contemporary sky survey with similar goals as LSST, uses a
scripted programming model on a cluster for image processing of telescope frames
to extract object attributes as a CSV file.  This is followed by followed by
scientific workflows, based on Trident, that load the files into individual
databases daily, and merges them with a master distributed-database on a weekly
basis on an HPC cluster~\cite{Simmhan:Building:2009}. Data extraction and mining
operations by astronomers are composed using a pseudo-SQL language with user
defined functions and submitted for execution on the master database using a
batch system.

%\subsubsection{old material follows}
%
%\begin{enumerate}
%\item What is the science/engineering problem? What is the purpose of the
application?
%
%Find and study variable objects and moving objects.
%
%2 How will the application solve the problem?
%
%Processing of data (data analysis and detection of new sources) from each image
should be done while the next image is being taken by the telescope.  This means
that if anything unusual is detected, normal observation can be
interrupted. Other observing resources can also be notified instantly, so they
can observe the same event. As data are collected, they will be added to all the
data previously detected from the same location of sky to create a very deep
master image. LSST will also build up a database of all known moving objects.
%
%Every time a new image of the sky is obtained, the master image will be
subtracted from it. The result is a subtracted image that contains the
difference between the current sky image and its average state.
%
%This subtracted image is then processed by a cluster of computers with the
following steps:
%\begin{enumerate}
%
%\item Query object catalog (knows orbits of all known objects) to find objects
which are expected to appear in subtracted image, given the area of sky, time of
day.
%
%\item Cross-match expected objects with sources in subtracted image. The result
is the unmatched-source catalog, which contains sources that can't be matched
with a previously known object
%
%\item Use unmatched-source catalogue to compute pairs of detections separated
by short time intervals, called tracklets. A tracklet is an observable, short
section of orbit. In order to create pairs of detections, all objects in the
unmatched-source catalog are queried against the orbit catalog, in an attempt to
obtain data about these objects from earlier observations. If earlier
observations and current observations can be linked to form tracklets, they are
stored in the tracklet catalog.
%
%\item Use tracklet catalog to attempt to link these tracklets over a larger
time window. This is achieved by again querying the orbit catalog, this time
looking for observations of the same objects even further back in time. If
matches can be found and the tracklets can be extrapolated out into longer
sections of orbit, then they are added as new orbits to the orbit catalog.
%
%\item Update orbit catalog with redetections of known objects. Each rediscovery
provides information that improves the known orbits.
%
%\item Attempt to classify all entries in the orbit catalog (i.e., all the
objects which now have known orbits). If any Near Earth Objects are detected to
be passing close to the earth, alerts are generated to astronomers so that
follow up observations can be scheduled.
%
%\item Compare with other observatories: If an unknown object is found that
cannot be classified locally
%\begin{enumerate}
%\item issue call for participation to set of possible observatories
%\item each observatory accepts or rejects the call
%\item local observatory chooses from accepting observatories one to obtain data
from
%\item that observatory takes data and sends it to local observatory
%\item local observatory decides what to do - classify the object, contact a
human, call for more observations (automatically to other observatories, that
decide whether or not to accept, and when)
%\end{enumerate}
%
%\end{enumerate}
%
%
%2a Which of the following describes the application
%
%large (exascale) data and large (exascale) computing (e.g., data mining,
analytics, bioinformatics)
%
%2b Can you quantify the amount of computing vs the amount of data?
%
%LSST will generate 36~GB of data every 30 seconds, and over a 10-hour winter
night, will collect up to 30~TB.
%
%3 What infrastructure will be used? (including compute, data, network,
instruments, etc.)
%
%A linked network of observatories and data and computing resources.
%
%(a) What will be distributed?  data, compute, both?
%
%both
%
%4 What data be used in the application?
%
%Fits images, and large databases for analysis of the discoveries.
%
% (a) What are the types of data
%
% (b) What is the size of the data set(s)?
%
% (c) Will any data be dynamic?
%
%5 How will the application get the data?
%
%Some will be analyzed at the mountain top, discoveries will be sent to other
centers.
%
%6 Are there time (or quality) constraints on the application?  (If so, what are
they?)
%
%Time component is critical. As real-time as possible.
%
%7 Please feel free to also talk about the current state of the application, if
it exists today, and any specific gaps that you know will need to be overcome.
%
%
%also, the Palomar Transient Factory is the best example to date:
%
%See:
%
%\url{http://www.astro.caltech.edu/ptf/}
%
%\url{http://xxx.lanl.gov/abs/0906.5350}
%
%\url{http://xxx.lanl.gov/abs/0906.5355}
%
%\url{http://arxiv.org/abs/1011.2199}
%

\subsection{SOA Astronomy \label{astroSOAAdam}}

%\katznote{from Adam Barker}
%
%\katznote{summary below assigned to Dan}
%\subsubsection{new summary}

% 1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

An increasingly important feature of astronomical research is the existence of
systematic sky surveys covering the whole sky (or large fractions thereof) in
many different wavebands. The principal outcome from these surveys are
astronomical source catalogues, which contain values for attributes
characterizing all the celestial sources detected in the sky survey dataset.
These are now typically implemented in relational databases, with SQL interfaces
implemented through web pages, and the principal goal of the Virtual Observatory
(VO) being developed by the International Virtual Observatory Alliance
(IVOA)~\cite{ivoa} is to standardize access to these, and other astronomical
data resources, so that astronomers can readily perform multi-wavelength
analyses, combining all the extant data on particular sources.  The previous
application (\S\ref{astroAdam}) was an example of a ``generator'' of such data
resources, this application is a typical ``user'' of them.  As such, it can
generally be considered as having just the last application stage: data
processing.  Thus it is an {\em infrastructural application}.  The particular
example described here calculates the photometric redshift for a given area of
sky using two tools: HyperZ and ANNz, both accessible via web services
interfaces~\cite{barker}.  However, it represents many service-oriented
architecture applications that are designed to work with the VO.

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}

The VO includes a registry that contains metadata describing all the data
resources published using VO data access standards, enabling the discovery of
datasets relevant to a particular analysis. Web service implementations of data
access protocols enable users to access data in these repositories
programmatically, and distributed scratch space is provided for the storage of
intermediate result sets, so the user does not need to route large data flows
through his/her own workstation. In the future, more and more data analysis
software will be made available through web services compliant with VO
standards, enabling more of the data integration and analysis process to be
combined in workflow.

The application orchestrates a set of services through a pipeline in order to
compute the redshift of a given area of sky.  It contains a set of components.
1) The Wide Field Survey Archive (WFS), which is an image catalog. 2) A
spectroscopic database, used to provide spectroscopic data for an area of
sky. 3) The SExtractor tool, which is an image processing that extracts all
objects of interest (stars, galaxies, etc.)  4) A cross matching tool that
compiles one table of all objects of interest over the five wavebands.  5)
Hyperz, the first photometric redshift estimation algorithm.  6) ANNz, the
second photometric redshift estimation algorithm, and 7) mySpace, the AstroGrid
storage service.

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below}

Images will be O(1) GB files.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

The data resources published to the VO exist in a number of data centers
distributed internationally. Major datasets---e.g. those from large sky
surveys---are typically implemented in relational databases running on high-spec
servers or clusters thereof.  The services in this application are also
distributed, with web service interfaces used to link them.

%\katznote{from answer to question 3 below}

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

The WFS catalog will be updated regularly, therefore a query at time $x$ will
not provide the same data as a query at time $y$.  In addition, the specific
queries that will be issued by users are unknown, and will change over time.

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

Enough data needs to be available for a given area of sky from both the WFS
archive and the spectroscopic archive.

%\subsubsection{old material follows}
%
%
%1. What is the purpose of the application?
%
%Calculates the photometric redshift for a given area of sky [RA, DEC] using two
tools: HyperZ and ANNz, both accessible via WS interfaces.
%
%2. How is the application used to do this?
%
%The application is a service-oriented workflow. Services are orchestrated
through a pipeline in order to compute the redshift of a given area of sky.
%
%3. What infrastructure is used? (including compute, data, network, instruments,
etc.)
%
%The Wide Field Survey Archive (WFS): image catalogue %Spectroscopic database:
used to provide spectroscopic data for an area of sky
%
%SExtractor tool: image processing, extracts all objects of interest (stars,
galaxies etc.)  %Cross matching tool: compiles one table of all objects of
interest over the 5 wavebands
%
%Hyperz: photometric redshift estimation algorithm 1 %ANNz: photometric redshift
estimation algorithm 2
%
%mySpace: AstroGrid storage service
%
%\katznote{checking more on this point}
%
%4. What dynamic data is used in the application?
%
%The WFS catalogue will be updated regularly, therefore a query at time x will
not provide the same data as a query at time y.
%
%(a) What are the types of data, (b) What is the size of the data set(s)?
%
%Data sets (images) will be in the GB range, I can attempt to clarify these
details.
%
%5. How does the application get the data?
%
%WS interfaces.
%
%6. What are the time (or quality) constraints on the application?
%
%Enough data needs to be available for a given area of sky from both the WFS
archive and the spectroscopic archive.

\subsection{Understand the Cosmic Microwave Background \label{astroJulian}}

%\katznote{from Julian Borrill}
%
%\katznote{summary below assigned to Dan}
%\subsubsection{new summary}

% 1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

This {\em traditional application} performs data simulation and analysis to
understand the Cosmic Microwave Background (CMB)~\cite{cmb}, which is an image
of the Universe as it was 400,000 years after the Big Bang. Tiny fluctuations in
the CMB temperature and polarization encode the fundamental parameters of
cosmology and, using the Big Bang as the ultimate particle accelerator,
ultra-high energy physics. Extracting this information from the data gathered by
current and anticipated CMB observations is an extremely computationally
intensive endeavor for which massively parallel tools have been developed.  This
application contains just one stage: data processing.

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}

The CMB community wants to extract cosmology and fundamental physics from the
observations gathered by the detectors as time-ordered sequences of O($10^{12}$
- $10^{15}$). These observations are reduced first to a map of O($10^6$ -
$10^8$) sky pixels, then to O($10^3$ - $10^4$) angular power spectrum
coefficients, and finally to O(10) cosmological parameters~\cite{cmb2}.

The central computing challenge for any CMB dataset is the simulation and
analysis of O($10^4$) synthetic observations, used to correct for biases and
quantify uncertainties in the analysis of the real data.  Preconditioned
conjugate gradient techniques are used to solve for the maximum likelihood sky
map given the input data (obtained by scanning the sky with hundreds to
thousands of detectors for weeks to years) and its piecewise stationary noise
statistics.  To avoid the I/O bottleneck inherent in the traditional
simulate/write/read/map paradigm, all simulations are performed on the fly only
when requested by the map-making code.

Going from the map to the angular power spectrum is the computationally most
expensive step. The exact solution scales with the cube of the number of pixels
(in the map), so going from map to angular power spectrum is ruled out now (but
was possible earlier for much smaller observations). The approximate solution
requires sets of O($10^4$) Monte Carlo realizations of the observed sky to
remove biases and quantify uncertainties, each of which involves simulating and
mapping the time-ordered data.

The map-making application is therefore applied to both real and simulated data,
but many more times to simulated data (which requires us to use the on-the-fly
simulation module too).

%The computing scales with the number of observation samples, while the
communication and I/O scale with the number of pixels in the map. For the
current Planck satellite experiment these are O($10^{12}$) and O($10^8$)
respectively; for the proposed CMBpol satellite they will be O($10^{15}$) and
O($10^8$); that is, a 1000x increase in calculation cost is expected over the
next 15 years (including intermediate suborbital pathfinder experiments) but the
communication and I/O costs will be fixed.

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below}

There is O(1 - 10) TB input data and a similar volume of output data.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

%\katznote{from answer to question 3 below}

This application is targeting the largest supercomputers available: Hopper,
Tianhe, Blue Waters, etc. It is all about the cycles, although as the
concurrency increases previously solved I/O and communication scaling
bottlenecks typically re-emerge.

The application currently uses a single HPC systems, but the scientists have
discussed using distributed systems, with a model of remote systems being used
for the data simulations.  Each simulation would be launched from the central
system that is building the map, and output data from the simulations would be
asynchronously delivered back to that central system as files that would be
incorporated in the map as they are produced.

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

Overall, this application can make use of whatever computing it can access in
order to run the simulations needed to flesh out the observed data (which can be
read from local disk).  The simulations are dynamic in time and in location.

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

A goal is to simulate and map O($10^2$) realizations of an experiment's data in
O(1) wallclock hour to provide 10\% errors during the early analysis stages, and
O($10^4$) realizations in O(100) wallclock hours for the final definitive
analysis.

%\subsubsection{old material follows}
%
%1. What is the science/engineering problem? What is the purpose of the
application?
%
%Cosmic Microwave Background (CMB) data simulation and analysis. The CMB is an
image of the Universe as it was only 400,000 years after the Big Bang. Tiny
fluctuations in the CMB temperature and polarization encode the fundamental
parameters of cosmology and, using the Big Bang as the ultimate particle
accelerator, ultra-high energy physics. Extracting this information from the
data gathered by current and anticipated CMB observations is an extremely
computationally intensive endeavor for which we have developed our massively
parallel tools.
%
%2. How will the application solve the problem?
%
%The central computing challenge for any CMB dataset is the simulation and
analysis of O($10^4$) synthetic observations, used to correct for biases and
quantify uncertainties in the analysis of the real data. We use preconditioned
conjugate gradient techniques to solve for the maximum likelihood sky map given
the input data (obtained by scanning the sky with hundreds to thousands of
detectors for weeks to years) and its piecewise stationary noise statistics. To
avoid the IO bottleneck inherent in the traditional simulate/write/read/map
paradigm, all simulations are performed on the fly only when requested by the
map-making code.
%
%(a) Which of the following describes the application
%
% i) large (exascale) data and large (exascale) computing (e.g., data mining,
analytics, bioinformatics)
%
% ii) large (exascale) data and modest computing (e.g., search)
%
% iii) modest data and large (exascale) computing (e.g., simulation/modeling)
%
%(iii) O(1 - 10) TB input data and a similar volume output.
%
%(b) Can you quantify the amount of computing vs the amount of data?
%
%The computing scales with the number of observation samples, while the
communication and IO scale with the number of pixels in the map. For the current
Planck satellite experiment these are O($10^12$) and O($10^8$) respectively; for
the proposed CMBpol satellite they will be O($10^15$) and O($10^8$) - that is,
we expect a 1000x increase in calculation cost over the next 15 years (including
intermediate suborbital pathfinder experiments) but fixed communication/IO
costs.
%
%3. What infrastructure will be used? (including compute, data, network,
instruments, etc.)
%
%We are targeting the largest supercomputers available - Hopper, Tianhe, Blue
Waters, etc. It's all about the cycles, although as the concurrency we work at
increases previously solved IO and communication scaling bottlenecks typically
re-emerge.
%
%(a) What will be distributed?  data, compute, both?
%
%Both.
%
%4. What data will be used in the application?
%
%Real and simulated CMB data.
%
%(a) What are the types of data
%
%(b) What is the size of the data set(s)?
%
%(c) Will any data be dynamic?
%
%See above.
%
%5. How will the application get the data?
%
%Real data are read from disk; simulated data are generated on the fly.
%
%6. Are there time (or quality) constraints on the application?  (If so, what
are they?)
%
%Typically we want to be able to simulate and map O($10^2$) realizations of an
experiment's data in O(1) wallclock hour to provide 10\% errors during the early
analysis stages, and O($10^4$) realizations in O(100) wallclock hours for the
final definitive analysis.
%
%7. Please feel free to also talk about the current state of the application, if
it exists today, and any specific gaps that you know will need to be overcome.
%
%Moving to the current and future generations of heterogeneous many-core systems
will require a significant re-working of our analysis codes, both to deal with
communication over hundreds of thousands of cores and to take advantage of
accelerator opportunities such as GPUs.

\subsection{Sensor Network Application \label{sensorSimon}}

%\katznote{from Simon Dobson}
%
%\katznote{summary below assigned to Yogesh}
%\subsubsection{new summary}

% 1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

Marine sensing covers a number of applications, including environmental
monitoring, remote exploration, marine life surveys, and habitat assessment. In
each case, the main challenge is collecting data reliably from a difficult
working environment, without interfering with the natural behaviors the animals
exhibit.

A canonical example is the monitoring of seal and other sea mammal populations
by the Scottish Oceans Institute (SOI), which tags animals with sensor packages
that can record dive behavior, speed, and movement~\cite{SMSSeal,SealContact}.
These datasets are then analyzed offline using traditional statistical
techniques and integrated with Google Earth to visualize animal tracks on a
large scale.  This is an {\em infrastructural application}. It begins after data
generation, and it includes the data storage and processing stages.
%\alnote{data transformation?} \katznote{I don't think so - I think all data %is
stored and then processed as users want} \alnote{I think most sensors do some
form of filtering/aggregation before transmitting things to a
backend...}\katznote{but I would argue that the application %starts after the
filtering/aggregation}

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}

The sensor packages on the marine mammals report back to base when the animal
crawls up a beach and comes within range of a cellular network. Data are
collected remotely and brought to a central site for analysis. Analysis includes
statistical applications and visualization is done using Google Earth.

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below}

The data sets collected are modest in size by scientific-data standards. They
include positions and motion vectors for sea mammals, and concentrations and
gradients for environmental missions.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

%\katznote{from answer to question 3 below}

Time series data is published from distributed sensors, but they are stored in a
central repository. The data is analyzed locally and visualized by distributed
users.

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

Time series data is published dynamically from the sensors when they are within
communication distance of cellular towers. So the time series data arrives at
the central repository asynchronously relative to the time of collection.

There is some basic adaptation to the resolution of the data collected to
account for different resolutions of the sensor packs.  Larger-scale
environmental sensing applications may make use of more structured and extensive
adaptation, such as changing the sampling frequency and other management
characteristics in response to the data being observed.

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

Collecting data reliably is challenging due to the distributed deployment of
sensors in hostile environments.

% \subsubsection{old material follows}

% \begin{itemize}
% \item Environmental (Marine) -- sensing within a hostile environment, % data
quality + availability of communication infrastructure (Simon)
% \item Patient monitoring -- data quality, type of data to transmit % (Omer)
% \end{itemize}

% \subsubsection{Answers from Simon}

% 1. What is the purpose of the application?

% Marine sensing covers a range of different applications, including %
environmental monitoring, remote exploration, marine life surveys and % habitat
assessment. In each case, the main challenges are collecting % data reliably in
what is an extremely challenging technical % environment---far more so than in
air---and (in the case of biological % surveillance), without interfering with
the natural behaviors the % animals exhibit.

% 2. How is the application used to do this?

% At present most applications are quite small-scale. A canonical example % is
the monitoring of seal and other sea mammal populations by the % Scottish Oceans
Institute (SOI), which works by tagging animals with sensor % packages that can
analyze dive behavior, speed, movement and the like, % and report back to base
when the animal comes within range of a cellular % network.

% 3. What infrastructure is used? (including compute, data, network,
instruments, etc.)

% The data sets collected are not enormous by scientific-data standards, % and
are analyzed offline using traditional statistical techniques.  % Recent work by
SOI has included integration with Google Earth to allow % animal tracks to be
visualized on a large scale.

% Data are collected remotely, then brought to a central site for analysis.

% 4. What dynamic data is used in the application?

% In the sea mammal case the data isn't all that dynamic, being mainly a % time
series collected with only basic adaptation to the resolution etc.  %
Larger-scale applications in environmental sensing in particular would % make
use of more structured and extensive adaptation (e.g., % changing the sampling
frequency and other management characteristics in % response to the data being
observed).

% (a) What are the types of data, (b) What is the size of the data set(s)?

% For sea mammals: positions, motion vectors. For environmental missions: %
concentrations and gradients. I'd have to enquire about dataset sizes, % but not
excessive.

% 5. How does the application get the data?

% For the sea mammals the animal crawls up a beach, the sensor packages % sees
the cellular network and transmits. For environmental missions % we're still
wondering about that ourselves :-)

% 6. What are the time (or quality) constraints on the application?

% Not sure I see what you mean here. Generally it's just a matter of % capturing
at whatever resolution device the sensor pack can support.

% I could probably get more complete -- perhaps *way* too complete :-) -- %
answers from Mike Fedak in Scottish Oceans if that'd help.

%\subsection{Bioinformatics Applications (Tom Slezak) \label{bioTom}}
%
%\katznote{summary below assigned to Shantenu - perhaps merge this with Sylvia's
app's section}
%\subsubsection{new summary}
%
%1. {\bf What does the application do? What problem does it solve?}
%
%\katznote{from answer to question 1}
%
%2. {\bf How does it do it?}
%
%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}
%
%3. {\bf What are the big data aspects?}
%
%\katznote{from answer to question 4b below}
%
%4. {\bf What are the distributed aspects?}
%
%\katznote{from answer to question 3 below}
%
%5. {\bf What are the dynamic aspects?}
%
%\katznote{from answer to question 4a below}
%
%6. {\bf What else is important?}
%
%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}
%
%
%\subsubsection{old material follows}
%
%1. What is the science/engineering problem? What is the purpose of the
application?
%
%Comparing metagenomic (DNA sequence) data sets against each %other. This could
be to distinguish human disease biomarkers among %different individuals,
determine different soil (or water or air or %...) microbial composition from
different samples, etc.
%
%2. How will the application solve the problem?
%
%In general this is a large (for biology -- don't think in terms of exascale!)
data problem with large (all-vs-all comparison) computing. There are strategies
to make this more scalable that will reduce the computing volume.
%
%(a) Which of the following describes the application
%
% i) large (exascale) data and large (exascale) computing (e.g., data mining,
analytics, bioinformatics)
%
% ii) large (exascale) data and modest computing (e.g., search)
%
% iii) modest data and large (exascale) computing (e.g., simulation/modeling)
%
%(b) Can you quantify the amount of computing vs the amount of data?
%
%Potentially peta-bytes or exa-bytes of data. Not sure how to quantify the
computing. (Huge memories will be potentially more valuable than huge numbers of
cores -- although this is heresy to the physics crowd.)
%
%3. What infrastructure will be used? (including compute, data, network,
instruments, etc.)
%
%(a) What will be distributed?  data, compute, both?
%
%Individual data sets will be terabyte-scale. There could be many thousands of
such sets to compare. Having them all local may not be feasible (the transfer
time from where they are generated by others may be prohibitive. Computing might
then have to be distributed. Biology is wrestling with this now.
%
%4. What data be used in the application?
%
%(a) What are the types of data
%
%Genomic sequence, billions of short (35-100?) bases of DNA; unassembled
%
%(b) What is the size of the data set(s)?
%
%Terabyte scale
%
%(c) Will any data be dynamic?
%
%Not for this problem
%
%5. How will the application get the data?
%
%That is a major part of the problem! Data will (in general) be generated from
other places around the world. Getting to the data will be an issue.
%
%6. Are there time (or quality) constraints on the application?  (If so, what
are they?)
%
%You'd like answers in under a day, generally.
%
%7. Please feel free to also talk about the current state of the application, if
it exists today, and any specific gaps that you know will need to be overcome.
%
%Biology is wrestling now with about 3 orders of magnitude increase in DNA
sequence data generation capability in the past 2 years. All algorithms,
architectures, storage, and networks have broken or will break soon. Some
biology problems (not necessarily the one I describe here) are not good fits for
core-rich and memory-poor modern architectures. For some problems (comparing
k-mers across metagenomic data sets, for instance) you need terabyte-scale
memory, not thousands of cores.

\subsection{Climate \label{climateDan}}

%\katznote{from Don Middleton, mostly}
%
%\katznote{summary below assigned to Dan}
%\subsubsection{new summary}

% 1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

This {\em infrastructural application} is aimed at supporting international
CMIP/IPCC~\cite{intergovernmental2007fourth,cmip} intercomparison activity,
which involves producing, storing, and analyzing data from multiple climate
simulations.  It includes data generation, storage, and processing stages.

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}

The overall system is comprised of three stages.  In the first stage, climate
centers run a prescribed set of common experiments that produce 2-10 PB of data.
Data can also come from sensors, in which case the center would still
post-process the data before it would be published.  Centers can then publish
their own output data, or send it to another center to publish.  The data is
generated over a roughly 2-year period (then post-processed and published over a
few more months).  This stage is characterized by being both distributed-compute
and distributed-data intensive.

The second stage is data storage.  The Earth System Grid Federation
(ESGF)~\cite{esgf} develops and deploys a federated network of gateways and
associated data nodes.  As models run at each center, the output data is
post-processed into packages with common formats and prescribed metadata
conventions.  Most centers deploy the ESGF data node software stack, and using
this, they manage the data from their experiments.  The data node software stack
scans the data, makes sure it has the right metadata fields, does QA/QC, and
builds a set of catalogs of the prepared data.  Minimally, the catalog provides
HTTP links to data elements, and it can also provide GridFTP endpoints, and/or
product services that abstract the dataset in other ways: get a whole file,
subset, browse, etc.  When the center is happy with the data/catalog in the data
node, they publish it to a host/affiliated gateway.  This submits the catalog to
the gateway.  The gateway then shares this catalog with other gateways so that
all gateways have a consistent view of all the published data.  In general, most
data is replicated at several sites.  The replication activity is manually
requested/initiated by a gateway owner.  The properties of this stage are that
it is distributed data-intensive (with distributed gateways and data nodes) and
that it is dynamic (data appears in the system over time).

The third stage is data analysis.  The approximately 20,000 users (as of
Dec. 2010) can browse/search a catalog at any gateway and locate data, which
might be hosted by a data node affiliated with another gateway. (This can output
a `wget' script that can later fetch the data.)  They can also authenticate, and
thus gain access to a group, which might have private data.  They can also
download data via http, GridFTP, or access product services (the latter of which
uses ESGF's data retrieval syntax---DRS---and which can be scripted.)  Some
users will analyze data from a single model.  However, many applications are
multi-model analyses, where many users want to look at the same parts of the
output of some/all of the models, to understand if and how the models differ.
Some centers will gather some/all of the core archive (plus more, perhaps) on
local systems for local users to perform ``power analyses.''  Each user's data
analyses are almost always done on a single system.  Because there is no
distributed computing infrastructure, this stage is search, access, and transfer
intensive.

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below}

The overall data generated and stored in the ESGF is 2-10 PB.  This includes a
core archive, which is 1-2 PB in size and contains the most popular data.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

%\katznote{from answer to question 3 below}

The data is generated by a distributed set of climate centers.  It is stored in
a distributed set of federated archives.  And it is used by a distributed set of
users, who either run data analyses on a climate center with which they are
associated, or they gather data from the ESGF to a local system for their
analyses.

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

Data is generated over time, so the data in the ESGF changes.  One might imagine
a future version of ESGF where the launching of data analysis jobs is automated,
in which case the location of those jobs would be dynamic, and the system might
also be able to respond (or optimize for) various types of applications.

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

This application/infrastructure serves a large group of users, and involves a
large amount of political and technical consensus to work.

%\subsubsection{old material follows}
%
%\subsubsection{CDAT} (from Dan)
%
%The Climate Data Analysis Tools (CDAT)~\cite{CDAT} are a scriptable %(Python)
set of packages and tools that are currently heavily used for %analysis of
climate data.  CDAT has been developed by LLNL's Program %for Climate Model
Diagnosis and Intercomparison (PCMDI), and is often %used with data from the
Earth System Grid (ESG)~\cite{ESG}.  The most %common current usage of these
tools with the data is that a user %downloads a set of data from ESG to a local
resource, then uses the %CDAT to analyze the data.  However, the PCMDI group is
now planning to %support a new usage mode, where data transfer can be scripted.
%Initial development of this capability is planned for February 2011, %using the
Globus online data transfer service~\cite{globusOnline} as a %scriptable
component.
%
%\subsubsection{ESGF} (from Don and Dan)
%
%Aimed at supporting international CMIP/IPCC intercomparison activity.  %Three
stages:
%
%{\bf Data Acquisition (Modeling)}
%
%\begin{itemize}
%\item Climate centers run prescribed set of common experiments; produce 2-10 PB
of data
%\begin{itemize}
%\item data can also come from sensors - here, the data would still be
post-processed before it would be published.
%\end{itemize}
%\item Centers can publish their own output data or send it to another center to
publish data generated over \~{}2 year period (then post-processed and published
over a few more months)
%\item Properties: data-intensive, distributed centers, lots of computing
%\end{itemize}
%
%
%
%{\bf Data storage (Infrastructure/system)}
%
%\begin{itemize}
%\item ESGF develops and deploys a federated network of gateways and associated
data nodes
%\item As models run at each center, output data is post-processed into packages
with common formats and prescribed metadata conventions
%\item Most centers will deploy data node software stack, and using this, they
manage the data from their experiments
% \begin{itemize}
% \item the data node software stack scans the data, makes sure it has the right
metadata fields, does QA/QC, build a set of catalogs of the prepared data
% \begin{itemize}
% \item minimally, catalog provides HTTP links to data elements
% \item also can provide GridFTP endpoints
% \item also can provide product services that abstract the dataset in other
ways - get whole file, subset, browse, etc.
% \end{itemize}
% \end{itemize}
% \item When the center is happy with the data/catalog in the data node, they
publish it to a host/affiliated gateway
%\begin{itemize}
% \item this submits the catalog to the gateway
% \item them the gateway shares this catalog with other gateways so that all
gateways have a consistent view of all the published data
%\end{itemize}
% \item Core archive (1-2 PB, most popular data)
%\begin{itemize}
% \item will be replicated at several sites
% \item replication activity manually requested/initiated by a gateway owner
% \item replication: copies catalog to replica site, do data transfer to new
data node, then publish catalog, push catalog to (this) gateway associated with
new data node
%\end{itemize}
% \item Currently building notification service
%\begin{itemize}
% \item track which users have downloaded data - if a problem is found in a data
set or if new data is added to a data set, notify users who have
downloaded/replicated it
%\end{itemize}
% \item Properties: data-intensive, distributed gateways and data nodes, data
appears in the system over time (dynamic)
%\end{itemize}
%
%
%
%{\bf Data processing (Application)}
%
%\begin{itemize}
% \item Users ($\sim$20000 as of Dec. 2010) can:
%\begin{itemize}
% \item browse/search a catalog at any gateway and locate data, which might be
hosted by a data node affiliated with another gateway. (can output a wget script
that can later fetch the data)
% \item authenticate, gain access to a group
% \item download data via http or GridFTP or access product services (uses data
retrieval syntax (DRS) - can be scripted)
% \end{itemize}
%\end{itemize}
%
%\begin{itemize}
% \item abstractions that can be used by a programmer: data is stored in NetCDF
format, uses CF metadata conventions, can be accessed using DRS
% \item Some users will analyze data from a single model
% \item Many applications are multi-model analyses - many users want to look at
the same parts of the output of some/all of the models, to understand if and how
the models differ
% \item Some centers will gather some/all of the core archive (plus more,
perhaps) on local systems for local users to perform ``power analysis''
%\end{itemize}

\subsection{Interactive Exploration of Environmental Data \label{envJon}}

%\katznote{from Jon Blower}
%
%\katznote{summary below assigned to Yogesh}
%\subsubsection{new summary}

% 1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

In the environmental sciences, the use of visualization techniques is vital for
understanding the ever-increasing volume and diversity of data that is being
produced by Earth observing systems and computer simulations.  The primary
purpose of visualization is to gain insight but the majority of scientific tools
generate static plots that neither the originating scientist nor the recipient
can easily customize to reveal new information. The issues with visualization of
a single dataset are compounded if multiple datasets are to be examined
simultaneously, as is common in environmental science for model validation,
``ground truthing,'' quality control, and data assimilation.

New, interactive modes of environmental data exploration and visualization based
on the principles of simplicity, open standards and user-friendliness are
required at all stages of scientific investigation.  The application described
here is an {\em infrastructural application} that uses stored data and consists
of one stage: data processing.  It represents a number of similar
applications~\cite{blower1,blower2}.

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}

The applications often take the form of graphical Geographic Information Systems
(GIS) tools, but with better support for large and multidimensional scientific
data. The general vision is of a map-based interface, onto which different
datasets can be overlain.  A processing step may be initiated by rubber-banding
an area of the map and selecting from a list of algorithms that process data
within the selected area, perhaps calculating statistics of the data.  In a
desktop application, this processing may take place on the user's desktop, but
in a web-based application the processing must take place on a server, which may
or may not be co-located with the datasets.  There is therefore the common
problem of moving large amounts of data around in a distributed system, to which
may be added concerns of security in cases in which the data in question are not
public.

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below}

Data from instruments is small (1--100~MB) but model output may be very big
(GB--TB). The data is multi-dimensional in nature. Data aggregated over time
from instruments may need to be visualized (e.g. as an animation), causing the
visualized data size to grow.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

%\katznote{from answer to question 3 below}

 Increasingly, environmental data are distributed in multiple locations.  The
applications are driven by data served through distributed web services, that
may in turn be ``fed'' instruments or computer simulation generated
data. Compute services need to be integrated with these data services in an
efficient, easy-to-use, and transparent way, allowing fast data processing.  The
processing algorithms are generally simple and high overhead scheduling systems
can cause a loss of responsiveness.  The infrastructure has more in common with
the Web (or cloud) than the grid.

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

Dynamic data are sometimes used, but may arise from real-time feeds from
instruments or from looking at live results from a model. Data hosted at
multiple locations may be frequently updated (often several times a day).  The
data services are designed to allow server-side subsetting of large datasets,
attempting to minimize the amount of unwanted data traveling across networks.
These make it hard to achieve scalability through caching.

In a Web environment, servers and proxy servers (which are sometimes beyond the
control of the data provider or the data user) may retain caches of data in a
strategy to reduce server load and network traffic.  For this reason, the HTTP
protocol provides mechanisms for defining expiration times for data resources,
specifying when a resource ought to be cleared from the cache.  In the
environmental and geospatial communities the Open Geospatial Consortium
specifications are being widely adopted alongside existing protocols such as
OPeNDAP~\cite{opendap}.  Unfortunately these protocols, although built atop
HTTP, do not make it easy for this versioning to be implemented correctly.  The
OPeNDAP protocol does not have the concept of a dataset version, meaning that
clients have no reliable or efficient means to detect that a dataset has
changed.  The OGC protocols, in general, \emph{do} provide versioning at the
level of the service endpoint, but provide no information on when a resource
should be considered expired.  Clients are therefore forced to poll the server
to check for updates.

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

The application aims for near-real-time interaction with the data, i.e the user
should not wait more than a few seconds for some kind of response to a request,
so performance with low latency is a key challenge.  Security is another serious
concern: many environmental datasets are held under access control, and
different providers often have very different access control policies.  In some
infrastructures, these different policies are handled via a role-mapping
mechanism~\cite{nerc_data}.  Concerns of access control---as well as data
volume---also make it difficult for data to be replicated to different
geographic locations (and hence different access control regimes).  In a
distributed environment, it is common for machines to access data on behalf of
end users (e.g., a processing service might download its input data from a
remote store).  Finding a means for the user to delegate his/her authority to a
multi-web-service infrastructure is a key current challenge.  The MashMyData
project~\cite{mashmydata} is investigating two solutions, based on Grid proxy
certificates and OAuth.

% \subsubsection{old material follows}

% (Here's a synopsis of the chapter I'm writing for Malcolm's book on
Data-Intensive Research.  % I'm not sure how to merge this with the answers to
the pro-forma questions below.  Also, % many of these sentences are
cut-and-pasted from the book chapter - future copyright issue?)

% In the environmental sciences the use of visualization techniques is vital for
understanding % the ever-increasing volume and diversity of data that is being
produced by Earth observing % systems and computer simulations.  The primary
purpose of visualization is to gain insight % \cite{card_readings_1999}, but the
majority of current tools in wide use are not adequate % for achieving this.
The time spent to create visualizations is too great, usually due to the %
crippling overhead of low-level technical tasks such as file-format and metadata
interpretation.  % Most scientific tools generate static plots, which are shared
in documents or emails, but % neither the originating scientist nor the
recipient can easily customize or adapt the visualization % to reveal new
information.  The context of how the visualization was generated is usually lost
% in the publication/sharing process, making interpretation and reproduction
much more difficult.

% Visualization of a single dataset is therefore more tedious and less useful
than it should be.  % These problems are compounded if the scientist wishes to
examine multiple datasets simultaneously.  % This is an extremely common task in
environmental science, which forms an important part of model validation, %
``ground truthing'', quality control and data assimilation.

% There is much current interest in developing new, interactive modes of
environmental data exploration and visualization, which can % play a key role at
all stages of scientific investigation.  We consider that, for the case of
interactive % data exploration, concerns of performance, % intuitiveness and
user engagement tend to outweigh the desire for sophistication in data
processing.  % We therefore argue that a new generation of ``4D GIS'' tools,
based on the principles of simplicity, % open standards and user-friendliness is
required to gain greater insight from the environmental data % deluge.  These
new tools will not replace existing approaches, but rather complement them.

% Increasingly, environmental data are distributed in multiple locations, and
are frequently updated % (often several times a day).  Interactive exploration
of such large, dynamic, distributed data is therefore % a strong technical
challenge.  Data infrastructures must support the ability to quickly extract %
subsets of large data, and any data processing must happen quickly, with the
minimum of overhead.  % Web services are therefore playing an
increasingly-important role, and there is much interest in % cloud technologies
to provide the necessary scalability at the back end.  Grid technologies, being
essentially batch-mode % systems, are less suitable for this kind of application
as the scheduling overhead tends to dominate % the processing time for simple
jobs, seriously hampering interactivity.

% Performance is therefore a key challenge, exacerbated by the dynamic nature of
the data, which makes % it hard to achieve scalability through caching (see
section~\ref{sec:dynamicdataweb}).  [Note: could probably % merge material from
that section here.]  Security is another serious concern: many environmental
datasets % are held under access control.  In a distributed environment, it is
common for machines to access data % on behalf of end users (e.g., a processing
service might download its input data from a remote store).  % Finding a means
for the user to delegate his/her authority to a multi-web-service infrastructure
is % a key current challenge.  The MashMyData project
(http://www.mashmydata.org) is investigating two solutions, based % respectively
on Grid proxy certificates and OAuth.

% \subsubsection{quick answers from Jon}

% 1.  What is the purpose of the application?

% We are actually concerned with a class of applications (highly % interactive,
graphical ones) rather than a specific one.  But % generally, the applications
involve the intercomparison of different % environmental datasets for purposes
such as model validation, ground % trothing of observations and data
assimilation.

% 2.  How is the application used to do this?

% The applications often take the form of graphical tools that share a lot in
common with Geographic Information Systems (GIS). However these applications are
much better suited to scientific data, which are large and multidimensional.
The general vision is of a map-based interface, onto which different datasets
can be overlain.  A processing step may be initiated by rubber-banding an area
of the map and selecting from a list of algorithms that process data within the
selected area, perhaps calculating statistics of the data.

% 3.  What infrastructure is used? (including compute, data, network,
instruments, etc.)

% Usually the applications are driven by data served through web services of
various kinds.  These web services may be 'fed' by data coming from instruments
or from computer simulations.  A key challenge is how to integrate compute
services with these data services in an efficient, easy-to-use and transparent
way, allowing fast data processing.  The processing algorithms are generally
quite simple and so it is frustrating to succumb to the high overhead of some
scheduling systems.  The infrastructure probably has more in common with the Web
(or perhaps cloud) than the Grid.

% 4. What dynamic data is used in the application?  % (a) What are the types of
data, (b) What is the size of the data set(s)?

% Dynamic data are only sometimes used, but may arise from real-time feeds from
instruments or from looking at live results from a model.  Data from instruments
are small 1~MB-100~MB but model output may be very big (GB-TB).

% 4.  How does the application get the data?

% Generally through web services.  These services are designed to allow
server-side subsetting of large datasets, attempting to minimize the amount of
unwanted data traveling across networks.

% 5.  What are the time (or quality) constraints on the application?

% We try to aim for near-real time interaction with the data, i.e the user
should not be waiting more than a few seconds for some kind of response to a
request.  We have largely solved this problem for simple visualization, but for
linking data access and processing we are not there yet.

%\subsection{Dynamic Environmental Data on the Web} \label{sec:dynamicdataweb}

% Key contents now merged into above section.  Specifics on the NERC VO % are
probably not relevant at the moment as they are not yet implemented.

% \jhanote{I have moved this from erstwhile \S 2. Jon: please either (i) %
convert into a stand-alone application, or (ii) merge with the % ``Interactive
Exploration of Environmental Data'' example}

% \katznote{summary below assigned to Yogesh}

% 1. {\bf What does the application do? What problem does it solve?}

% The NERC\footnote{UK Natural Environment Research Council} Virtual %
Observatory\footnote{http://www.nerc.ac.uk/research/programmes/virtualobservatory/}
% is a new initiative to create a pilot cloud-based environment in which %
different practitioners will discover, visualize and process data % relating to
the UK's system of rivers and soils.  It will provide % access to many different
kinds of distributed hydrological data, % including sensor data and numerical
models, many of which produce % rapidly-changing, dynamic information.

% 2. {\bf How does it do it?}

% There are many scientific % challenges to be faced, notably including the
coupling of the models % and observations; this includes the possibility that
sensors can be % dynamically tasked in response to model forecasts, or that
models may % adapt dynamically to new information from sensors.

% 3. {\bf What are the big data aspects?}

% Not known.

% 4. {\bf What are the distributed aspects?}

% Hydrological data, % including sensor data and numerical models, are
distributed.  % Principles of the % Web and ``linked data'' may help to solve
some of the challenges in % informatics, and % Web Services are commonly used to
distribute data through HTTP.

% 5. {\bf What are the dynamic aspects?}

% Sensor data and numerical models produce % rapidly-changing, dynamic
information.  % A key characteristic of dynamic data, of course, is that it
changes with % time.  It is of great importance for the user to be confident
that she % is accessing the correct version of the data; this will often be the
% most up-to-date data, or may be a specific historical version.  In a % Web
environment, servers and proxy servers (which are sometimes beyond % the control
of the data provider or the data user) may retain caches % of data in a strategy
to reduce server load and network traffic.  For % this reason, the HTTP protocol
provides mechanisms for defining expiration % times for data resources,
specifying when a resource ought to be % cleared from the cache.

% In the environmental and geospatial communities the Open Geospatial %
Consortium specifications are being widely adopted alongside existing %
protocols such as OPeNDAP.  Unfortunately these protocols, although % built atop
HTTP, do not make it easy for this versioning to be % implemented correctly.
The OPeNDAP protocol does not have the concept % of a dataset version, meaning
that clients have no reliable or % efficient means to detect that a dataset has
changed.  The OGC % protocols in general %[TODO: look into this], % \emph{do}
provide % versioning at the level of the service endpoint, but provide no %
information on when a resource should be considered expired.  Clients % are
therefore forced to poll the server to check for updates.

% Many technical challenges will remain regarding the % handling of dynamic
data, including keeping careful records of its % provenance in order to record
the steps taken in making a decision.

% 6. {\bf What else is important?}

\subsection{Power Grids \label{power}}

%\katznote{from Yogesh}
%
%\katznote{summary below assigned to Yogesh}
%\subsubsection{new summary}

% 1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

Demand response (DR) optimization~\cite{ferc2011DR} in smart power grids deals
with ensuring an adequate supply of electricity by targeted curtailment of power
load by consumers during periods of peak load to avoid blackouts. One
application used for DR optimization is power demand forecasting at coarse and
fine temporal and spatial granularities to allow load curtailment operations to
be triggered in advance of a demand-supply mismatch situation.  This is a {\em
infrastructural application}.  It begins after data generation, and includes the
data transformation \& decisions, storage, and processing stages.

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}

Power usage information is aggregated from smart meters at consumer premises
into the utility's data center~\cite{Simmhan:buildsys:2011,simmhan2011TR}. Data
is collected at approximately 15-min intervals and transmitted to the utility
through cellular and wireless networks using proprietary protocols. Complex
event pattern matching over the power usage events and other information source
such as weather and scheduling data accessed using web services help perform
near term load forecasting. Long term power forecasting at the city-scale or at
a smaller micro-grid scale (e.g. industry complex, university campus) is done
using power usage models built using machine learning algorithms running at the
utility's or micro-grid's datacenter using historical usage
data~\cite{Aman:dddm:2011}. These applications run on private clouds and are
composed using DAG, Complex Event Processing and MapReduce
structures~\cite{zhou:2012:scepter}.

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below}

Constructing forecast models using machine learning requires access to
historical data that can grow to large sizes~\cite{Yin:mapreduce:2012}. Several
hundred TB per year of smart meter power usage data can accumulate for a large
city with millions of customers. Other data used in predictive modeling, such as
event/people schedules and building/facility features, are smaller, on the order
of GBs per year. In future, social network data from consumers may also be mined
to learn about power usage patterns. These can also be on the order of GBs.

Real-time forecasting uses smaller data---events on the order of 1~KB, but
potentially from millions of sources---that are passed to complex event
pipelines or predictive models that run continuously. These pipelines can
process GBs of data per hour.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

%\katznote{from answer to question 3 below}

Data from domains such as power systems, building management systems, weather
and traffic, and social networks need to be integrated. While the data are
generated and/or stored in distributed locations, they are typically aggregated
at the utility's datacenter for analysis. Limited processing may happen locally
at the smart meters or local micro-grids with the results pushed as events to
the utility.

The applications themselves run on private cloud platforms across hundreds of
VMs. Communication between the utility and smart meters is bidirectional,
allowing control and pricing signals to be sent from the datacenter to the smart
meters. Integrated data will also need to be shared with external applications
running on, say, mobile platforms or portals, and with regulatory agencies.

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

The data rates for events published by the sensors may change, either passively
due to local measurement limitations or actively throttled by the utility to
meet the application's accuracy needs/resource
constraints~\cite{Simmhan:sciencecloud:2011}. The data from smart meters can be
sent synchronously as they are generated or in batch mode once a day to conserve
bandwidth. The application processing itself is done elastic resources, scaling
up or down to meet latency requirements. Some of the forecasting models can also
decide to listen to additional data streams on-demand when higher accuracy is
required or special situations are encountered.

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

Semantic information integration is being performed using an ontology composed
together from individual domains~\cite{Zhou:itng:2012}. Machine learning models
using Matlab, Weka and Hadoop are being used. Complex event processing is using
the ``Siddhi'' CEP engine with semantic support added~\cite{siddhi}. Stream
processing on Eucalyptus using IBM Infosphere Streams is being performed.

% \subsubsection{old material follows}

% Consider the energy informatics domain and smart power grids in %
particular. For data continuously arriving from 2 million smart meters % in
large city, households will soon need to be continuously analyzed % in order to
detect impending peak power usage in the smart power grid % and notify the
utility to respond by either spinning up additional % power sources or by
triggering load curtailment operations to reduce % the demand. This closed loop
cyber-physical application, modeled as a % workflow, needs to combine streaming
data arriving from sensors with % historic data available in file archives along
with structured % collections of weather forecast data that helps the large
scale % computational model make an energy use prediction in near real time.

% A workflow framework that supports this data model diversity, % including
streaming data, structured collections and files, and the % ability to execute
reliably and scalably on resources is required.

% In general, smart grids - respond to varying data loads, spinning up %
additional power sources or triggering load curtailment operations.  % E.g., of
a closed-loop cyber-physical system modeled as a workflow % needs to combine
streaming data arriving from sensors with historic % data available in archives,
along with structured collections of % weather forecast data. Need a workflow
model that supports such % requirements. Taken from ref (Towards Reliable
Performant Workflows % for Streaming Applications on Cloud Platforms).

% {\it Requirements: } (i) Applications will have to span desktops, %
workstations and clouds, (ii) streaming data models and stream % programming
abstractions that are more than TCP sockets, (iii) % reliability of VMs hosting
the workflow and avoid costly duplicate % movement of the same logical stream
(see paper in WORKS by Zinn et % al.)

% Scientific workflows are commonplace in eScience applications. Yet, % the lack
of integrated support for data models, including streaming % data, structured
collections and files, is limiting the ability of % workflows to support
emerging applications in energy informatics that % are stream oriented. This is
compounded by the absence of Cloud data % services that support reliable and
performant streams.  Need a % scientific workflow framework that supports
streams as first class % data, and is optimized for performant and reliable
execution across % desktop and Cloud platforms.

% \subsubsection{old version of answers}

% 1. What is the purpose of the application?

% Demand response optimization (DR) in smart power grids. Specifically, %
accurate forecasting of power demand at coarse and fine temporal and spatial %
granularities, and launching power load curtailment operations when a peak %
load situation is forecast.

% 2. How is the application used to do this?

% Several approaches exist for performing DR.

% (1) Perform realtime aggregation and analysis on power usage information %
arriving into the utility's data center from smart meters installed at %
consumer premises for low latency detection of current (peak) load;

% (2) Perform complex event pattern matching and mining over an event cloud %
that includes event streams from multiple information source types to detect %
pattern that forecast a peak load event or opportunity of energy % conservation.

% (3) Construct building and campus energy models using machine learning to %
predict future energy use.

% 3. What infrastructure is used? (including compute, data, network, %
instruments, etc.)

% Compute: 128-core Eucalyptus private cloud; Public Cloud (TBD)

% Data: Local disk, others (TBD)

% Software/Services: Eucalyptus Cloud fabric, IBM InfoSphere Streams, %
Esper/Siddhi Complex Event Processing (CEP) system, MatLab toolkit, Workflow %
system (TBD), custom research prototypes.

% Network: Ethernet, others (TBD) % Instruments: Smart meters (consumer power
usage), Power sensors (fan speed, % room temperature, thermostat setting,
appliance state, etc.), Facility % sensors (people counter, ambient light, etc.)

% 4. What dynamic data is used in the application?  % (a) What are the types of
data, (b) What is the size of the data set(s)?

% Current and expected data \& rates:

% Smart meter power usage data: 1.4M Consumers x $\sim$1 KB event/min
($\sim$700~TB/yr (max))

% Weather: KB/hr

% Traffic: KB/min

% Consumer Calendar Schedules: MB/hr

% Facility Calendar Schedule: MB/day

% Twitter/Facebook: MB/min

% 5. How does the application get the data?

% Web services, file transfer, proprietary network protocols

% 6. What are the time (or quality) constraints on the application?

% O(secs) to O(hrs) for load forecasting and response, depending on accuracy %
of prediction and how far ahead the prediction is. There is a tradeoff in % the
quality constraints based on available power capacity, data, % infrastructure
resources, granularity of prediction and response, etc. The % absolute quality
constraint is that the peak power load must {\bf not} exceed % the available
power capacity. So the power forecast, load curtailment % activation, and the
impact of the curtailment have to be in a closed loop % cycle whose equilibrium
is maintained.

% 7. How much diverse data integration is involved?

% 8. How diverse is the data?

% Data from domains such as power systems, building management, weather and %
traffic, and social networks need to be integrated. We use semantic %
information integration with an ontology composed together from individual %
domains.

% 9. Please feel free to also talk about the current state of the % application,
if it % exists today, and any specific gaps that you know need to be overcome.

% Currently, application for constructing Building and Campus energy models %
using machine learning and forecasting power usage exist. These over power %
usage data, building construction features, academic schedule, weather, etc.  %
Uniform information integration is currently underway to experiment with %
multiple models.

% Initial versions of stream processing on the private Cloud over campus level %
smart meter data using information from 100+ buildings exist with adaptive %
logic for stream rate control.

% Research into semantic CEP is underway, with domain patterns being defined.  %
Gaps include scalable stream and CEP processing over clouds (including % advance
resource provisioning and task scheduling), running ``online'' % workflows with
support for stream inputs that run continuously (to improve % forecast models
and make predictions), resource selection based on data % privacy and latency
constraints, and reliable execution with graceful % degradation of critical
analysis operations.

\subsection{Fusion (ITER) \label{fusionScott}}

%\katznote{from Scott Klasky}
%
%\katznote{summary below assigned to Dan}
%\subsubsection{new summary}

% 1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

With ITER~\cite{iter} scheduled to operate before 2020, plasma fusion has become
one of the highest priority research items for the US DOE. Scalability of
first-principles fusion codes is an important part of fusion research needed to
obtain predictive capability of plasma performance with first-principles physics
fidelity. The goal is to build a realistic full-distribution kinetic code, in
realistic ITER geometry, combining the strengths of each of the existing
gyrokinetic codes (GEM, GTS, XGC, and GTC), and utilizing the most up-to-date
computer science, hardware design, and applied mathematics technologies
developed. This {\em traditional application} is intended to be capable of 1)
discovering macroscopic magnetohydrodynamics activity in burning plasmas, 2)
understanding Energetic particle effects in fusion reactors, 3) understanding
the heating in a fusion reactor, and 4) understanding the interaction of the
plasma at the material boundary. This applications just has the data processing
stage, along with associated data movement.

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}

The applications that are currently run (XGC1, XGCP, GTC, GTS) run on all of the
DOE and TeraGrid leadership class facilities, and require very large resources
for the modest problems that are currently being solved. In the next generation
(exascale), not only will all of the computing power that will be available at
the time be needed to understand ITER, all of the data that will be generated
during the simulation also will have to be understood. The community's
techniques have been to examine and understand the data in situ, to reduce the
total amount of data produced in the simulation.

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below}

Today the team has produced O(100) TB per week for a simulation using the ADIOS
middleware~\cite{adios} incorporated in the simulations. Using staging
techniques, they believe that they will only produce O(10) PB on an exascale
machine.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

%\katznote{from answer to question 3 below}

The team always runs on all available resources, so if they have computer time
across the country, then their data will be distributed, and so will the
computation.

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

The primary aspect that is dynamic is that the set of computing resources that
are used for a given run are based on those that are available at the time of
that run.  As part of this, data is streamed from one running simulation to
another, and is transformed in-flight.

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

The applications run on a distributed set of leadership-class facilities, using
advance reservations to co-schedule the simulations.  Each code reads and writes
data files, using ADIOS and HDF5.  Files output by each code are transformed and
transferred to be used as inputs by other codes, linking the codes into a single
coupled simulation.  Because the overall data generated is so large, it can not
all be written to disk for post-run analysis, so in-situ analysis and
visualization tools are being developed.

%\subsubsection{old material follows} %1. What is the science/engineering
problem? What is the purpose of the %application?
%
%With ITER scheduled to operate before 2020, plasma fusion has become %one of
the highest priority research items for DOE. Scalability of %first-principles
fusion codes is an important part of fusion research %needed to obtain
predictive capability of plasma performance with %first-principles physics
fidelity. We want to build a realistic %full-distribution kinetic code, in
realistic ITER geometry, combining %the strengths of each of the existing
gyrokinetic codes (GEM, GTS, %XGC, and GTC), and utilize the most up-to-date
computer science, %hardware design and applied mathematics technologies
developed. The %application that we will develop will be capable of 1)
discovering %Macroscopic Magneto Hydro Dynamics activity in burning plasmas, 2)
%Understand Energetic particle effects in fusion reactors, 3) %Understand the
heating in a fusion reactor, and 4) Understand the %interaction of the plasma at
the material boundary.
%
%2. How will the application solve the problem?
%
%(a) Which of the following describes the application
%
% i) large (exascale) data and large (exascale) computing (e.g., data mining,
analytics, bioinformatics)
%
% ii) large (exascale) data and modest computing (e.g., search)
%
% iii) modest data and large (exascale) computing (e.g., simulation/modeling)
%
%The applications we currently run (XGC1, XGCP, GTC, GTS) run on all of the
leadership class facilities, and require very large resources for the modest
problems which we are currently solving. In the next generation (exascale) we
will not only need all of the computing power that will be available at the time
to understand ITER, we will also need to understand all of the data that will be
generated during the simulation. Our techniques have been to examine and
understand the data in situ, to reduce the total amount of data produced in the
simulation. Today we have produced O(100 TB/week) for a simulation using our
ADIOS middleware incorporated in our simulations. Using our staging techniques,
we believe that we will only produce O(10 PB) on an exascale machine. (I don't
know if this is considered modest dat, or extreme size data)?
%
%(b) Can you quantify the amount of computing vs the amount of data?
%
%1 PB of data -- 10 PB of data, for an exascale size computation run in O(100
hours).  %Today we achieve about 6\% of peak with excellent scalability to 220K
%cores on the Cray XT systems. It is %difficult to predict how we will run on
%accelerators, and next generation machines.
%
%3. What infrastructure will be used? (including compute, data, %network,
instruments, etc.)
%
%For Validation, we will need to compare our data from the computer to %ITER (or
current fusion reactors). Typically the amount of data used %in comparison is
tiny by comparison (people only look at MB's form %experiments).
%
%(a) What will be distributed?  data, compute, both?
%
%We always run on all available resources, so if we have computer time %across
the country, then our data will be distributed, and so will our
%computation. ITER will have data that will be distributed, but it %still
unknown.
%
%4. What data be used in the application?
%
%I don't understand this question? We will input data (from perhaps a restart),
and this data is typically in the ADIOS-BP file format for large data. Smaller
data sets are typically stored in HDF5, and the smallest are stored in MDS+.
%
%(a) What are the types of data
%
%(b) What is the size of the data set(s)?
%
%Depends on what we want to understand --- For GTS/GTC we have typically %looked
at the particle data to compute fluxes, etc. and this data can %be large, but
due to the Monte Carlo error, and ways to generated 5D %distribution functions
from the particles, we reduce the particles to %a manageable amount (300 GB per
time step x 1 K time steps). Our %current I/O is capable of reading this in $<$
100 s per time step, which %allows us to analyze the data in about 1-2 days.
%
%(c) Will any data be dynamic?
%
%Don't understand the question. Features change and are dynamic. The physics
will get more complex as we continue to evolve our codes, and will be more
dynamic in nature.
%
%5. How will the application get the data?
%
%Reading the data from disk --- (Perhaps I don't understand the question).  It
will also couple to other codes, and the coupling will be in-memory data
exchange.
%
%6. Are there time (or quality) constraints on the application?  (If so, what
are they?)
%
%We always need to run as fast as possible since we have limited computing time
(our team does have about 100M hours FY 11, but this still puts severe
restrictions on what we run, so the faster we can run, the more science we can
get done.
%
%7. Please feel free to also talk about the current state of the application, if
it exists today, and any specific gaps that you know will need to be overcome.
%
%Our biggest gaps will be in accelerating the code on next generation exascale
platforms, and in processing as much data as possible in situ.

\subsection{Industrial Incident Notification and Response \label{Kees}}

%\katznote{from Kees}
%
%\katznote{summary below assigned to Yogesh}

%\katznote{Yogesh, some parts are repetitive - can you fix them?}

%\subsubsection{new summary}

% 1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

Applications such as large scale crisis management, maritime security, disease
and pollution monitoring, and environmental control need to fuse large amounts
of heterogeneous data to form virtual information sources that mitigate
noise~\cite{emergency1,emergency2,emergency3,emergency4}.  The area of greater
Rotterdam in the Netherlands is a chemical hub and houses more than 15\% of the
Dutch population, making safety and security a major concern.  Some leakage of
chemicals into the air is an everyday, unavoidable event.

The environmental protection agency (EVA) maintains a command and control call
center staffed by chemical hazard experts to (1) monitor the air quality and
sewage, (2) in case of an incident locate the source, (3) identify the threat
and (4) provide the public authorities, the emergency response organization, and
the public with situation awareness and suggested actions.  The information that
is shared is often incomplete, uncertain and inconclusive. Knowledge within the
EVA as well as from outside experts and industries must be accessed to
meaningfully use that information.  This subsection describes the {\em
traditional application} that the EVA uses for these tasks.  The application
starts after data generation, and includes data transformation \& decisions,
storage, and processing.

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}

For monitoring and incident detection, continuous streams of data are processed
and scanned for anomalies from a small number of physical sensors and air
analysis data (gas-chromatographs outputs).  If an incident or an anomaly is
detected, new streams of data become available from the general public and field
inspectors and must be processed. Static information on weather and about
chemicals are also needed.  These may be stored in different places, access may
not be granted automatically, and validity and update frequencies may not match
expectations. Thus specialized pre-processing is needed.

Different knowledge-based processing models must be used account for uncertainty
in the data and sensor behavior, and provide robust fusion (processing) for
situation awareness and decision support. Some confidential data must be
processed \emph{in situ}. New intermediate and final data that are generated may
have different retention characteristics and need to be stored and made
accessible for visualization and analysis in different workflows.

Processing in done in semi-real-time or in batch mode processing on a small,
private computer infrastructure. Therefore, scalability is lacking and the
dynamics of the workflows is kept under strict human control to match the
capabilities of a set of static resources.

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below}

The data sizes are not very large.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

%\katznote{from answer to question 3 below}

A number of special data and sensor interface processes will be distributed over
various platforms and fixed in place.  A low frequency stream of reasonable
fidelity data from a small number of physical sensors and an even lower
frequency stream of air analysis data (gas-chromatographs outputs) is
continuously trickling in.  Low fidelity human provided data from local people
complaining to the call center, field inspectors calling in with observations
and measurements, information from social media, pictures from cell phones and
emails sent to a centralized email account are entered manually into a local
databases.  Static information such as current weather conditions and forecast
and information about chemicals are also needed.  These data are made available
via a distributed shared data space.  Some confidential data must be processed
remotely and only aggregated results communicated back to other entities. The
information processing algorithms are packaged as autonomous software agents
that are managed by a multi agent system infrastructure (middleware) and humans
(via GUIs). Processing algorithms need to access multiple concurrently and
remote data stores.

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

Both the data and the processing are dynamic in nature. Data that cannot be
moved due to confidentiality have to be processed remotely by moving
applications to them.  A process integration framework will use agents that
provide autonomous and composable resources to dynamically construct workflows
for data driven processing.  If an incident happens or an anomaly is detected,
it triggers new streams of data about other anomalies that must be processed.
Another source of dynamic behavior is when the event is escalated and additional
management processes come into play, each requesting and adding their own sort
of processing to the mix.

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

Quality of service aspects that are important for real world and real-time
applications are the dynamics of scheduling from an application perspective,
multi independent level of security support, processing service discovery
support, autonomous reconfiguration support and logging-for-traceability.  Most
of the dynamic data that these applications use has a short validity period,
i.e., it must be processed now or we process the next update coming in. The data
is uncertain, i.e., it may have dynamically changing signal to noise ratio and
although expected at some base frequency, may not be available for some period
of time.

% \subsubsection{old material follows}

% Efficient Processing in Real-Time Complex Distributed Information Systems

% Dr. Kees Nieuwenhuis, Thales Nederland, D-CIS Lab

% Acknowledgement: The use case and working prototype system discussed here are
% realized with the help of the DIADEM project grant from the EC (FP7 grant
224318).

% Introduction

% Contemporary net-centric applications, such as large scale crisis management,
% maritime security, disease and pollution monitoring, environmental control and
% management etc. cab greatly benefit from the fusion of large amounts of
heterogeneous % information, coming from static and mobile sensors as well as
intelligence obtained % from human observers, various databases and the World
Wide Web. This is because % new AI based techniques are available that help to
combine multiple information % modalities and thereby increase the amount of
information for (real-time) fusion % systems, which in turn can mitigate the
impact of noise and reduce ambiguities and % overcome the unavailability of
specific information sources.

% However, the compute and networking requirements for this new type of
applications % that we need to build are significantly different from the more
traditional massive data % processing or distributed computing applications that
are known and deployed in % professional business and scientific domains today.

% Use case description: DIADEM

% The area of greater Rotterdam in the Netherlands is a highly industrialized
area with the % largest harbor in Europe, very many chemical, petrochemical,
pharmaceutical and other % processing plants. Economically, it is the most
important area of the country. In addition, % the same area is home to more than
15\% of the Dutch population. Safety and security % are therefore a major
concern, especially with the type of hazards involved and % because major
evacuations are just not an option. Some leakage of chemicals into the % air
however, is an every day event and cannot be avoided. Therefore, the %
environmental protection agency (EVA) is actively monitoring air quality and
incidents % (that by law must be reported immediately) from a command and
control center which % includes a dedicated call center for the population. They
employ a number of chemical % and other hazard experts that can be deployed in
the field for inspection (prevention) % and measurements. Due to the sheer size
of the area and the (so far) inadequate % techniques and resources for
continuous data gathering, data sharing, data processing % and sensor
management, there are very few sensors in the area that the EVA has % continuous
access to.

% The tasks of the EVA are to 1) monitor the environment (mainly air quality and
sewage), % 2) in case of an incident locate the source, 3) identify the threat
and 4) provide the % public authorities, the emergency response organization and
the public with situation % awareness and possible courses of action. For that,
they must liaise and coordinate % with other expert organizations (health and
hazard experts, public order and safety % experts, the harbor authority, traffic
experts, etc.) and share and exchange information % and knowledge. An important
aspect of the latter is that the information that is available % is always
incomplete, uncertain and no single element is conclusive by itself. In order to
% process and use that information, inside and outside knowledge must be
accessed.

% Dynamic versus static data

% For continuous monitoring and incident detection continuous streams of data
are % necessary that need to be processed and scanned for anomalies. Typically,
a low % frequency stream of reasonable fidelity data from a small number of
physical sensors % and an even lower frequency stream of air analysis data
(gas-chromatographs outputs) % is trickling in. A much larger amount of data
from private sensors from hazard sources is % potentially available, but
currently not accessible. If an incident happens or an anomaly % is detected,
new streams of data become available and must be processed, such as low %
fidelity data from people in the area calling the call centre with complaints,
field(ed) % inspectors calling in with results of expert observations and
additional high fidelity % measurements, information from social media and maybe
pictures from cell phones % from individuals in the area (solicited or
not-solicited). In case of an incident, suddenly % also other more static
information such as current weather conditions and forecast and % information
about chemicals (hazards, risks, air-borne or water-borne behavior, health %
aspects, neutralizing methods and agents, etc.) is needed for further
processing. This % other information may be stored in different places and
access may or may not be % granted automatically and formats may be special and
validity and update frequencies % may not match expectations or needs and thus
demand for specialized pre-processing % emerges.

% Use of data

% The real-time data is mainly stochastic in nature. The fidelity is variable,
which means % that different knowledge based processing models must be used
concurrently to take % into account not only the uncertainty in the data /
measurements itself, but also to take % into account sensor behavior. Most of
the data are observations of the effects of the root % cause and therefore
robust fusion (processing) with the help of knowledge based % models is
necessary to provide situation awareness and decision support. Some fusion %
requires data that is confidential in nature or commercial in confidence and may
not % leave the place where it is stored. Therefore the special event based
processing % algorithms must be brought to the data and only aggregated results
can be % communicated to other processing entities. This means that both the
data and the % processing are dynamic in nature, which the application and its
computer infrastructure % must support. Relationships with static or semi-static
(e.g., weather) data are often built % in to the processing algorithms, which
means that these data sources need to be % accessed during real-time processing
also. This requires multiple simultaneous % accesses to remote data stores. The
processing will also produce new intermediate and % final data, sometimes with
different retention characteristics, which need to be stored % and made
accessible to special visualization and analysis applications in different %
workflows.

% Infrastructure use

% In current applications the three main network technologies used are the
telephone and % cell phone network and a centralized computer network with
internet connection (for % emails). Data is filtered and entered manually into
locally housed databases and files % and processed in semi-real-time or batch
mode processing on a private and limited % computer infrastructure. Therefore
scalability is lacking and dynamics of workflows is % kept under strict human
control to match the capabilities of a set of static resources.

% The need however, is to have more scalable solutions where resource usages can
% actually follow demand from an escalating (and de-escalating) incident and
information % availability point of view. This is complemented by the need to
support the dynamic % creation of workflows in a multi-stakeholder (and no
longer single stakeholder) setting.  % In addition, security is an important
aspects that drives a separate set of requirements % which impact data
availability and processing availability that must comply with a % dynamic
network centric processing model. That model must support secure access at % the
data-item level and without restrictions at the network or compute platform
level. A % mixture of the Cloud and Grid approaches may offer a solution here.

% Use of dynamic data

% The use of dynamic data has two origins: the first is the real world situation
that may % begin with a small scale event such as a minor gas leak. If such an
event is undetected % at first, it may escalate and trigger more and more
disturbances that add data about % different other anomalies to the data
streams. The second source of dynamic behavior % is that with the escalation of
the event, more and more special repression an % management process come into
play, each requesting and adding their own sort of % processing to the mix.

% How does the application get the data?

% In a typical application a number of special data and sensor interface
processes will be % part of the development. Whether packaged as software agents
or some other form of % software process, they will normally be distributed over
various platforms and fixed in % place. There task is to make the data available
via a distributed shared data space % (used here as a logical concept that can
be implemented in different ways) that % abstracts from physical storage. The
information processing algorithms are packaged % as autonomous processing
entities such as software agents, that are managed by a % multi agent system
infrastructure (middleware) and humans (via GUIs).

% QoS aspects

% Quality of service aspects that are important for real world and real time
applications are % the dynamics of scheduling from an application perspective,
multi independent level of % security support, processing service discovery
support, autonomous reconfiguration % support and
logging-for-traceability. Because of the nature of the application aspects %
like sustained communication bandwidth, processor availability,
checkpoint-restart etc.  % are not important. Most of the dynamic data that
these applications use has a short % validity period, i.e., it must be processed
now or we process the next update coming in.  % The data is uncertain, i.e., it
may have dynamically changing S/N ratio and although % expected at some base
frequency, may not be available for some period of time.

\subsection{MODIS Data Processing \label{modisKeith}}

%\katznote{from Keith Jackson}
%
%\katznote{summary below assigned to Yogesh}
%\subsubsection{new summary}

%\katznote{Yogesh, some parts are repetitive - can you fix them?}

% 1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

This {\em traditional application} allows the coordinated usage of
MODIS~\cite{modis} satellite imagery, and its integration with ground based
sensors and models for environmental applications.  It consists of data storage
and processing stages.

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}

The application~\cite{modisAzure,modis:ipdps:2010} is used to handle the processes of downloading
data from different sites and integrating them. The integration consists of
reprojecting the data from one map projection to another, spatial and temporal
resampling, and gap filling. The application itself is a loosely-coupled data
pipeline that coordinates these processes.

In addition, scientists can submit Matlab scripts for further custom processing
on subsets of the data that has been integrated. This includes some
summarization processing and plotting routines.  The only time constraints are
on the debugging cycle for debugging the scientist-supplied reduction
codes. These are purely driven by the users' patience.

The application has been implemented in both the Windows Azure cloud
environment~\cite{win_azure} and a standard HPC environment.

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below}

There is approximately 25~TB of data for one global year.  The main data is the
various MODIS satellite data products, of which 11 are used by this
application. These are in approximately 3,000,000 files for one global year. A
smaller set of sensor data from the fluxnet data network is used as input to the
evapotransporation calculation, along with a set of 20-30 files containing
various metatdata like climate regions used by the calculation.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

%\katznote{from answer to question 3 below}

 The data is collected from a wide variety of FTP servers managed by different
groups and brought into the cloud/HPC environment. Custom software tooling
manages and validates this download process. Reliable cloud message queues are
used to manage the coordination between the different components in the
multi-stage pipeline. Simple NoSQL tabular storage is used for logging and fault
recovery. Data is stored in cloud blobs and NoSQL table storage. Application
scripts on the user's desktop can be uploaded into the cloud as post-processing
routines and analyses on the integrated data.

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

 The data is currently static and the application acquires the data from a set
of FTP servers on a schedule or trigger. But as the pipeline is expanded to
handle other data, the usage of real-time streaming sensor data is also
expected. There is infrastructure dynamism through the use of cloud virtual
machines, acquired and released on-demand, to execute tasks in the pipeline and
to execute the user's application scripts. While the pipeline structure itself
is static, there is a limited form of application dynamism through the use of
user specified scripts pushed into the cloud for execution over the integrated
data.

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

The data is in one of two map projections, and multiple temporal and spatial
resolutions. The fluxnet sensor data that is used required significant data
integration, but was handled in a separate project.

The quality constraints are driven by the needs of the scientific reduction
application. Significant effort is required to ensure that the data handled to
the scientific application meets these requirements.
% \subsubsection{old material follows}

% 1. What is the purpose of the application?

% The purpose is to allow the coordinated usage of MODIS satellite imagery, and
to allow its integration with ground based sensors, models, etc.

% 2. How is the application used to do this?

% The application is used to handle the processes of downloading data and data
integration. The data integration consists of the ability to reproject the data
from one map projection to another, spatial and temporal resampling, and gap
filling. The application consists of a loosely-coupled data pipeline that
coordinates this process.

% 3. What infrastructure is used? (including compute, data, network,
instruments, etc.)

% The application has been implemented in both the Windows Azure cloud
environment and a standard HPC environment. The data is collected from a wide
variety of ftp servers managed by different groups. Significant custom
infrastructure is required to manage and validate this download
process. Reliable message queues are used to manage the coordination between the
different components in the pipeline, and simple NoSql tabular storage is used
for logging and fault recovery.

% 4. What dynamic data is used in the application?

% The data is currently static, but as we expand the pipeline to handle other
data we expect to see the usage of real-time streaming sensor data.

% (a) What are the types of data

% The main data is the various MODIS data products, of which we use 11. These
are in approximately 3,000,000 files for one global year. A smaller set of
sensor data from the fluxnet data network is used as input to the
evapotransporation calculation, along with a set of 20-30 files containing
various metatdata like climate regions used by the calculation.

% (b) What is the size of the data set(s)?

% There is approximately 25~TB of data for one global year.

% 5. How does the application get the data?

% The application dynamically acquires the data from a set of ftp servers.

% 6. What are the time (or quality) constraints on the application?

% The only time constraints are on the debugging cycle for debugging the
scientist-supplied reduction codes. These are purely driven by the users'
patience. The quality constraints are driven by the needs of the scientific
reduction application. Significant effort is required to ensure that the data
handled to the scientific application meets these requirements.

% 7. How much diverse data integration is involved?

% The data is in one of two map projections, and multiple temporal and spatial
resolutions. The fluxnet sensor data that is used required significant data
integration, but was handled in a separate project.

% 8. How diverse is the data?

% 9. Please feel free to also talk about the current state of the application,
if it exists today, and any specific gaps that you know need to be overcome.

% \katznote{from talking to Keith: There is a pipeline of initial standardized
processing steps on the data.  Then, scientists can submit executables that do
further custom processing on subsets of the data, which likely include some
summarization processing (building graphs).}

%\subsection{Floating Sensors \label{floaterKeith}}
%
%%\katznote{from Keith Jackson}
%%
%%\katznote{summary below assigned to Yogesh}
%
%\katznote{Yogesh, some stuff is missing - can you fix any of it?}
%\ysnote{Unfortunately, thats all the info I have from Keith. No response to a
reminder I sent him earlier.}  %\katznote{I will talk to Alex, the PI of this
project, on Feb 2.}
%
%%\subsubsection{new summary}
%
%% 1. {\bf What does the application do? What problem does it solve?}
%
%%\katznote{from answer to question 1}
%
%This {\em traditional application} constructs real-time hydrodynamics models,
and forecasts the behavior of a complicated water system: the Sacramento %- San
Joaquin River delta.  It begins after data generation and includes data
transformation \& decisions, storage, and processing.
%
%\subsubsection*{Application description} %%2. {\bf How does it do it?}
%
%%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}
%
%The application integrates both static and dynamic sensors streaming %real-time
data with a high-resolution hydrodynamics model running on the NERSC Magellan
cloud %testbed. The model uses sensor data to initialize and shape the run, and
the model output is in turn %used to drive the real-time positioning of the
dynamic sensors. The output of this process is a map %of the current and future
behavior of the modeled watershed.
%
%Custom built dynamic sensors are deployed in the watershed along with
traditional statically placed %sensors that provide data. These sensors
communicate over a variety of different network %infrastructures, e.g., GSM,
CDMA, 802.11, wired ethernet, etc.  and the application acquires the %data over
the internet from the sensor platforms.  The application processes the sensor
data in %real-time as it is streamed.
%
% \subsubsection*{Big data aspects} %%3. {\bf What are the big data aspects?}
%
%
%%\katznote{from answer to question 4b below}
%
%Not known.
%
% \subsubsection*{Distributed aspects} % %4. {\bf What are the distributed
aspects?}
%
%
%%\katznote{from answer to question 3 below}
%
%Not known.
%
% \subsubsection*{Dynamic aspects} %%5. {\bf What are the dynamic aspects?}
%
%
%%\katznote{from answer to question 4a below}
%
%The sensor are all streaming real-time data to the application. The position of
the dynamic sensors %themselves change over time.
%
%
% \subsubsection*{Other important issues} %%6. {\bf What else is important?}
%
%
%% \katznote{from answers to questions 7,8 below, also address programming
systems (how was the app written; what tools does it use)}
%
%The first full, end-to-end, field test of the system is scheduled for May 2011.
%

% \subsubsection{old material follows}

% 1. What is the purpose of the application?

% To model in real-time, and forecast, the behavior of a complicated water
system like the Sacramento - San Joaquin River delta.

% 2. How is the application used to do this?

% The application integrated both static (Eulerian) and dynamic (Lagrangian)
sensors streaming real-time data with a high-resolution hydrodynamics model
running at the NERSC supercomputing center. The model uses sensor data to
initialize and shape the run, and model output is used to drive the real-time
positioning of the Lagrangian sensors. The output of this process is a map of
the current and future behavior of the modeled watershed.

% 3. What infrastructure is used? (including compute, data, network,
instruments, etc.)

% Custom built Lagrangian sensors are deployed in the watershed along with
traditional statically placed sensors. These sensors communicated over a variety
of different network infrastructures, e.g., GSM, CDMA, 802.11, wired ethernet,
etc. The model is running on the NERSC Magellan cloud testbed.

% 4. What dynamic data is used in the application?

% The sensor streams are all real-time streaming data.

% (a) What are the types of data % (b) What is the size of the data set(s)?

% \katznote{Keith: I need to chat with Alex to get you these answers Dan. Please
remind me if I don't get back to you on this.}

% 5. How does the application get the data?

% The application is acquired over the internet from the various sensor
platforms.

% 6. What are the time (or quality) constraints on the application?

% The application is processing the sensor data in real-time as it is streamed.

% 7. How much diverse data integration is involved?

% 8. How diverse is the data?

% \katznote{Keith: I need to chat with Alex about the data diversity also.}

% 9. Please feel free to also talk about the current state of the application,
if it exists today, and any specific gaps that you know need to be overcome.

% The first full, end-to-end, field test of the system is scheduled for May.

\subsection{Distributed Network Intrusion Detection \label{detectionJon}}

%\katznote{from Jon Weissman}
%
%\katznote{summary below assigned to Jon}

%\subsubsection{new summary}

% 1. {\bf What does the application do? What problem does it solve?}

%\katznote{from answer to question 1}

This {\em traditional application} performs detection of distributed network
intrusion and distributed denial of service attacks~\cite{jon}.  Network TCP/UDP
traffic data is analyzed/mined at each site. Potential alerts are collected
centrally and further analysis is done to see if an attack may be underway by
correlating alerts, and events are then communicated back to local sites. The
following features are extracted: 1) basic features such as source/destination
IP address pair, port number, and protocols and 2) some additional features such
as the number of flows from the same source to specific destination IP addresses
for a predetermined time period. The extracted features are input for the MINDS
system. While well-known intrusions are detected by the known attack detection
module, the anomaly detection module is applied to the remaining network
connections. During the process of anomaly detection, a training set is
generated if not given as input. The anomaly detection module scores each
network connection to reflect the level of anomalousness of the network
connection compared to the normal network traffic. The highly scored network
connections are analyzed by the association pattern analysis module to produce
summary and characterization of the possible attacks. This summary and
characterization is used to create new signatures and models for unknown
emerging attacks.  MINDS contains various modules for collecting and analyzing
massive amounts of network traffic.  Typical analyses include behavioral anomaly
detection, summarization, scan detection and profiling.  The application
includes all four stages: data generation, transformation \& decisions, storage,
and processing.

\subsubsection*{Application description} %2. {\bf How does it do it?}

%\katznote{from answers to questions 2,3,5,6 below, also address
cyberinfrastructure (where does the app run) -- perhaps have a diagram here of
what happens in the app}

MINDS is deployed at each site as a grid service. The analysis application
continuously scans local logs, and reports interesting events to a centralized
analyzer, and alerts are reported back to local sites.

Each site uses a data capturing device such as network flow tools or network
monitoring tools like tcpdump to collect the network traffic data from routers.
This collected data is first filtered to remove unimportant network connections
due to the large data volume.

All communication in the system is accomplished using messages on top of HTTP.

 \subsubsection*{Big data aspects} %3. {\bf What are the big data aspects?}

%\katznote{from answer to question 4b below}

For local data, this is site dependent. It can be MB/hour/site.

 \subsubsection*{Distributed aspects} %4. {\bf What are the distributed
aspects?}

%\katznote{from answer to question 3 below}

The system is deployed across clusters located at a set of participating sites
(University of Minnesota, University of Florida, and University of Illinois,
Chicago) connected by the public Internet. In addition, the data generated is
also produced (distributed) at this set of sites.

 \subsubsection*{Dynamic aspects} %5. {\bf What are the dynamic aspects?}

%\katznote{from answer to question 4a below}

Network traffic data is highly dynamic and constantly streaming in to the
application. The data consists of TCP flow and UDP records, which are very
structured.  They contain IP addresses, ports, various headers, and
payloads. Interesting events are propagated across sites. These are very small,
a few KB at most.

%\ysnote{Some of the applications here like behavior anomaly detection seems to
be triggered on demand, and they consume additional data. This may be similar to
the LSST case where the application is dynamic/steered based on incoming data or
situations.}  %\katznote{Yogesh, what is your suggested action to do with this
comment?}

 \subsubsection*{Other important issues} %6. {\bf What else is important?}

%\katznote{from answers to questions 7,8 below, also address programming systems
(how was the app written; what tools does it use)}

 The application is built using Globus Grid Services for wrapping the MINDS
analysis applications and for providing storage services for events. At the time
the application was written, the project had to develop tools such as remote
launching of Grid services themselves. MINDS was also a separate software
package that was utilized. The project also built an authorization service for
controlling access to data products. Other noteworthy points: real-time is
ultimate goal of the project and false positives are more tolerable than false
negatives.  Finally, the project found it to be very difficult to obtain network
data at each site due to privacy concerns.

 IDS systems also use pattern detection tools such as Bro and Snort, which
perform simple string and regular expression pattern matching on IP packet
headers and body. These are monolithic tools that run standalone, without any
comprehensive programming support to automate updating the patterns or taking
action upon pattern detection.

%\subsubsection{old material follows}
%
%
%1. What is the purpose of the application?
%
%Distributed network intrusion detection and distributed denial of service
attacks.  Network TCP/UDP traffic data is analyzed/mined at each site. Potential
alerts are collected centrally and further analysis is done to see if an attack
may be under way by correlating alerts and events are then communicated back to
local sites. The following features are extracted:1) basic features such as
source/destination IP address pair, port number, and protocols and 2) some
additional features such as the number of flows from the same source to specific
destination IP addresses for a predetermined time period. The extracted features
are input for MINDS system. While well-known intrusions are detected by the
known attack detection module, the anomaly detection module is applied to the
remaining network connections. During the process of anomaly detection, a
training set is generated if not given as input. The anomaly detection module
scores each network connection to reflect the level of anomalousness of the
network connection compared to the normal network traffic. The highly scored
network connections are analyzed by the association pattern analysis module to
produce summary and characterization of the possible attacks. This summary and
characterization is used to create new signatures and models for unknown
emerging attacks. MINDS contains various modules for collecting and analyzing
massive amounts of network traffic. Typical analyses include behavioral anomaly
detection, summarization, scan detection and profiling.
%
%2. How is the application used to do this?
%
%MINDS is deployed at each site as a service. The analysis application
continuously scans local logs, and reports interesting events to a centralized
analyzer, and alerts are reported back to local sites.  These communication
steps use messages.
%
%3. What infrastructure is used? (including compute, data, network, instruments,
etc.)
%
%Small clusters at a set of participating sites (UMn, UFL, UIC). Public Internet
connectivity.
%
%4. What dynamic data is used in the application?
%
%Network traffic data is highly dynamic.
%
%(a) What are the types of data
%
%TCP flow and UDP records -- very structured: contains IP addresses, ports,
various headers, and payloads. Interesting events are propagated across
sites. These are very small, a few KB at most.
%
%(b) What is the size of the data set(s)?
%
%For local data, this is site dependent. Can be MBs/hour/site.
%
%5. How does the application get the data?
%
%The data capturing device such as network flow tools or network monitoring
tools like tcpdump are used to collect the network traffic data from routers.
The collected data is first filtered to remove unimportant network connections.
%
%6. What are the time (or quality) constraints on the application?
%
%Real-time is ultimate goal. False positives are more tolerable than false
negatives.
%
%Gaps: The project had to develop tools such as remote launching of Grid
services themselves. They also found it to be very difficult to obtain network
data at each site due to privacy concerns.
%

%\subsection{Wikianalytics -- Yike}
%
%Wikianalytics offers a new methodology and platform for 3D applications.
%
%Wikianalytics is based on the concept of collective intelligence and social
computing.  %In Wikianalytics, complex and large scale data is analyzed via a
group of analysts %collectively in an interactive and collaborative way. The
data mining life cycle is not a %linear process but a dynamic composition of
concurrent interactive processes with %the joint effort of participating
analysts.  Each analyst contributes data, prior %knowledge, expertise,
algorithms to build, validate, improve, integrate models %(knowledge).  Also,
modern cloud computing and social computing model provides %an ideal platform
for Wikianalytics.
%
%Why is Wikianalytics suitable for 3D:
%
%Data intensive: Data intensive is not just about the quantity but also the
complexity.  %Given the human genome example, the pervasive use of human genome
for %healthcare requires the analysis to be done with medical knowledge.
Analyzing %human genome is not just the issue of dealing with the huge amount of
sequence %data but also the need of deep analysis the biological and medical
implication of the %data. To meet this challenge, collective effort in analysis
is essential.
%
%Distributed: Distributed feature has three means : 1) data is distributed 2)
data owner/ %analysts are distributed and 2) data analysis environment is
distributed.  All three %aspects requires the analysis to be a compositional
process. That is, distributed data %can be analyzed in a distributed way and the
analytical products ( information or %knowledge) can then be combined
(integrated). Such a composition of distributed %process is a ``communicating
parallel composition'' in the sense that for time to time %each individual
component process will communicate with others to exchange the %information and
coordinate the actions. Such a distributed analytics process can be %realized
well by the Wiki collective intelligence principle.
%
%Dynamical: Dynamical nature of data intensive computing comes the dynamism in
%data availability in data comprehension and also in the inherent dynamism in
the %applications.  From data analysis point of view, the dynamism is reflected
by the fact %that an analytical process cannot be captured by a static workflow,
rather it has to be %dynamically constructed based on the partial information
derived during the analysis.  %Such an adaptive nature of analysis is described
in Bayesian statistics as the %posterior probability where knowledge (belief) is
dynamically updated based on the %new available and analyzed data.  Wikianalysis
is not based on the concept of %organizing data via workflow, rather its
collective mechanism reflects precisely the %dynamism needed in analyzing
complex and dynamic data where the analysis %process requires to be built
adaptively based on the up-to-date analytical results gain %from the joint
effort.

\subsection{Application Summaries}

Here we provide short summaries of the computational (as opposed to scientific)
aspects of each application in preparation for the analysis that follows in the
rest of this document, which is motivated by and centered around these set of
applications.

{\bf NGS Analytics (\S\ref{bioSilvia})}: Data is processed through a customized
pipeline of analysis tools, such as BWA, BFAST, Bowtie; processing of each data
element is independent of other elements, so processing is
decomposed/parallelized over the data elements.  All data is stored in files.
Later analysis may want to operate on files from multiple data stores
distributed around the world, e.g., to compare O(1000s) of TB-scale data sets
(metagenomics) with each other.  Dynamic elements include the choice of
executables in each instance of the pipeline, the design of the pipeline itself,
and the choice of computing infrastructure to provide the best resource usage
and turnaround time.  The solution to this problem is not clear.  There is a
need for flexible programming systems and associated runtime environments for
hardware-software co-optimization.

{\bf ATLAS/WLCG (\S\ref{WLCGSteve})}: There is a hierarchy of systems. Data are
centrally stored, and locally cached (and copied to where they likely will be
used), perhaps at various levels of the hierarchy. Processing is done by
applications that are independent of each other. Processing of one data file is
independent of processing of another file, but groups of processing results are
collected to obtain statistical outputs about the data.

{\bf LSST (\S\ref{astroAdam})}: Data is taken by a telescope. Quick analysis is
done at the telescope site for interesting (urgent) events (which may involve
comparing new data with previous data).  The system can get more data from other
observatories if needed, request other observatories to take more data, or call
a human. Data (in files) is then transferred to an archive site, which may be at
the observatory. At the archive site, the data are analyzed, reduced, and
classified, some of which may be farmed out to grid resources.  Detailed
analysis of new data vs. archived data is performed.  Reanalysis of all data is
done periodically.  Data are stored in files and databases.

{\bf SOA Astronomy (\S\ref{astroSOAAdam})}: Services are orchestrated through a
pipeline, including a data retrieval service that is used to share data across
VO sites, as well as analysis and image processing services.  Data (files) are
moved through the pipeline, and intermediate and final products can be stored in
AstroGrid storage service, such as mySpace.  Scratch space is provided for
storage of intermediate results.

{\bf Cosmic Microwave Background (\S\ref{astroJulian})}: This application builds
a sky map.  It reads data from detectors, and obtains data from simulations that
are performed on-the-fly, when requested by the map-making component (in a
master-worker sense), one of a set of components that gradually reduces the data
from the detected and simulated data to a small number of cosmological
parameters.

{\bf Sensor Network Application (\S\ref{sensorSimon})}: Data are collected from
mobile sensors that periodically transmit when they come within communication
distance of receivers. Fault tolerance (collecting data in a hostile
environment) is an important component. The data, which are very diverse,
because they can involved different resolutions, sampling frequencies and other
management characteristics., are brought to a central site. Stored data are
analyzed using statistical techniques, then visualized with tools such as Google
Earth.

{\bf Climate (\S\ref{climateDan})}: Climate simulations are run by climate
centers.  Data (outputs of climate simulations) are distributed in multiple
federated stores.  Current data analysis and visualization tools require the
user to bring the data they want to analyze to a local system, then run the
tools locally.  New tools will include the capability to add data transfer.
This will allow the set of tools to iterate: transfer data, do analysis, etc.

{\bf Interactive Exploration of Environmental Data (\S\ref{envJon})}:
Applications are driven from map-based graphical tools.  The inputs are
geographic areas and algorithms to run on the data in those areas. Data can come
through web services from stored data, sensors, or simulations (possibly
running).  The services can perform server-side processing to reduce the size of
data files to be transferred. Responses are provided in near-real-time.
Security also supported through a delegation-based model.

{\bf Power Grids (\S\ref{power})}: Diverse streams arrive at a central utility
private cloud over cellular and wireless networks at dynamic rates controlled by
the application. Communication is bi-directional, with pricing and control
signals sent to customers power meters.  A real-time event detection pipeline,
involving complex event processing and machine learning/data analysis algorithms
and using other stored information such as weather data, can trigger load
curtailment operations. Data mining is performed on current and historical data
for forecasting. Partial application execution on remote micro-grid sites is
possible.

{\bf Fusion (\S\ref{fusionScott})}: Multiple computational intensive physics
simulation codes run concurrently on a distributed set of parallel computers.
Data from some codes are streamed to other codes to link them into a single
simulation.  Data are transformed in flight into needed inputs.  Data are also
examined in situ to understand the overall state of the simulation.  Some data
are stored for later data mining and visualization.

{\bf Industrial Incident Notification and Response (\S\ref{Kees})}: Data are
regularly streamed from diverse sources, and sometimes manually entered into the
system. Initial disaster detection causes additional information sources to be
requested from that region and one or more applications to be composed based on
available data. Some applications run on remote sites (where the data is stored)
for data privacy. Escalation can cause more humans in the loop and additional
operations.

{\bf MODIS Data Processing (\S\ref{modisKeith})}: Data are brought into system
from various FTP servers.  A pipeline of initial standardized processing steps
on data is done on clouds or HPC resources.  Scientists can then submit
executables and Matlab scripts that do further custom processing on subsets of
the data, which likely include some summarization processing (building graphs).
The application has been run on a cloud and in an HPC environment.

%Floating Sensors (\S\ref{floaterKeith}): Data from static and dynamic %sensors
are brought to a central location for processing by models. The results %of
modeling and forecasting can cause the sensor locations to change.

{\bf Distributed Network Intrusion Detection (\S\ref{detectionJon})}: Data are
analyzed or mined by a service running at a number of local sites. The results
are sent (via HTTP messages) to a central site for analysis. Events are then
communicated (via HTTP messages) back to local sites.

\section{Understanding Distributed Dynamic Data}\label{sec:distdyndata}

% \jhanote{AL and I agree the structure should be: 3.1. (i) Define % Dynamic,
(ii) Formalism/types (iii) % Applications/tables/discussions. 3.2 (i) Define
distributed, (ii) % discuss two stages (a) structuring (static) (b) processing %
(dynamic), (iii) Formalism/table, (iv) Application discussion; 3.3 % discussion}

In the previous section we described thirteen applications and analyzed them
along an analytical space spanned by the three ``D''s of data-intensive,
distributed, and dynamic behavior.

\emph{Distribution} refers to the presence of application data in different
physical or logical locations. Distributed data may arise naturally in that the
data sources themselves are geographically separated, such as from remotely
located sensors.  In some cases, data must remain distributed for privacy and
policy reasons, or due to constraints in sharing.  In other cases, a full
dataset may not fit in one location either on disk or in memory and must be
decomposed to enable storage and/or computation. Distributed data may be driven
by a need for performance, scalability localization, reliability, or
availability considerations. For example, copies of data may be generated and
distributed to maintain data availability or load balancing.  In general, a
\emph{distributed application} is one that needs, or would benefit from the use
of, multiple resource~\cite{dpa_surveypaper}. Example benefits include increased
throughput, decreased time to solution, and increased reliability.  A
distributed application rarely consist of a single component that is
distributed; most often a distributed application is composed of smaller and
functionally self-contained, if not entirely independent, components that can
execute on different distributed resources.

\emph{Dynamism} refers to changes in the behavior and characteristics of an
entity % \jhanote{``of an entity'', does this % sound right?} \katznote{fine
with me} over time and/or space.  Over a long enough timescale, almost all
infrastructure, and especially distributed infrastructure, will exhibit some
form of dynamism, as a consequence of failure, varying load, etc. However, we
are interested in dynamism on timescales that influence application behavior,
scheduling decisions, etc. Some prominent examples are resource fluctuation,
data availability, and change in application execution characteristics.

In this section we will try to extract commonalities in the dynamic behavior of
applications studied in the previous section, as well as properties and types of
distribution seen in these applications.

% \jhanote{needs fixing} After discussing the different kinds of % dynamism, we
will apply this framework to the applications.  This will % permit an extraction
of common features and possibly \katznote{either % we can do this or not...} a
derivation of requirements for multiple % applications.

\subsection{Types of Distribution \label{sec:distTypes}} %Distributed
\emph{data} occurs e.\,g.\

There are many scenarios under which distribution of data is either necessary or
desired. In addition to the reasons behind distribution presented at the
beginning of this section (viz., large volumes of data) and reasons that
constrain or localize data (viz., privacy/policy issues), another common
situation under which data is distributed occurs when data comes from multiple
sources, for example multiple sensor streams or data centers, but needs to be
processed collectively.

% and (ii) where a data % set is too large for a single storage module and thus
must be stored % in a distributed % way.  % \alnote{application
stages?}\jhanote{Andre, please clarify?}

%{\it Application structure and distribution of data:}

% One or multiple data sources (M) are expanded to one or more % locations (N),
M$ <$ N, for processing.  Scattering is commonly used % to facilitate
distributed processing of data.

There are many ways to address distributed data.  These include but are not
limited to: % Ways to manage \katznote{?} the % distribution include but are not
limited to:
\begin{itemize}
\item \textbf{Replication:} Data is duplicated on multiple nodes. When fully
replicated, each execution unit has access to the same data. Replication can be
partial, i.e., data can be replicated to a subset of resources. The degree of
replication quantifies the average of the number of copies of each data
item. Replication may be non-uniform, e.g., if only a subset of data (perhaps
``hot'' data) is replicated. Caching is a lightweight form of replication.
\item \textbf{Partitioning:} Data is divided and distributed across multiple
nodes. Different forms of partitioning exist.  For example, data can be
thematically partitioned with respect to its type, such as sensor data
vs. simulated data, or data can be partitioned according to a spatial or
temporal attributes. This can be an opaque (e.g., ID) and or a domain dependent
(e.g., spatial) attribute.
\item \textbf{Streaming}: A form of distributed data processing in which data
arrives from a source; often the stream is a steady, continuous flow, but often
it fluctuates both spatially and temporally. Also, typically, the stream of data
must be processed in near real-time. Streaming thus also has attributes of
dynamism, which is discussed in the next subsection.
\end{itemize}

Which of the above approaches is employed cannot be reduced to a simple
rule. Which option can or should be used often depends upon a plethora of
co-dependent factors, such as the data volumes, the {\it degree of distribution}
(which describes the number of data sources and sites used for storing and
processing the data), the availability of compute resources, or the desired
level of reliability. Data distribution can be difficult to manage for reasons
of performance: latencies and bandwidths may vary unpredictably.  Further,
issues associated with reliability may also need to considered. An example is
the occurrence of a partial failure, where some (but not all) of the required
data remains available, leads to applications having to decide whether to
continue with reduced data or exit without processing what remains available.

%\jhanote{I am not sure this is the best way to express co-dependence % of
reliability and performance}\alnote{refined}

Viewed from an application's perspective, there are at least two application
characteristics that influence the distribution of data. The first is the stage
and the second is the degree of coupling between components of a distributed
application. Distribution in the first stage is primarily determined by the
structure and layout of the data sources. For example, the placement of the data
generated in the climate application (\S\ref{climateDan}) is based on which site
was assigned to run the particular model that generates the data. But analysis
of the data from multiple models requires bringing the data together to a single
site.  Although it is difficult to formalize the difference due to the large
fluctuations between different applications, but in general for a given
application, there are differences between the distribution prior to the
application's processing stage and the distribution of data within the
processing stage, e.g., to aid data-parallel processing.

% \jhanote{can we % provide an example or some further application discussion
for the % previous point. it would help immensely}

The second application factor that determines distribution of data is the type
of separation and coupling of the components of the distributed
application. Different coupling types place different constraints on data
distribution, either due to latency, or due to data volumes that must be
coordinated. % Different types of coupling in % turn influences the distribution
of data.  \emph{Loosely-coupled} distribution involves coordination and
cooperation between largely independent application components, as might be
found in a workflow system. \emph{Tightly-coupled} applications involve greater
constraints between the components, either in latency tolerance, coordination
requirements or relative scheduling. \emph{Decoupled} distribution results in
the separation of application components elements of the same task, for example,
sharing a spatial data set between components that cooperate to perform a single
logical task~\cite{dpa_surveypaper,thebookthatchangedtheworld}.  Most of the
applications in \S\ref{sec:scenarios} are loosely coupled, with the CMB
(\S\ref{astroJulian}) and Fusion (\S\ref{fusionScott}) applications as examples
of tight coupling.

% \jhanote{again, can we relate the above general paragraph to % specific
application examples from the previous section} % \katznote{tried to come up
with an application example, but instead, % realized that I was very confused by
this paragraph.}\jhanote{Any % better? If you could be more specific that'll
help}

\begin{table}[h]
  \begin{scriptsize}
    \begin{center}
      \caption{Distributed Processing}\label{tab:app_distributed_processing}
      \begin{tabular}{|p{3.5cm}|p{10.2cm}|} \hline \textbf{Application}
&\textbf{Distributed Characteristics} \\
% \endhead \hline Next Generation Sequencing (NGS) Analytics & Data sources (NGS
machine storage) and compute not always co-located; multiple data sources in
case of metagenomics \\ \hline ATLAS (an example of use of the WLCG) &Source
data is generated by one central data source. Data needs to be distributed to
match distribution of computing resources. During processing data is copied from
closest replica to the resource where compute job is run. \\ \hline LSST
&Multiple different kinds of data sources (observatories, data storage);
multi-step workflow that processes and archives data, carried on distributed
compute resources.\\ \hline SOA Astronomy Applications &Data resides in multiple
data centers; services are placed closed to the data.\\ \hline Cosmic Microwave
Background &Replication of subset of data to compute resources for processing.\\
\hline Sensor Network Application & Data is moved from different data sources to
a central analysis system for processing\\ \hline Climate (ESGF) &A subset of
the data is commonly moved to compute (desktop or grid resource)\\ \hline
Interactive Exploration of Environmental Data &Data is brought to grid resource
or desktop \\ \hline PowerGrids &Data is brought to central private cloud for
analysis\\ \hline Fusion (ITER) &Simulations run wherever resources are
available; in-situ analysis of large output data; partitioning/replication of
data for data-parallel processing\\ \hline Industrial Incident Notification and
Response &Distributed data sources. Processing takes place on multiple
distributed resources. \\ \hline MODIS Data Processing & Data is copied
(replicated) from different sources for processing.  Processing can take place
in different locations, e.\,g.\ cloud and user desktop. \\ \hline Distributed
Network Intrusion Detection &Distributed data and compute resources.\\ \hline
     \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

%\subsection{Characterizing Distributing Dynamic Data \label{sec:appChar}}

% \katznote{without the table that was here, we are now defining % parameters
($\alpha$, $\beta$) that we don't really do anything % with.  We should think
about what part(s) of this subsection make % sense to keep.} \jhanote{the tables
have been moved (and reference % to them fixed). Admittedly we don't use
$\alpha$ or $\beta$ % again..}  \katznote{then why do we define them?}

\subsection{Types of Dynamism \label{sec:dynTypes}}

\begin{table}[h]
  \begin{scriptsize}
    \begin{center}
      \caption{Understanding Dynamic Characteristics. This table highlights
different types and scenarios for dynamic data in the application set
investigated %\katznote{some explanation %here might
help}\jhanote{OK?}\katznote{yes} }
          \label{tab:app_dyn_char}
      \begin{tabular}{|p{4cm}|p{9.5cm}|} \hline \textbf{Application}
&\textbf{Dynamic Characteristics} \\
% \endhead \hline Next Generation Sequencing (NGS) Analytics & Data Dynamism:
different data formats and sources need to be processed\\ \hline ATLAS (an
Example of Use of the WLCG) & Data Dynamism: new raw data is steadily
collected\\ \hline LSST & Infrastructural Dynamism: Decision making based on
observed data\\ \hline SOA Astronomy Applications & Application and Data
Dynamism: data is continuously updated, dynamic queries\\ \hline Cosmic
Microwave Background Mapping & Application Dynamism: dynamic utilization of
resources\\ \hline Sensor Network Application & Data Dynamism: rate and volumes
of data dynamic \\ \hline Climate (ESGF) & Data \&Infrastructural Dynamism: new
data is generated over time. Dynamic distributed infrastructure for storing and
processing of data %\jhanote{please check} \\ \hline Interactive Exploration of
Environmental Data & Infrastructural and Data Dynamism: Various data sources,
varying input rates and different queries.\\ \hline PowerGrids & Application,
Infrastructural and Data Dynamism: data itself changes, data rates may change
(throttling), data sources change (traffic, weather, social networking), data
queries respond to seen data (e.g.  accuracy adjustment) \\ \hline Fusion (ITER)
& Application and Infrastructural Dynamism: data streamed between simulations
(multiple distinct data flows) \\ \hline Industrial Incident Notification and
Response & Application and Data Dynamism: data itself and data rate changes,
dynamic data sources, processing changes with data\\ % \jhanote{Omer needs to
clarify if application %dynamism too?} \\ \hline MODIS Data Processing &
Application and Data Dynamism: data that is operated upon changes,
infrastructure config changes\\ \hline Distributed Network Intrusion Detection &
Infrastructural, Application and Data Dynamism: data highly dynamically
changing, data rate variable, application responds to data, on-the-flight
processing\\ \hline
     \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

% \jhanote{Modify this paragraph to reference table 1 and discuss the % kinds of
dynamism} There are often spatio-temporal variability in the values and
volumes~\cite{deRoos2011UnderstandingBigData} of data.  The applications in
\S\ref{sec:scenarios} have variability in the production/generation rate of
data, and in the configuration of data production/generation.  Data dynamism may
arise from extrinsic factors, for example because sensors begin to observe more
events in the same data stream. Contrariwise, data dynamism may occur because an
application component requested an increased sampling frequency.  Similarly
application dynamism may occur through managed replication of a
compute-intensive component or from the demands of a particularly complex
query. And infrastructure dynamism may arise because the infrastructure expands
to meet additional requirements, or contracts because of device failure.  A
structured analysis of the applications in \S\ref{sec:scenarios} suggests that
the sources of dynamism can be classified into three categories.
\begin{enumerate}

\item \emph{Data} dynamism arises when some property of the input data or its
delivery changes, for example in terms of arrival rate, provenance, burstiness,
or source.  There may be variability in the structure of the data, for example,
data schema, file formats, ontologies, etc.

  % \alnote{Not sure, but maybe changes in the structure of data would % better
fit into category 1 or 3 since the structure is typically % not related to the
infrastructure}\ysnote{I agree with Andre.}  % \jhanote{In order to % be
consistent with paragraph above, the third type of dynamism % should focus on
the infrastructure -- resources not on data any % more.}

\item % Dynamism in the Application, % Software/Services use of the data:
\emph{Application} dynamism involves changing the processes or components
applied to a data set, and might encapsulate changes in workflow, \textit{ad
hoc} queries, or parameterization.  There may be variability in the data queries
enacted by the application, or in the behavior and/or characteristics of the
application.  % \ysnote{this may affect % data access as well as generation of
derived data}

\item \emph{Infrastructure} dynamism occurs when the system wants to, or is
forced to, change its demands on its underlying platform, for example through
elastic computation or partial failure.  Changes in the structure and load of
the infrastructure can be associated with data.  Examples include variability in
the producer/provenance of the data, in the performance, and in the quality of
service of the infrastructure.

\end{enumerate}

Based upon this classification, in \tablename~\ref{tab:app_dyn_char}, we
describe the applications discussed in \S\ref{sec:scenarios}.  Many of these
dynamic aspects are related: an increase in the rate of data generation may lead
an application to increase its degree of parallelism to maintain throughput,
which may in turn trigger the allocation of new compute nodes. (In the other,
less attractive direction, decreasing computational capability might cause a
decrease in parallelism and a consequent need to drop data. Avoiding these cases
means treating dynamism equally in both directions.)

This should not be construed to imply that every data-intensive application has
all elements of data and compute dynamism and is always distributed, but is a
reflection that at extreme scales, dynamism and distribution become increasingly
important issues that are often correlated.  As alluded to, a high degree of
distribution often results in a high amount of infrastructure dynamism, %
\katznote{I don't really like this forward-looking link, or the one in % the
next paragraph}\jhanote{Not forward looking anymore} caused by, for example,
resource fluctuations and a higher failure probability.  Similarly, high levels
of dynamism, for example, in the data sources, data rates, and queries, usually
correlate with a high degree of distribution.  For example, the Interactive
Exploration of Environmental Data application (\S\ref{envJon}) manages various
sources of data, from archived data sets to real-time feeds.  Depending on the
spatiotemporal properties of data, computation can be carried out on a set of
distributed grid resources or the user's desktop, i.e., distributed processing
is used to efficiently handle dynamic data.

\subsection{Distributed Dynamic Data Observations}

%\jhanote{needs text to establish that this is the end of section 3} \katznote{I
think this should be a short subsection, as I've done here - can be removed...}

Reviewing \tablename{s}~\ref{tab:app_distributed_processing} and
\ref{tab:app_dyn_char}, which have summarized the distributed and dynamic
properties of the D3 applications, we observe:

%find that commonly re-occurring patterns %\jhanote{should we be calling the
following patterns? I don't think % they are patterns, but possibly
observations?} \katznote{yes, `observations'} in the data %distribution and
processing are:

\begin{itemize}
\item Data generation and consumption are commonly distributed from each other
if the volume permits.  The Fusion application is a counterexample, due to data
volumes and velocity.

\item Details of the storage system are often not known. Often data is stored in
a distributed way, e.g., on a distributed file system, which commonly fits into
the medium degree of distribution (the same data center).

%\item Mapping: The most common pattern is to copy data to an available %
compute resource. \katznote{I'm not sure about this bullet: 1. It's not a
sentence; % 2. I don't know what it's saying.}

\item Aggregation of data can either be the equivalent of gather or reduce
(using the language of collective communication). Gather is in many ways the
reverse process of distribution (called broadcast in the language of collective
communication). Similar to distribution, aggregation is a commonly occurring
pattern for data analysis.

\end{itemize}

%\subsubsection{Characterizing Dynamic Data \label{sec:dynDataChar}} %
\jhanote{AL: Next sentences OK to go?} Also, in lieu of quantitative %
estimates, we attempt classification of these ratios into categories % of
high/medium/low.  It is natural to expect that applications will % cluster
around certain {\it trends}.

% \jhanote{AL/SJ: The following should be deleted?} As a guide to how % the
metrics proposed may be useful, we characterize a couple of the % applications
investigated using $\alpha$ and $\beta$: % \katznote{actually, neither the
examples here nor the table mention % $\alpha$}

% \begin{itemize}
% \item NGS: $\beta_1$ is high, $\beta_2$ is high. Degree of % Distribution is
Medium. The number of distributed sites does not % scale with data.

% \item Atlas: $\beta_1$ is Medium, $\beta_2$ is Medium. Degree of %
Distribution is High. The number of distributed sites scales with % data.
% \end{itemize}

% \jhanote{Is LSST like NGS or Atlas or in its own class?}  % \jhanote{Is
PowerGrid more like NGS or Atlas or in its own class?}  % \note{metrics for
distribute data} \jhanote{No idea who put the % previous note there or why? Help
please?}

% \jhanote{not sure where this goes?} replicated $\rightarrow$ average %
replication factor of a data element/N ; coupled $\rightarrow$ M/N

% \jhanote{Not sure this is correct?}  % Degree of distribution: latency between
primary data source and data % processing element

% \alnote{Refine description, define view point} \note{Is the control of % the
sensors from the application a case of 5 or is it something % different?
Feedback to the data source (other areas, sources of % data, accuracy)?}
\note{most applications commonly operate in a % local setting, but can also
operate in a distributed settings.}  % \alnote{How to measure degree of
distribution?} \jhanote{these notes % should move}
% \begin{table}[h!]

% \begin{table}
% \begin{scriptsize}
% \begin{center}
% \caption{Dynamic and Distributed Characteristics of Example
Applications} \label{tab:app_characteristics}
% \begin{tabular}{|p{3cm}|p{5.0cm}|p{5.1cm}|} % %\begin{footnotesize}
% %% \begin{tabular}{|p{4cm}|p{5cm}|p{5cm}|}
% % \begin{longtable}[h!]{|p{3cm}|p{5.0cm}|p{5.1cm}|} % % \caption{Dynamic and
Distributed Characteristics of Example
Applications} \label{tab:app_characteristics} \\ % \hline % \textbf{Application}
&\textbf{Dynamic Properties} &\textbf{Distributed Properties} \\ % % \endhead %
\hline % Next Generation Sequencing (NGS) Analytics & Infrastructure dynamics
(2) % &degree-of-distribution: medium; $\beta_1$: high; $\beta_2$: high \\ %
\hline % ATLAS (an example of use of the WLCG) &New raw data is collected %
steadily (1) % &degree-of-distribution: high; $\beta_1$: medium; $\beta_2$:
medium \\ % \hline % LSST &DE decisions based on observed data (3)
&degree-of-distributed: % medium, multiple different kinds of data sources
(observatories, data % storage); $\beta_1$: ?; $\beta_2$: ? \\ % \hline % SOA
Astronomy Applications &data is continuously updated (1), thus % dynamic data
(pre-defined schedules or on-demand), queries change (3) %
&degree-of-distribution: high, multiple kinds of data sources\\ % \hline %
Cosmic Microwave Background &dynamic utilization of resources (2 or 3?)  %
&degree-of-distribution: low--medium\\ % \hline % Sensor Network Application
&data rate, volume dynamic (1) % & degree-of-distribution: high, multiple kinds
of data sources\\ % \hline % Climate (ESGF) &New data generated over time
(ESGF-wide - is this in % scope of the application?) (1)
&degree-of-distribution: high, multiple % kinds of data sources\\ % \hline %
Interactive Exploration of Environmental Data & data rates and % data
inputs/outputs change (1), queries change (3) % & degree-of-distribution: high
\\ % \hline % % Dynamic Environmental Data on the Web &data operated on changes
% % (realtime feeds) (2) % % &natural (sensor sources, simulations), (sub-set)
replication\\ % % \hline % PowerGrids &data itself changes, % data rates may
change (throttling) (1), % data sources change (traffic, weather, social %
networking) and infrastructure dynamics (2), response to data (e.g. accuracy
adjustment) (3) % &degree-of-distribution: high, large number of sensors;
$\beta_1$: medium; % $\beta_2$: medium \\ % \hline % Fusion (ITER) &data
streamed between simulations (multiple distinct data % flows) (1), on-the-fly
transformation (3) % &degree-of-distribution: medium\\ % \hline % Industrial
Incident Notification and Response &data itself and % data rate changes (1),
dynamic data sources (2), processing changes % based on data (3)
&degree-of-distribution: high\\ % \hline % MODIS Data Processing & dataset that
is operated upon changes (1), % infrastructure config changes (7) %
&degree-of-distribution: high, multiple data sources\\ % \hline % % Floating
Sensors &data itself (1), data rate (2), application causes % % data changes (5)
% % &degree-of-distribution: high, multiple kinds of data sources; % %
aggregation \\ % % \hline % Distributed Network Intrusion Detection &data %
highly dynamically changing, data rate variable (1), application % responds to
data, on-the-flight processing (3) % &degree-of-distribution: high\\ % \hline %
% % \end{table} % % \end{longtable} % %% \end{tabular} % %\end{footnotesize}
%
% \end{tabular}
% \end{center}
% \end{scriptsize}
% \end{table}

\section{Infrastructure}\label{sec:infrastructure}

%\alnote{relate infrastructures back to applications: in particular, how do
specific purpose built infrastructures as the LHCG or ESGF fit here?}

We define {\em infrastructure} as the hardware and software that is provided to
support an application, rather than being explicitly created or hosted by the
user themselves. In many cases the precise boundary between infrastructure and
application is blurred, e.g., a catalog is a component with functionality that
may be similar to that of similar components in other applications, but is often
implemented specifically by an application rather than being provided as part of
the infrastructure. This may also be the case when reconfiguring an
infrastructure service for an application, when it is more onerous than
reimplementing the service specifically for that application. We also introduce
the concept of an ``infrastructure'' application as a collection of
infrastructure components that are packaged so that they can be used by many
applications in defined ways. The difference between these and ``traditional''
applications is often delineated by who owns and administers the application,
with traditional applications ``owned'' entirely by the user, and infrastructure
applications owned by someone other than the user but provided for use by
others. Infrastructure can be specialist as opposed to generic, for instance the
purpose-built infrastructure in ESGF (\S\ref{climateDan}) that is provided to
support the climate community.

In this section, we focus on the {\em software} infrastructure that is provided
to support D3 applications, specifically components that deal with distributed
or dynamic data. The components that make up the infrastructure can be
categorized based on the function provided, namely (i) infrastructure that
supports {\it data management}, (ii) infrastructure that supports {\it data
  analysis}, and (iii) infrastructure that supports the \emph{coordination
  within applications}. Each of these components may be present at three
different levels of the infrastructure: programming frameworks, services and
tools (e.g., MapReduce, workflow engines); middleware and platform services
(e.g., databases, message queues); and system-level software capabilities (e.g.,
notifications, file system consistency).

\subsection{Infrastructure to Support Data Management}
\label{sec:infra_mgt}

% \katznote{we should think about how the items in this section map to the boxes
and arrows in Figure~\ref{fig:figures_application-stages}.  For data sources,
storage, access, and movement, it's fairly clear.  For discovery and
notification, it's less clear.}

{\em Data Sources} are used in the acquisition stage as inputs of data into the
application. Broadly speaking these sources include types such as files, servers,
sensors, instruments, simulations and models.  Additionally, databases are often
considered as an intermediate data source in types of system. The dynamic
characteristics are primarily due to availability. These are often accessed in
the Data Generation stage of the application
(\figurename~\ref{fig:figures_application-stages}).

{\em Data Storage} components are used to provide persistent access to data by
the application. They may be structured (e.g., databases, No-SQL tables) or
unstructured (e.g., collections of audio recordings). The dynamic
characteristics are primarily due to the level of persistence, e.g., temporary,
durable, permanent (archived or published), or replicated; and also in the
structure of the data stored, e.g., schema free tables as opposed to relational
tables. There may also be dynamic characteristics of storage due to rate of
growth of data, e.g., a field survey may bring in a lot of data during summer
but none during other times.  These are commonly used in the Data Storage stage
of the application (\figurename~\ref{fig:figures_application-stages}), but may
also be present at the boundaries between the stages for transient staging of
data.

{\em Data Access} components are gateways to Data Sources and Data Storage
components, and provide interfaces that often enforce security controls to
restrict access, provide dynamic information about the available data sources
and/or data analysis components, and may also perform logging of requests and
data movement (e.g., tracking provenance for auditing). A side effect of
auditing/logging may be the growth of provenance metadata simply due to access
transactions even if no new data is created. Data Access components usually
serve at the interface between the application stages, before and after data
movement.

{\em Data Movement} components are used for bulk or continuous movement of data
between parts of the application, or to and from the application. % (often
called staging) Data movement may be explicit and visible to the application
(e.g., staging), or implicit (e.g., automatic replication of files). Components
may operate on streams or blocks (files). Dynamic characteristics include
partitioning, scheduling and planning (e.g., edge caching of ``hot'' data, where
datasets that are being used frequently during some period are replicated close
to the application). Data movement can be considered to be real-time (i.e., as
it happens, e.g., sensor data), as real-time (e.g., a storage buffer has been
introduced into the transport system, but the data source is still real-time, as
with time-series data), or asynchronous (where the ordering of the data movement
is independent of the data itself).  Latency may be considered as an important
characteristic of the data movement infrastructure depending on the application
e.g. real-time processing of sensor data for disaster analysis. Other dynamic
aspects of data movement include data rates, their variability from data
sources, and the ``freshness'' of data. These infrastructure appear on the edges
between the application stages.

{\em Data Discovery} components assist the access to and use of data sources,
and include Catalogs, Information Services, and Metadata stores. The dynamic
characteristics pertain to the evolution or recalibration of information held in
these components. These components are orthogonal to the applications stages and
can be used by any of the stages to guide their selection or identification of
incoming data sources and outgoing data sinks.  %\ysnote{rate of growth of data
may be an issue.  %Catalogs that perform nightly ``crawling'' or data loading
(e.g., sky surveys) may not be suitable if the %latest data needs to be
accessible to applications. Catalog update will have to keep up with data
%update. However, batch processing applications can survive less frequent
updates.}

{\em Notification} components respond to changes in the application or data and
trigger other functionality. These include publish-subscribe message brokers,
and complex event processing. By nature, these components support the dynamic
aspects of an application.  These components again are orthogonal to the
applications stages and can be used for coordination of the stages and data
movement.

In each case, these components can be single or centralized, distributed,
partitioned or federated. These are also typically offered as middleware and
platform services, or system-level software capabilities.

%\nchnote{where do we put data publishing? Calibration? provenance?}\ysnote{They
can be incorporated in data discovery and data access}

\subsection{Infrastructure to Support Data Analysis}
\label{sec:infra_analysis}

% \alnote{Should we also include infrastructure for data exploration, data %
mining, machine learning, visualization?}  %\katznote{we should think about how
the items in this section map to the boxes and arrows in figure 1.  The first
three map to both the data transformation \& decisions and data processing
boxes.  I have no idea what should happen with workflow} % \alnote{should we
also talk about the computational aspects of the infrastructure?  % At some
point compute and data need to come together.}

Analysis components include functionality such as comparison, statistical
analysis, event analysis (i.e., analyzing information derived from data rather
than raw data) and visualization. These components leverage the computational
aspects of the infrastructure.  Based on the application stages outlined in
\figurename~\ref{fig:figures_application-stages}, we identify a number of
infrastructure services that can be associated with these stages.

{\em Conversion} components (associated with the Data Transformation and
Decisions stage) deal with the changing of data into a different form, including
file conversion, (re-)formatting, transformation and metadata
extraction. Heterogeneity of formats and representations is common in many
scientific disciplines and such components often provide essential
pre-processing of data prior to analysis. These are often referred to as adaptor
services (or ``shim'' services in the Taverna workflow system~\cite{Taverna}).

{\em Enrichment} components take data and attach additional detail to them,
which might include filtering, image processing, information integration, data
tagging, semantic annotation, data augmentation, or data fusion (e.g.,
OLAP). Many of these components may deal with distributed sources of data. It is
useful to note that these operations may not be fully automated and may involve
a human expert. These components would be associated with the Data
Transformation and Decisions stage.

%\ysnote{addresses Andre's commment} {\em Analytics and visualization}
components help extract domain knowledge and assist with exploration of the
data. These are often closely supervised by the end-user and may even be
interactive. The proliferation of ``big data'' has caused data mining libraries,
machine-learning toolboxes and visualization environments to become part of the
infrastructure, some of which are deployed for specific domains but others which
are more generic (e.g., Weka, GraphLab). These components are associated with
the Data Processing stage.

\subsection{Infrastructure to Support Coordination within Applications}
\label{sec:infra_coord}

Programming abstractions and orchestration frameworks cut across both data
management and analysis functions. While different in character from the other
infrastructure discussed in this section, they enable the coordination of the
different stages of the application, sometimes enforcing a particular constraint
to improve, e.g., scalability. These approaches can allow the composition of
sequential data processing pipelines and further offer the flexibility of
modeling more complex constructs.  Standardization of APIs and components within
these frameworks can allow interchangeability of components with similar
functionality, such as those described in \S\ref{sec:infra_mgt} and
\S\ref{sec:infra_analysis}.

{\em Orchestration} components handle the management and choreography of data
placement, code placement, communication, and messaging to allow complex
analyses to be performed on distributed data sources. They may also be dynamic
if they have the capability to adapt their behavior based on earlier parts of
the workflow or can be steered based on external notification received during
execution. In addition to supporting choreography, often making use of one or
more resource manager/scheduler components, an orchestration component may also
include specialist support for conversion and analysis components discussed
above.

{\em Bulk Data-parallel} frameworks use ``Bulk'' operators to express
computations over collections of data which are executed in a loosely-coupled,
data-parallel manner.  The {\em MapReduce} paradigm~\cite{MapReduce} and its
variants~\cite{twister,hmr} are the most prominent among these. MapReduce forces
programmers to model computation in two stages: a \emph{map} phase in which a
transformation is applied to every element of a collection independently of
every other; and a \emph{reduce} phase in which the transformed results are
aggregated through a binary function that is associative and commutative.  These
properties serve to guarantee that the MapReduce computation can be efficiently
parallelized, by performing the map phase concurrently and then reducing without
the reduction order affecting the final result.

{\em Stream processing} (sometimes called event processing) is an approach
modeled on signal processing, in which a stream of data is transformed
``on-the-fly'' to produce one or more further streams of values. As such it is a
programming abstraction suitable for continuous data feeds, or when results need
to be pushed to users of interactive applications. Since the data is not assumed
to be available \textit{en masse}, it can be used in situations where storing
the entire data set is undesirable, unnecessary, or impossible.

{\em Sensor networks} and other intelligent front-end data collection systems
are sometimes seen as an extension of stream processing and application
pipelines.  They often have the feature of being \emph{adaptive}, in that they
can vary their exact behavior according either to external control stimuli or
from observations made of the data they are themselves observing. Adaptation may
also be through users interactively changing the the pipeline, either by
changing parameters or by adding or removing tasks in the pipeline. This allows
decision logic to be initiated by the pipeline itself, or using an external
control signal that may come from users or an instrument. This form of
``computational steering'' is especially useful in systems being used to explore
a data set. In these cases, it is important to note that the infrastructure is
adaptive, rather than the scenario itself.

%Gomes et al.~\cite{gomes08} describe how a pipeline may be %dynamically
modified using structural and behavioral operators.  %This also provides an
end-to-end description of the entire %scientific process. This is valuable both
for documentation purposes, %but also in pursuing issues of component and data
re-use.

\subsection{Examples of Specific Cyberinfrastructure Usage}

We conclude this section by revisiting the application scenarios of
\S\ref{sec:scenarios} in terms of the infrastructure and coordination approaches
that they use currently, and discuss elements that could usefully be deployed.

The software infrastructure components we have classified above address
different characteristics of D3
applications. \tablename~\ref{tab:inf_app_mapping} summarizes the role of these
components in addressing dynamic and distributed characteristics of the D3
applications in \S\ref{sec:scenarios}. Each application makes use of different
components, though it can be seen that applications that utilize stream or event
processing %(see \ref{tab:progSystems} \katznote{this table name doesn't
resolve...}  have larger requirements on notification and orchestration
components.

\begin{table}[h]
  \begin{scriptsize}
    \begin{center}
      \caption{Infrastructure used to support dynamic and distributed properties
of application scenarios % \jhanote{Are all the terms % in parenthisis
describe/defined. I think not, e.g., % orchestration? Also, I removed analysis
as a % category.. needs fixing here too} \ysnote{fixed} }
      \label{tab:inf_app_mapping}
      \begin{tabular}{|p{6.7cm}|p{6.7cm}|}

\hline \multicolumn{2}{|l|}{{\textbf{Application Name}}} \\ \hline %\nchnote{For
example, workload decomposition and distribution must be determined dynamically
in order to optimally use resources, e.g., compute resource selection can be
based on data location, processing profile and/or network capacity. Workflows?
Resource selection?}  Infrastructure for Dynamic Properties & Infrastructure for
Distributed Properties \\ \hline
%\begin{footnotesize}
% \begin{longtable}[h!]{|p{6.7cm}|p{6.7cm}|}
% \caption{Infrastructure used to support dynamic and distributed properties of
application scenarios} \label{tab:inf_app_mapping} \\ % \hline %
\textbf{Infrastructure to support Dynamism} % &\textbf{Infrastructure to support
Distributedness} \\
% \endhead \hline \multicolumn{2}{|l|}{{\textbf{Next Generation Sequencing (NGS)
Analytics}}} \\ \hline %\nchnote{For example, workload decomposition and
distribution must be determined dynamically in order to optimally use resources,
e.g., compute resource selection can be based on data location, processing
profile and/or network capacity. Workflows? Resource selection?}  Dynamic
workload decomposition and distribution for resource optimization
\emph{(Orchestration)} & Distributing data from source (generation) to the
computing destination (analysis) \emph{(Data Discovery, Movement)} % No unifying
single requirement or single solution at present (or possible at all?)  %
Normally data local and CI doesn't easily support movement of data to compute %
\ysnote{We may want to remove the preceding lines since we're talking about
existing infrastructure, not missing ones at this point.}  % naturally
distributed: data source (NGS machine storage) and compute on % different
resources (typically not co-located); loosely coupled (many % machines,
different repositories) \\ \hline \multicolumn{2}{|l|}{{\textbf{ATLAS (an
example of use of the WLCG) Reconstruction app}}} \\ \hline Manage gradual
growth of data using SRM \emph{(Data Storage, Access)} % Management, subset
replication and partitioning to go inf components: data replication, (gLite),
data movement % specify SRM as standard and then they leave choice of
implementation to the organisations as part of consortium.  % New raw data is
collected % steadily (2) \note{timescale needs to worked out. What happens if
something is found?}  & Management, subset replication, data movement and
partitioning using gLite and SRM \emph{(Data Storage, Access, Movement,
Discovery)} % , data movement % natural, (sub-set) replicated, loosely coupled
\note{Is natural the % right word? Artificial % distribution (the source is
natural, but the intermediate points are % artificial), data not
partitioned}\ysnote{These details are not fully present in the actual
application description of its dynamism (Sec 2.2). Should we trim there and move
some content there? } \\ \hline \multicolumn{2}{|l|}{{\textbf{LSST}}} \\ \hline

Steering observations, Databases to manage data growth \emph{(Notification, Data
Storage)} % DE decisions based on observed data (2, 5) \ysnote{what do these
numbers mean? specific % characteristics?}  & Coordinate classification,
distributed databases, data movement to distributed observatories
\emph{(Notification, Data Access, Data Movement, Data Storage)} %Analysis, % ,
natural distribution of telescopes \\ \hline \multicolumn{2}{|l|}{{\textbf{SOA
Astronomy Applications}}} \\ \hline

Catalog contents updated, Query mix changes \emph{(Data Discovery)} % data is
continuously updated, thus dynamic data (pre-defined schedules or on-demand),
queries change (7)\note{timescale clarification!}  & Distributed VO Services for
data storage and access \emph{(Data Discovery, Data Access, Data Movement)} %
moving data sources, services and storage are distributed across institutions \\
\hline \multicolumn{2}{|l|}{{\textbf{Cosmic Microwave Background}}} \\ \hline
Dynamic resource selection %dynamic utilization of resources (4) & Compute on
distributed HPC centers \emph{(Data Movement)} % (sub-set) replication (during
application run) \note{What phases of the application do we consider? data
acquisition? app run?}\ysnote{Can any of the supercomputers be selected? If so,
the data is moved to distributed compute (or it is dynamic resource selection)}
\\ \hline \multicolumn{2}{|l|}{{\textbf{Sensor Network Application}}} \\ \hline
Notification of sensor data when in range, data rates and volumes \emph{(Data
Movement, Notification)} %data rate, volume dynamic (1-2) &Aggregating
distributed sensor data \emph{(Data Movement, Notification)} \\ \hline
\multicolumn{2}{|l|}{{\textbf{Climate (ESGF)}}} \\ \hline Manage data growth %
New data generated over time (ESGF-wide - is this in scope of the application?)
& Distributed, replicated catalogs using ESGF software stack (GridFTP,
wget/HTTP) \emph{(Data Access, Data Discovery, Data Movement, Data Conversion)}
\\ % multiple climate centers (natural), users can request data from ESGF
(sub-set replication), coupled % Can come in via website - cookie / OpenID %
Come in through command line script - grid cert \hline
\multicolumn{2}{|l|}{{\textbf{Interactive Exploration of Environmental Data}}}
\\ \hline On-demand data subsetting, caching and versioning, real-time
instrument data \emph{(Data Discovery, Data Movement, Data Sources)} % at moment
services are hosted on a single address. There may be farming for load balancing
in the future.  % data rates change (1), % data inputs/outputs change (3),
queries change (7) & Distributed Data Services, Access Restrictions \emph{(Data
Movement, Data Access)}

% data frequently located, but data sources themselves don't change
locations. Some replication for bigger community datasets. No replication
service - just another website!

% distribution of data has security constraints. Access rights which means you
can't just copy data. Data providers like to assert these rights at source.

% Increasingly user tokens being passed to data providers, but previously
application had tokens and trust required on part of data providers. Data
providers want to have user tokjens so they can tell usage, and whether its
researchers or commercial.

% Web server accessing data on your behalf. Not often directly accessing data -
going through an intermediary service. trying grid proxy certs and OAuth.

% If you had freely available data you could build a better system.

% As a response to query you get a subset of data % Virtualization and
aggregation of files - see single logical dataset rather than the thousands of
files its stored as.

% OpenDAP binary stream for data transport. logical OpenDAp data set may be
collection of files.

% If file changes (e.g. due to satellite reimaging) however difficult to tell
when data set has changed as versioning not well supported.

\\ \hline % Dynamic Environmental Data on the Web &data operated on changes %
(realtime feeds) (2) % &natural (sensor sources, simulations), (sub-set)
replication\\ % \hline \multicolumn{2}{|l|}{{\textbf{PowerGrids}}} \\ \hline

Stream processing over variable sources, rates and volumes, Event processing
over different patterns, Trigger analysis on-demand \emph{(Data Sources,
Notification, Enrichment, Orchestration)} %Analysis, % data changing at the
because application choosing more accurate sources % How is the selection made?
Is there a registry?  % There is a registry /catalog, it is distributed (sensors
catalogued by powergrid, buildings in campus database). Require distributed data
colletion to create a virtual single catalog.  % Using cloud computing to handle
scaling/elasticity for peaks. Not automated as yet but expected for the future.
& Dispersed event sources and sinks, Launch ensemble modeling on distributed
clusters and cloud \emph{(Data Sources, Notification, Orchestration)} %Analysis,

% There is a registry /catalog, it is distributed (sensors catalogued by
powergrid, buildings in campus database). Require distributed data colletion to
create a virtual single catalog.

% Output of computation itself is distributed - sent to various neighbourhoods /
locations.  % Some sent to human beings, sms, emails, many formats. Best effort
message delivery.

 \\

		\hline \multicolumn{2}{|l|}{{\textbf{Fusion (ITER)}}} \\ \hline
Stream and transform data between simulation sites on-demand \emph{(Data
Movement, Enrichment)} % (multiple distinct data % flows), on-the-fly %
transformation (8) & Run analysis on available, distributed resources
\emph{(Data Movement)} %Analysis % natural, coupled, transient sub-set
replicated \\ \hline \multicolumn{2}{|l|}{{\textbf{Industrial Incident
Notification and Response}}} \\ \hline Acquire different data sources and rates,
Processing determined by data \emph{(Data Sources, Data Access, Notification,
Orchestration)} %Analysis, % (1), % data itself changes (2), dynamic data
sources (3), processing changes % based on data %(5) & Distributed data sources,
Analysis by agents across institutions \emph{(Data Sources, Data Access)} % ,
Analysis % natural, coupled \\ \hline \multicolumn{2}{|l|}{{\textbf{MODIS Data
Processing}}} \\ \hline

% Satellite is already well known - static set of FTP sites.  -- % Pipeline is a
static structure - knows what it needs to do. Three different tasks looslely
coupled.  % Pipeline is setup for a community and once up and running will be
available to users for a while.

% How are users authorised to use the system? Portal, allows them to choose a
region. Basic authN?

		& Data from distributed FTP sources, Cloud queues for
coordination, Partitioned NoSQL data \emph{(Data movement, Data Storage,
Orchestration)} %Reliable message queues used as you don't want to have a black
spot in the sky - is it a real feature or a missing data. Try not to miss any
tiles!  \\ \hline %\multicolumn{3}{|l|}{{\textbf{Floating Sensors}}} \\ \hline %
&data rate (1), data itself (2), application causes % data changes (5) % &
spatial-temporal correlated, coupled \\ % \hline

\multicolumn{2}{|l|}{{\textbf{Distributed Network Intrusion Detection}}} \\
\hline Analyze dynamic network traffic rates \emph{(Data Movement)} % , Analysis
% data rate variable (1), data % highly dynamic changing (2), application
responds to data (5), % on-the-flight processing (8) & Run analysis at different
sites, Distribute generated data \emph{(Data Movement)} % Analysis, % natural,
data/compute tightly coupled \\ \hline
% \end{longtable}
%\end{footnotesize}

      \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

% in the process pose two questions:
% \begin{enumerate}
% \item What programming systems and techniques are \emph{currently} % used in
implementing the application scenario?
% \item What elements identified in this section \emph{could} usefully % be
deployed or further explored in each application scenario?
% \end{enumerate}

% \jhanote{I think the second question is not addressed in the remainder % of
this subsection? Yogesh/Neil?}  % \ysnote{removed the "questions" and rephrased
it as a discussion}

\tablename~\ref{table:progSystems} summarizes the coordination approaches
currently dominant in each application case study. It is worth noting that, in
common with most long-lived systems, many of the cases have accreted substantial
code bases over time and so show little in the way of systematic
architecture. (One reason for the dominance of workflow pipelines may be that
such architectures are quite suitable for adding extra processing steps to
existing computations.)

\begin{table}
  \begin{scriptsize}
    \begin{center}
      \caption{Coordination approaches used by the
applications\label{table:progSystems}}
      \begin{tabular}{|p{6.3cm}|p{7.2cm}|} \hline Application & Current
approaches \\ \hline \hline NGS Analytics & pipelines that ship with
sequencers~\cite{illumina}; workflows~\cite{Taverna} \\ \hline ATLAS & custom
algorithms; job submission tied to WLCG \\ \hline LSST & custom layer; scripted
workflows; queries \\ \hline SOA Astronomy & web services \\ \hline Cosmic
Microwave Background & \textit{ad hoc} \\ \hline Sensor Network Application &
focused on data collection \\ \hline Climate & \textit{ad hoc} \\ \hline
Interactive Exploration of Environmental Data & visualization tools \\ \hline
Power Grids & stream processing; data mining \\ \hline Fusion (ITER) & scripted
pipeline \\ \hline Industrial Incident Notification and Response & agent-based
\\ \hline MODIS Data Processing & scripted pipeline \\ % \hline Floating Sensors
& \textit{ad hoc} \\ \hline Distributed Network Intrusion Detection & monolithic
packages \\ \hline
      \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

The applications in \tablename~\ref{table:progSystems}, which make use of
scripted or \textit{ad hoc} pipelines, could in most cases be easily replaced
with a structured workflow engine. Doing so would bring immediate benefits in
terms of reproducibility, but also in terms of standardization of components and
interfaces.

%Standardization is not simply a software issue; it has %scientific significance
in the sense that data produced using %well-documented, well-used, and
well-debugged suites of components is %often more trusted -- and trustworthy --
than data produced using %locally-produced components that, while they may be
more innovative %and targeted, also run the risk of undetected assumptions and
errors.  %\jhanote{i think standardization of whatever data/component etc as an
% issue is independent of structured workflow engine. in fact i would % argue
that the causative arrow flows in the opposite direction: % structured workflow
engines do not help standardization but % standardizations does help structured
workflow engines..}  %\jhanote{IMO also still need to make it relevant to the
sub-section % aims?}

%\jhanote{NCH, we don't talk about standardization before this do we? so why %
further? Are we still trying to understand table IV? if yes, how % does this
relate?}  A further implication of standardization is that component reuse is
greatly improved. This is perhaps best illustrated by observing the increased
ability to automate the processing step over repeated runs.  The use of standard
component-based software techniques such as web services, coupled with web
technologies for metadata representation, assist in reuse, re-purposing and
integration, for instance into different workflow systems.

The {\em MapReduce} paradigm~\cite{MapReduce} and its
variants~\cite{twister,hmr} are the most prominent among the bulk data-parallel
frameworks which use bulk operators to express computations over collections of
data in an loosely coupled manner.  Applications such as Next Generation
Sequencing (NGS) (\S\ref{bioSilvia}) and Power Grids (\S\ref{power}) use
MapReduce for their
analytics~\cite{cloudburst,genometoolkit,Yin:mapreduce:2012}.

There are many examples of stream~\cite{chandrasekaran:telegraphcq2003,
abadi:borealis2005, liew:streamgraph2010} and event~\cite{ESPER, siddhi,
MicrosoftInsight} processing frameworks, and recent works have combined
streaming with more traditional file-based
workflows~\cite{ZinnCluster,Herath:CCGrid:2010}.  Applications such as the
Marine Sensor Network (\S\ref{sensorSimon}), Distributed Network Intrusion
Detection (\S\ref{detectionJon}), Power Grids (\S\ref{power}), and Industrial
Incident Notification and Response (\S\ref{Kees}) use stream and event
processing abstractions.

The view of scientific applications as pipelines hides significant differences
between (and within) domains, since the different stages of the pipeline receive
different emphases in different applications. Some applications must collect and
store all the incoming data, with only minor pre-processing to clean the data
ahead of the main processing. For others, a major constraint is to reduce the
incoming data stream as quickly as possible, perhaps discarding a large fraction
of the data before any processing occurs. In spite of these differences, we
believe the use of the pipeline abstraction is still useful.  Taking some of the
examples from \S\ref{sec:scenarios}, medical imaging and astrophysics often fall
into the first category, while particle physics (especially in the case of the
Large Hadron Collider) often falls dramatically into the second, to the extent
that principled data reduction forms one of the major scientific challenges for
the project.

There are also significant variations in where ``the pipeline'' begins and
ends. This represents the difficulty in differentiating between ``traditional''
applications and ``infrastructure'' applications. Many systems are driven by
data sets collected and shared widely, against which several different
applications may be applied.  Here the infrastructure application is the
collection of components used to allow the data sets to be accessed and analyzed
by the different applications.  Climate science is a good example of this
approach. In this case the data is ``pre-collected'' in the sense of existing
outside most of the applications, although its collection may have been a
complex task. The raw data required for gene sequencing (\S\ref{bioSilvia}), for
example, while being collected using a dedicated and very sophisticated
computerized processing system, would generally be regarded as ``given'' and not
part of the processing pipeline, since the pipeline would not affect the
collection task. Another way to look at this is that data is collected
\emph{independently} of its further processing.

Alternatively the application may be driven directly from a live data stream,
with processing happening on the fly. The front-end data collection subsystem is
becoming increasingly flexible and capable, with the introduction of sensor
networks to replace more passive approaches, and it may often make sense to
include the sensor network into the application workflow pipeline, such as for
Power Grids and Industrial Incident Response. All these choices have
implications for design and provisioning. For instance, the choice of which sort
of pipeline a particular application uses may well be fixed, or will change only
slowly as the experimental environment evolves. For example, we see some systems
that add simulation alongside experimental data collection to act as a
verification and validation step.

Revisiting the overall issue of infrastructure, which we defined as the hardware
and software needed to run the applications, we note that the specific hardware
isn't a primary issue.  The applications we've discussed all use a mix of the
systems that exist today, which have aspects of web, grid, and cloud, and can be
considered a hybrid of the three.  The key infrastructure components are really
the software and services on top of the hardware.  In this section, we've
discussed infrastructure to support data management and data analysis, as well
as the tools and systems used to program applications, and the specific
components used for the application in \S\ref{sec:scenarios}.  However, there is
not much commonality in the software and services that the applications use, so
the software and programming infrastructure we've described here is much more
broad than what is used by the applications we've studied.

%``This paper presents the design and the implementation of an %architecture for
the analysis of data streams in distributed %environments. In particular, data
stream analysis has been carried out %for the computation of items and itemsets
that exceed a frequency %threshold. The mining approach is hybrid, that is,
frequent items are %calculated with a single pass, using a sketch algorithm,
while %frequent itemsets are calculated offline by a further analysis. The
%architecture combines parallel and distributed processing to keep the %pace
with the rate of distributed data streams. In order to keep %computation close
to data, miners are distributed among the domains %where data streams are
generated. The paper also reports the %experiment results obtained with a
prototype of the architecture, %tested on a Grid composed of two domains
handling two different data %streams. Results confirm the scalability of the
approach and show that %it is possible to keep the pace of data production and
reduce the %communication overhead.''
%
%``.. In some cases streams of data must be analyzed in real time to %provide
information about trends, outlier values or regularities that %must be signaled
as soon as possible.  The need for online computation %is a notable challenge
with respect to classical data mining %algorithms [1], [2].  Important
application fields for stream mining %are as diverse as financial applications,
network monitoring, security %problems, telecommunication networks, Web
applications, sensor %networks, analysis of atmospheric data, etc.  A further
difficulty %originates when streams are distributed, and mining models must be
%derived not only for the data of a single stream, but for the %integration of
multiple and heterogeneous data streams [3]. This %scenario can occur in all the
application domains mentioned %before. For example, in a Content Distribution
Network, user requests %delivered to a Web system can be forwarded to any of
several servers %located in different and possibly distant places, in order to
serve %requests more efficiently and balance the load.''
%
%``..the architecture combines the parallel and distributed %paradigms, the
first to keep the pace with the rate of a %single data stream, by using multiple
miners (processors %or cores), the second to cope with the distributed %nature
of data streams. Miners are distributed among %the domains where data streams
are generated, in order %to keep computation close to data.''
%
%
%
%FlumeJava~\cite{flumeJava} is a distributed, reliable, and available %service
for efficiently collecting, aggregating, and moving large %amounts of log
data. It has a simple and flexible architecture based %on streaming data
flows. It is robust and fault tolerant with tunable %reliability mechanisms and
many failover and recovery mechanisms. The %system is centrally managed and
allows for intelligent dynamic %management. It uses a simple extensible data
model that allows for %online analytic application.

% Programming % Factored out by sd for various not-very-good reasons....
%\input{3dpas_programming}

\section{Discussion and Conclusions} \label{sec:conclusions}

% \jhanote{Andre L suggests a schematic to capture the main ideas from %
applications and the data properties. I think it is a good % idea. Omer: do you
want to think about this as you review/refine % this section}

%\jhanote{main message: (i) to application users/developers, you're not % alone,
(ii) infrastructure scientists: think of abstractions, and % how to build tools,
(iii) domain/application scientists message.}

%\jhanote{summarize contribution at front before 5.1 begins}

This section brings together some key observations from the applications and
infrastructure analysis carried out in this paper.  %as take away messages.
Section~\ref{sec:observations} summarize the dynamic and distributed
characteristics of data associated with the applications we've studied in this
work, attempting to infer more general characteristics that %can be applied may
apply to other similar applications.  In \S\ref{sec:discussion}, we suggest two
metrics that may be associated with such applications, primarily as a basis to
compare how compute and data processing (in terms of their number and
distribution) may be used to find common features between them. Such feature
analysis is useful, we believe, for two main reasons: (i) to identify
similarities between the applications covered in \S\ref{sec:scenarios} with
others that are in the design phase or under development, so that common
requirements can be identified; and (ii) to identify gaps in current software
tools and programming systems that such applications use, thereby leading to the
development of new tools and systems. As our application and infrastructure
coverage is, by necessity, limited, identifying common features also enables the
lessons learned to be generalized and applied more widely. We subsequently
discuss a set of architectures for D3 science, and pose some questions we
believe are pertinent as the community begins considering the next-generation
architectures for D3 applications. In \S\ref{sec:conclusion} we conclude by
providing a number of observations to inform applications and
cyberinfrastructure developers, based on common trends found in our study.

An important motivation for this work is to capture the state of the art in the
design and development of distributed dynamic applications and the associated
infrastructure currently used for their execution. We have aimed to try to
discern common approaches and challenges, and to identify any obvious gaps (or
opportunities). Using a common vocabulary and terminology to describe otherwise
distinct and unrelated applications allows potential application
users/developers to benefit from the insight that a common analytical framework
provides. For example, seeing how applications with similar patterns (e.g.,
pipelines) have addressed issues of dynamism and distribution enables common
solutions and facilitates the emergence of best practices even in the absence of
formal methods.  Similarly, our analysis sends a useful message to domain and
application scientists that the challenges and barriers they face are not unique
even if current solutions require a level of customization.

%\subsection{Common Characteristics}\label{sec:vectors}

% Given the application descriptions in \S\ref{sec:scenarios} and % subsequent
discussion about infrastructure characteristics in % \S\ref{sec:infrastructure},
we identify features that may be used to % compare these applications. Such
feature analysis is useful, we % believe, for two main reasons: (i) to identify
similarities between % the applications covered in \S\ref{sec:scenarios} with
others that are % in the design phase or under development, to enable common %
requirements to be identified; (ii) to identify gaps in current % software tools
and programming systems that such applications make use % of, thereby leading to
the development of new tools and systems. As % our application and
infrastructure coverage, by necessity, is limited, % identifying common features
also enables the lessons learned to be % generalized and applied more
widely. One important result is that we % can identify when particular
applications suit particular data-centric % programming models, e.g., stream
processing, MapReduce and Peer-2-Peer % networks.

%\item mechanism for supporting in-situ analysis at point of data % capture
(includes decision making to decide where analysis is % done)\note{move to
vector section.}
%\end{itemize}

%The first of the objectives above makes use of the four axes that %characterize
applications in the DPA survey paper, the intention of %which was to ascertain
how a number of real world applications made %use of distributed computing
infrastructure. Although we consider a %different set of applications here, the
aims are somewhat similar: to %both assess how existing real world applications
utilize distributed %computing infrastructure and also how they adapt to various
dynamic %characteristics that exist in their operating environment.

%Table~\ref{tab:characteristics} compares the applications across the four axes
identified at the beginning of this section.

\subsection{Observations}
\label{sec:observations}

In this subsection we summarize the source and nature of dynamic data that we
observe in the application surveyed in \S\ref{sec:scenarios}, and the types of
data distribution that we observe in such applications.

\subsubsection{Understanding Dynamic Data } %\alnote{in section 3 the three
classes for dynamism were different}

The applications in \S\ref{sec:scenarios} show that the dynamic characteristics
of many applications arise in two ways, due to data generation and placement or
due to the coordination mechanisms of the application.  The first of these
refers to data dynamism and the second to application dynamism (outlined in
\S\ref{sec:dynTypes}).

%\jhanote{Omer: this is inconsistent -- at least in terminology with % the 3
types of dynamism identified in Section 3: data, application % and
infrastructure} \rananote{Shantenu: this has been updated to % align with the
terminology in section 3. This is primarily referring % to the first two kinds
of dynamism mentioned there.}

\emph{Data generation, placement and consumption}: In many applications, data is
dynamically generated in response to a user request or when a particular
phenomenon of interest has been observed. Often this data is large in size and
must be placed close to the user undertaking the analysis. In some instances, it
may not be known where (and when) an analysis request may be made, and it is
therefore necessary to identify how replicas of the data could be produced and
placed across multiple possible locations.  In general, dynamic data arises due
to changing data properties, changing data volume (generation, consumption, or
processing), and changing data location or distribution.

% The coordination process of managing interaction between % components
undertaking data analysis, or the interaction of the % components and the
execution environment in which they are % deployed.

% Coordination mechanisms to support dynamic behavior remains an % important
feature across D3 Science applications.

\emph{Application Dynamism: } In many scientific applications the workflow
structure is well defined, but either the components used within this structure
or the location of the data that feeds the workflow may change dynamically.
Data coordination and planning (identifying how data is managed from ingest
through analysis to output) within such applications can range from statically
defined interactions between execution units and data sources to a fully
adaptable interaction using multi-agent systems.  As outlined in the Industrial
Incident Notification and Response application in (\S\ref{Kees}), the workflow
structure may also change based on a user request or the event being
observed. Data relating to the coordination of a components that make up the
application can therefore also be dynamic and change.  Exposing the data
planning used within an application makes it possible to modify or adapt it over
time.

%In the Industrial Incident Notification and Response
application\S\ref{sec:scenarios}.

%\jhanote{AL and I think this needs attention still. Either discuss % under data
planning or subsume dynamic coordination under % application dynamism}
\rananote{Agree with Dan's paragraph below as % a replacement -- and agree that
it should be subsumed under % application dynamism.}

%\katznote{SJ to move this paragraph to under application dynamism, and % merge
as needed.}\jhanote{done. NCH Can you check I have not mangled % the definition
of data planning please?}

%Exposing the data coordination/planning %used within an application (composed
of a number of different %components) to can be beneficial.  By exposing the
coordination %mechanism used, it becomes possible to modify or adapt it over
time.

%\jhanote{Omer: this is talking about changes in the infrastructure as % being
the cause and coordination changes as being the effect. The % last sentence of
the opening paragraph of this subsection, talks % about coordination being the
cause} \rananote{Shantenu: reference to % infrastructure has been removed to
avoid the confusion.}

%Data coordination within such applications can range from statically %defined
interactions between execution units and data sources to a %fully adaptable
interaction using multi-agent systems, as outlined in %the Industrial Incident
Notification and Response application in %\S\ref{sec:scenarios}.

%\jhanote{Omer: Multi-agent systems are not % called-out/discussed as such in
Section 2} \rananote{Shantenu: This % has been updated to refer to the
particular application when agents % have been described -- rather than the
section as a whole.}

% Workflow techniques and tools provide one of the most widely used % approaches
to expose such coordination to the user.

% \item An analysis of the characteristics of the D3 Science % applications in
\S\ref{sec:distdyndata} suggests that important % issues \jhanote{Important
issues of coordination? distribution? or % dynamism?}  become evident when
considering the ``data topology'' % of the application.

\subsubsection{Understanding Distributed Data}

From \S\ref{sec:distTypes} and \figurename~\ref{fig:figures_application}, we
find that applications with distributed data differ based on the amount or
volume of data distributed, the degree of distribution (scattered over one site
or many sites), and the dynamic aspects associated with distribution, e.g.,
latency and relative scheduling flexibility (and thus tolerance or constraints
on data transfer/aggregation).  Applications processing the largest volumes of
data (O(PB)) typically run on specialized data cyberinfrastructure (e.g., the
Atlas LHC Grid). Applications such as Fusion and NGS that run on smaller data
sets mainly utilize HPC environments. On the other end of the spectrum are
highly dynamic applications that process sensor data and make near-realtime
decisions based on streaming data, e.g., the power grid and sensor network
applications. These applications process much less data; often, sensor data is
also archived and aggregated for historical analyses, which leads to higher data
volumes.

\begin{figure}[ht] \centering
		\includegraphics[width=.7\textwidth]{figures/application.pdf}
                \caption{The distribution, dynamism, data volume characteristics
of the surveyed D3 applications. The clustering of applications with high
dynamism in the bottom right hand corner indicates a similar structure amongst
these applications, namely multiple (sensor-based) sources of data generation
with temporal variation in data generation. Although we do not explicitly
color/shade traditional and infrastructural applications, it is worth mentioning
that there is no correlation between application type and any of the three
``D''s; we attribute this to a strong dependence of degree-of-distribution
infrastructure available for both types of applications.
                  \label{fig:figures_application}}
\end{figure}

% \alnote{Think about how we can add a infrastructure dimension to this %
figure.} \jhanote{As AL and I discussed on a call, in addition to % AL's
suggestion, we need to also use this figure to integrate % section 2-4.}
\jhanote{Also, how would we provide intuitive % explanation of high versus low
dynamism?}  \jhanote{not to endlessly % quibble, but shouldn't ATLAS be more to
the right (and higher % dynamism?) and NGS more the left?}

%\jhanote{the following does not make sense to % me..}  Additional temporal
aspects pertain to the stage(s) of an %application in which data can be
distributed.  Because application stages often have a distribution that makes
sense to that stage, and because that distribution may not be appropriate for
all other phases, applications must either use one particular distribution that
may not work well in all phases or transform the data on input to some stages.
In particular, infrastructural applications require a specific distribution for
the data storage, and thus the stages that read and process the data will need
to make this choice.

As seen, how application data is distributed depends upon which stage of an
application is being considered. In spite of such variation, {\it data topology}
can be used to identify the flow of data through the application, with nodes
representing execution units and edges representing communication features.  %
The overall shape of such a data connectivity pattern gives a sense % of {\it
distributedness}; the thickness of edges gives a sense of % how data intensive
the application is; and the change in the shape % and thickness (and rate) and
number of edges gives a sense of the % dynamism of the application. \katznote{it
seems odd that we describe % a picture here but don't actually have any such
picture in the % paper.}  Additionally, how the application reacts to and
handles dynamism is an important factor and may also indicate that some aspects
of the data topology are unknown.

One aspect of managing data topology is ``data planning," which refers to how
the data is coordinated from ingest through analysis to output. In all cases,
there is a coordination of the data with code used to analyze or transform
it. In some cases these form clear, integrated stages where data flows from one
to another. In others the way that data enters and flows through the application
is more loosely defined and the execution of code is independent. Some examples
include the ATLAS experiments, which have a single point of primary data
generation which fans out to multiple analysis sites followed by mass exchange
of derived data, and sensor grids, which have multiple sources of data
generation collected to a single point of analysis followed by constrained point
of reuse.

We observe that pipelines provide a useful abstraction to represent the
application stages identified in
\figurename~\ref{fig:figures_application-stages}.  Many workflow tools also
enable simple data capture and processing pipelines to be defined, where each
pipeline stage may range from a complete application to a subroutine in Matlab
(for instance).  It is useful to note that pipelines may also be dynamically
constructed and adapted using context and monitoring information, as undertaken
in the Industrial Incident Notification and Response application studied in this
paper. In this instance, the actual pipeline stage to be invoked is triggered by
the availability of particular types of data, a feature also relevant in the
Power Grids application.  However, it is useful to note that different stages of
the pipeline receive different emphases in different applications. Some
applications must collect and store all incoming data, with only minor
pre-processing to clean the data ahead of the main processing. For others, a
major constraint is to reduce the incoming data stream as quickly as possible,
perhaps discarding a large fraction of the data before any processing
occurs. Reuse of pipeline stages across different applications must therefore be
undertaken with care.

Although dynamic and distributed properties have been characterized separately,
it is important to remember, as the above patterns indicate, that distribution
and dynamism become increasingly correlated at extreme scales.  For example,
applications that run on distributed infrastructures are often also highly
dynamic and vice versa.

% \jhanote{There is a problem here. I am not saying the following is %
incorrect, but ultimately coordination mechanisms are being touted % as a way to
support dynamic behaviour if not respond to dynamic % behaviour. However in
Section 3, this is not even hinted at. Needs % fixing IMO. The following should
be trivial: we need to relate which % dynamic behaviour -- infrastructure,
application or data dynamism, % is coordination influencing?}

% \jhanote{this is the first occurence of topology. can this be removed?  % SJ
to reduce last 2 paragraphs into a few sentences}

% \rananote{Update description of data topology -- relate it to % pipelines and
workflows mentioned earlier in the paper. Avoid % confusion in terminology}.
\jhanote{Also can this be related to % section 5.2?}

% Nonetheless, understanding the trend of say, $\beta_1$ as the degree % of
distribution changes can be insightful.

\subsection{Discussion}
\label{sec:discussion}

Now that we have discussed dynamic data and distributed data in the context of
specific applications, we can use this experience to understand the
characteristics that can be associated with D3 science applications, and the
characteristics of a computational infrastructure that may be most suited for D3
science applications.  We believe these are general trends that can be applied
to other applications we have not considered.

\subsubsection{Understanding Distributed Dynamic Applications: }
\label{sec:quantifyingdistapps}

In \S\ref{sec:distdyndata} we characterized dynamic and distributed properties
of the D3 Science applications. Our analysis was qualitative by necessity, due
to the fact that approaches to quantify dynamism and distributed properties were
often intimately tied to application/scenario definition and specific usage
modes.  Notwithstanding, to quantify better dynamic properties of an
application, we believe a term analogous to the use of Amdahl's
number~\cite{10.1109/ICDE.2000.839382} could be used to characterize `static'
data-intensive applications and systems:

\begin{itemize}
\item[] $\alpha$ = Num. of Compute Operations / Amount of Data (Consumed,
Generated or Transformed)
\end{itemize}

Note that this term is scale-invariant, i.e., a value of 1 could correspond to
very small or very large values of both numerator and denominator, and that it
does not capture the time scales over which data changes. However, we believe it
captures most of the non-temporal {\it dynamic} properties.

% \jhanote{we might consider removing the definitions. More text about % what is
$\alpha$ and $\beta$ and say it is early work..}

% defined above for dynamic characteristics,

Similarly, another term can be used to quantify the ``scale'' of distribution,
namely the degree of distribution as a function of the amount of data
transferred, where the degree of distribution can be defined as the number of
distinct data locations:

\begin{itemize}
\item[] $\beta$ = Amount of Data Transferred / Degree of Distribution
% \item[] $\beta_2$ = Number of Compute Operations / Degree of % Distribution
\end{itemize}

%Specifically, we propose $\beta$ as a ratio of the amount of data %transferred
to the number of distinct sites (i.e., degree) that it was %distributed over.

% either the computation involved or For a given application, the % degree of
distribution is often dependent or correlated with the % total number of
computational operations involved or the amount of % data transferred, e.g., for
small computational operations or % data-volumes there is typically a limited
amount of flexibility in % distribution before the overhead of distribution
(either source or % destination) dominates.

% and $\beta_2$ as a ratio of the number of compute operations to the % degree
of distribution; as can be expected, for low amounts of % compute units the
extent to which the tasks can be distributed is % limited, as overhead costs may
dominate.

%Notwithstanding, in this subsection, we provided a framework for

It has proven difficult to provide ``global'' quantitative values of these
ratios, as often the value of the numerators is subject to how an
application/scenario is defined, even more so for a ``class of applications'' as
opposed to a specific application. We believe, nonetheless, that these measures
suggest a possible way to quantify an application's dynamism and distribution
and thereby find possible common approaches solutions.  We plan to explore this
in our future work, and also hope others will consider building on these initial
concepts.

%and thus providing well-defined numbers is difficult,

\subsubsection{Towards an Architecture for D3 Science?}

\jhanote{Note to self: need smoothening}

A survey of data-intensive distributed applications on production
cyberinfrastructure leads us to posit that there are three primary macroscopic
architectures~\cite{pilot_clouds} and approaches for distributed data-intensive
applications: (i) localize all data into a ``cloud'' (or a large analytical
engine) -- the paradigm adopted by several genome projects and the cancer atlas,
(ii) decompose and distribute data to an appropriate number of
computing/analytical engines as available -- the paradigm employed by particle
physics for the discovery of the Higgs Boson (similar to the application in
\S\ref{WLCGSteve}), and (iii) a hybrid of the above two paradigms, wherein data
is decomposed and committed to several infrastructure, which in turn could be a
combination of either of the first two paradigms, similar to what is done in the
Power grids application (\S\ref{power}). \jhanote{improve description....replace
hybrid by bottom-up? }

It is obvious that the first architecture can be used for large volumes of data,
but surely there are limitations to the scalability or validity of this
model. How does this limitation change, qualitatively or quantitatively, if
dynamic data and/or applications are considered? This begets the question: how
fundamental an issue is {\it dynamic}?

The high-energy physics (HEP) community has self-organized towards the second
architectural approach.  Although distributed computing infrastructures
representing this approach, such as OSG and EGI, have been successful for this
community and application, broad uptake across other application types, modes,
and data volumes has not taken place. What lessons does the HEP experience teach
us for the other communities, such as the bioinformatics?  Can an infrastructure
conforming to a particular architecture and developed for a specific application
domain be generalized to another science domain, using the same underlying macro
architecture but by providing a different set of services and specialized data
cyberinfrastructure? Or is the architecture specific to a {\it fixed} range of
compute-data characteristics? If so, what are these characteristics?

%\subsection{Observations}
%
%\katznote{some text needed here.  Note that I promoted this from 5.1.5 to 5.2,
thinking that 5 might be best as discussion, observations, conclusions - how we
order 5.1 and 5.2 is still a question... Maybe call this `takeaway points'?}
%
%\katznote{Omer to go through and move the text/points to either Discussion or
Conclusions... This section will then empty/removed}

%\begin{itemize}

%\item In \S\ref{sec:scenarios}, we divided applications into groups: %
traditional and infrastructural applications.  Traditional % applications have
been built by one or more authors in order to % solve a particular problem or to
answer a science question.  % Infrastructural applications have been built by
groups or % communities to allow the solution of a set of problems, or to answer
% a set of science questions.  In both cases, we can observe that the % same set
of applications stages (as shown in %
\figurename~\ref{fig:figures_application-stages}) can be applied, % with APIs
defined as the inputs and outputs of those stages.  These % APIs are very
informal and loose for the traditional applications, % but more fixed and
conservative for the infrastructural % applications.  These APIs limit how new
paradigms can be used for % the infrastructural applications, unless those
paradigms happen to % fit the existing APIs.  Another way of saying this is that
the % abstractions for dynamism and distribution that can be used in the %
infrastructural applications are those abstractions that the % developers of the
original applications were aware of at the time % the infrastructural
applications were defined.  This leads to a % possible mismatch, where computer
scientists have new ideas that % could be applied to the class of problems that
an infrastructural % application was designed to solve, but that the
infrastructural % application doesn't allow, due to the fixed APIs.  Workflow
systems % are a possible way of solving this problem, as they are often %
sufficiently domain-independent that there does not seem to be a tie % of a
particular system to a particular infrastructural application.

%\item There exist patterns in the described applications; however, %
implementation and deployment specific details often mask these % patterns.  For
example, applications often differ in the way they % handle distribution and
dynamism, such as managing such as % data/compute localities. This makes it
difficult to support patterns % of distribution and dynamism in a
general-purpose fashion, for % anything but the simplest patterns.
%
%\item It is useful to observe that pipelines provide a useful % abstraction to
represent the application stages identified in %
\figurename~\ref{fig:figures_application-stages}.  Many workflow % tools also
enable simple data capture and processing pipelines to be % defined, where each
pipeline stage may range from a complete % application to a subroutine in Matlab
(for instance).  It is useful % to note that pipelines may also be dynamically
constructed and % adapted using context and monitoring information, as
undertaken in % the Industrial Incident Notification and Response application %
studied in this paper. In this instance, the actual pipeline stage % to be
invoked is triggered by the availability of particular types % of data, a
feature also relevant in the Power Grids application.  % However, it is useful
to note that different stages of the pipeline % receive different emphases in
different applications. Some % applications must collect and store all incoming
data, with only % minor pre-processing to clean the data ahead of the main %
processing. For others, a major constraint is to reduce the incoming % data
stream as quickly as possible, perhaps discarding a large % fraction of the data
before any processing occurs. Reuse of pipeline % stages across different
applications must therefore be undertaken % with care.

  % Managing and coordinating the behavior of pipeline execution, for %
supporting both file-based and streaming data, has been % extensively studied in
computer science (supporting synchronous % (blocking), asynchronous, concurrent,
threading semantics, for % instance).

%\item

%\jhanote{mismatch between abstraction levels, reference SAGA % paper ``leaky
abstractions''} \katznote{Shantenu to incorporate this % into the leaky
abstraction stuff above} \katznote{A possible problem is % that distributed
systems (infrastructures and perhaps applications) % are so different from each
other that there is no common system % abstraction that is both general enough
to work on all of them and % specific enough to be useful on any of them.}

%\item The trend toward more unstructured and more dynamic data sources lead to
%higher demands on data processing infrastructure.  % provide simple
abstractions or frameworks to % Higher-level abstraction need to provide
sufficient level of control % of infrastructure trade-offs (such as data/compute
localities), % while hiding a lot of the complexities (e.g.., the management of
% replicas).\jhanote{andre to try to link back to section 4 and % elaborate}

% This indicates the difficulty of % using existing abstractions and
infrastructures.

%\jhanote{Worth having in this section. Needs integration}

% \item Often the coupling between the various components/services that % are
part of the workflow can differ -- ranging from hardwired % interactions between
components to those that can be defined and % modified by an application
developer. We can therefore distinguish % between workflows that are embedded
within an application, to those % that are explicitly defined using a
composition tool (such as % Taverna, Triana, Kepler, etc). Workflow systems
often utilize a % variety of abstractions related to: (i) mapping of %
services/components to computational/data resources; (ii) data %
movement/migration; and (iii) coordination between various % components.  And, %
in many cases, they do not appear to be widely used where we might % think they
would be, perhaps due to lack of publicity, steep % learning curves, or other
issues.

%\end{itemize}

%\subsection{Points + Hypotheses}

\subsection{Conclusions}
\label{sec:conclusion}

Many fields, such as biology, that until recently were not characterized by
their intensive use of data are rapidly becoming data-driven.  The growth in
data has profound implications for the design and implementation of services and
infrastructure.  For example, there is a lifecycle to data that has often been
ignored until recently: the decreasing cost of storage has made it attractive to
archive \emph{everything}, just in case it becomes useful. This approach is
becoming untenable in many areas: some datasets are simply too big to archive,
and so must be summarized and their raw form discarded (possibly in real
time). Others are easier to regenerate than to store.  Still others are too
transient to justify their storage, and may in aggregate be too big to store
even when being small individually.

Additionally, it is useful to observe that the trend toward more unstructured
and more dynamic data sources lead to higher demands on data processing
infrastructure.~\cite{snia_2012} Any data science architecture needs to
explicitly recognize, capture, and manage the lifecycle requirements of
different datasets, within the constraints of available time and
storage. Associated with the growth in data volumes are increasing levels of
dynamism and distribution; sometimes distribution is logical, but sometimes the
data is physically distributed for reasons of performance or convenience. At
large scales, dynamism becomes intrinsic to the application and infrastructure;
sometimes it occurs because of changing experimental or analytic conditions,
sometimes due to the availability of infrastructure and resources.  %We close
out this paper with a brief %discussion of a range of issues associated with the
distribution and %dynamism of data.

We have identified many different application approaches, programming systems,
and infrastructure techniques applicable to dynamic data.  Consistent with
experience in other related fields, there seems no reason to expect that these
approaches and techniques will disappear in the near future. While a convergence
in approaches and techniques might be expected, this has proved to be elusive;
there remains a diversity in the design, implementation, and execution of
scientific codes.  It may even be argued that it is premature to have
convergence; after all, D3 Science is still an emerging domain.

From the previous discussion we can draw out some conclusions about the trends
within dynamic, distributed, data-intensive science. For instance, in
\S\ref{sec:scenarios}, we divided applications into groups: traditional and
infrastructural applications.  Traditional applications have been built by one
or more authors in order to solve a particular problem or to answer a science
question.  Infrastructural applications have been built by groups or communities
to allow the solution of a set of problems, or to answer a set of science
questions.  In both cases, we can observe that the same set of applications
stages (as shown in \figurename~\ref{fig:figures_application-stages}) can be
applied, with APIs defined as the inputs and outputs of those stages.  These
APIs are very informal and loose for the traditional applications, but more
fixed and conservative for the infrastructural applications.  % \jhanote{is
there a basis for these conjectures?} \katznote{just the % evidence of the
applications themselves} These APIs limit how new application paradigms
%\jhanote{new paradigms of what?}  can be introduced for the infrastructural
applications, unless those paradigms happen to fit the existing APIs.  Another
way of saying this is that the abstractions for dynamism and distribution that
can be used in the infrastructural applications are those abstractions that the
developers of the original applications were aware of at the time the
infrastructural applications were defined.  This leads to a possible mismatch,
where computer scientists have new ideas that could be applied to the class of
problems that an infrastructural application was designed to solve, but that the
infrastructural application doesn't allow, due to the fixed APIs.  New tools may
be developed to address a narrow set of features while ignoring the larger set
of problems that prevented earlier tools from being used.  These tools then may
be used be new applications as those applications are developed; however, they
are unlikely be used by the existing applications they were meant to help.

This is because applications (particularly infrastructural applications) aren't
developed in a vacuum, and that the constraints on the applications lead to new
tools that are or could be developed not being able to be applied in practice.
For example, while workflow systems are often sufficiently domain-independent
that there does not seem to be a tie of a particular system to a particular
infrastructural application, there are a very large number of workflow systems
that exist, and it is common in the authors' experience to hear about new ones
being developed from scratch because ``the ones that exist don't quite do what
is needed for my application or system.''

\katznote{previous 2 paragraphs to be re-read.  Old comments about them are
available as commented out comments.}  %\jhanote{Irrespective, I find this past
paragraph % somewhat speculative. And I don't see the point of it? WHat is the %
latter half trying to say: there are problems with APIs and % interfaces of 3D
applications and systems?} \katznote{It's saying % that applications aren't
developed in a vacuum, particularly % infrastructural applications, and that the
constraints on the % applications leads to new tools that are or could be
developed not % being able to be applied in practice.  In some ways, it's a call
for % an application abstraction or middleware, between the applications % and
the `infrastructure' as defined in \S4, much as perhaps AIMES is % trying to
build}\jhanote{Dan, I find your explanatory note more % convincing than the
sentence on workflow systems! Also there is no % mention of middleware as the
``impedance mismatch lowering layer'' % in the above. I propose we rewrite the
sentence starting, ``Workflow % systems are a possible way..} \katznote{ok, I
used some of my text % and turned around the sentence from Omer's optimistic
view to % my more pessimistic view.  If others disagree, we can change it back %
or compromise.}

As part of this survey, we also observe that there exist patterns in the
described applications; however, implementation and deployment specific details
can often mask these patterns.  For example, applications often differ in the
way they handle distribution and dynamism, such as managing such as data/compute
localities. This makes it difficult to support patterns of distribution and
dynamism in a general-purpose fashion, for anything but the simplest
patterns. In \S\ref{sec:quantifyingdistapps} we have attempted to initiate a
discussion in this context however by considering two ratios %(identified as
parameters $\alpha$ and $\beta$) to relate data volume and distribution with
compute operations performed by an application. We believe determining values
for these parameters would provide a useful starting point for comparing
infrastructure requirements of Big Data applications.  %\jhanote{I transposed
the last two paragraphs}

% \item Workflows seem to be a natural idea for expressing both % traditional
and infrastructural applications, and indeed, there are % a number of workflow
tools that are used in both cases.  It might % also seem that an infrastructural
application would have a specific % workflow system that is in common use, since
different % developers/users are all doing similar computational work, but this
% does not appear to be true, at least in the applications we studied.

  % \katznote{I would like to say that specific workflow systems are %
successful when embedded inside traditional applications, % because they then
become part of the application, but I am not % sure I have a good example here
from \S\ref{sec:scenarios}.  % Does anyone else?}

% \begin{itemize}

% \item \jhanote{Is the following observation, hypothesis and how do we % link
them?} \katznote{Can we say something about applications that % deal with files
vs those that deal with streams (stream-based are: % PowerGrid, Fusion, Incident
Response, network monitoring) PowerGrid % and Incident apps have centralized
stream processing;} Fusion and % network have distributed stream processing.
Not sure what to say % here.  Maybe this point isn't a point... \jhanote{Look at
tables % \ref{tab:app_distributed_processing} and %
\ref{table:progSystems}. From these tables can we say something % about
``common'' ways of say distribution for ``stream based % applications''}
\jhanote{analogy with abstractions for workflows: % num. of data elements, num
of streams, num. dynamic resources}

% \end{itemize}

\section*{Acknowledgements} This paper is an outcome of the UK e-Science
Institute sponsored Research Theme on Dynamic Distributed Data-Intensive
Programming Abstractions and Systems. This work is also sponsored by NSF
OCI-1059635.  We thank Jon Weissman and Jon Blower for useful initial
discussions, as well as 3DPAS workshop attendees. Discussions of the selected
applications benefited from the generous help of Adam Barker, Jon Blower, Julian
Borrill, Silvia Delgado Olabarriaga, Steve Fisher, Keith Jackson, Scott Klasky,
Bob Mann, Don Middleton, Kees Nieuwenhuis, Manish Parashar, Stephen Pascoe, and
Jon Weissman, though any errors are the responsibility of the authors.  Some of
the work by Katz was supported by the National Science Foundation while working
at the Foundation; any opinion, finding, and conclusions or recommendations
expressed in this material are those of the author(s) and do not necessarily
reflect the views of the National Science Foundation.

\bibliographystyle{unsrt} \bibliography{3dpas}

\end{document}

{\em Workflow} systems have already received significant attention in the
collaborative e-Science domain.  Such systems provide a useful means to
aggregate capability and subsequently deploy the resulting workflow over
multiple, distributed resources. Workflows can be ``abstract'', i.e., only the
connectivity graph is identified but is not bound to particular executables and
particular resources, or ``concrete'', i.e., the executable components are
identified along with their execution environment. Generally, an abstract
workflow needs to be mapped to a concrete workflow graph to enable
execution. This may be achieved dynamically using a variety of different
user-specified criteria, ranging from cost of execution (such as with the
availability of remotely accessed resources from Amazon.com~\cite{berriman11}),
Quality of Service, etc. This mapping can also be adaptive, especially if each
concrete workflow is dynamically composed based on event triggers from the
execution environment or constraints identified by a user. We may therefore
identify a range of possibilities:

\begin{itemize}

\item An abstract workflow graph is statically defined and mapped to a known
resource/infrastructure pool.

\item An abstract workflow graph is dynamically mapped to an infrastructure
based on user-specified constraints (related to cost of access, Quality of
Service, etc.)

\item An abstract workflow graph is dynamically enacted over a varying
infrastructure pool (often making use of an external resource management
system.)

\item An abstract workflow graph is used to dynamically compose a concrete
workflow using techniques from AI planning and model checking.

\end{itemize}

Workflow tools such as Taverna~\cite{Taverna}, Kepler~\cite{Kepler},
Trident~\cite{Trident} and others~\cite{iantaylorsbook} \jhanote{Are any of
these used by applications? If not I propose they not be mentioned} enable
reusable code for scientific analysis by expressing it as an explicit dataflow
or control flow composition of components, each serving a dedicated
function. This has simplified the development and the maintenance of complex
codes, both by allowing common component re-use and by highlighting the
structure of the processing pipeline. Workflow tools therefore provide both
infrastructure for supporting the construction and enactment/execution of
scientific applications and components that be used alongside application
specific infrastructure. The granularity of a workflow tool can therefore range
from simple components to complete infrastructures (i.e. enabling integration of
other components and supporting their enactment/execution).