### 2015-05-29 old-sections/3dpas_programming.tex

\section{Programming Systems and Abstractions}
\label{sec:prog_sys}

% \note{The issue of programming systems for scientific applications is always
% vexing. On the one hand, scientific codes place requirements on
% applications that are often substantially different from commercial
% applications in terms of their data volumes and flow rates -- although
% this is becoming less true as industry discovers the value of ``big
% data''; on the other hand, scientific codes must increasingly
% interoperate with other services on the web, and so must appeal to
% standard protocols, data formats and libraries. The latter can be at
% odds with addressing the former, since it complicates the use of
% radically different programing models and tools.  However, the
% convergence of scientific and commercial interest in large-scale
% dynamic data should lead to improvements in the approaches applied in
% both domains.}

%\subsection{Architecture}

% Top-level views of a system is typically as a processing pipeline from
% collection, through pre-processing, to storage and then
% post-processing and querying. (Collection includes both simulation and
% sensing.) Different applications place different emphasis on the
% different stages: some collect and store all the data they can, with
% only minor pre-processing to remove noise; others must perform
% enormous pre-processing to reduce the data volumes to manageable
% dimensions. Another point in this spectrum is a classifier system that
% uses a data stream to recognise higher-level events which it then
% propagates, so it's the events that are interesting, not the data.

% \note{Scientific codes have long lifespans, as tools are refined and
% extended along with improving theoretical models and experimental
% techniques. Many experiments are inherently long-lived, for example in
% fundamental physics or space science, with the associated analysis
% software often being refined over the length of the mission.}

The design of scientific software evolved from an early emphasis on fully-functioning 
  monolithic applications to a library-based
  approach with a core of reusable code by one or more,
  perhaps changing, user environments. More recently there has been a
  further move to re-use through the use of workflow tools such as
  Taverna~\cite{Taverna}, Kepler~\cite{Kepler}, and
  Trident~\cite{Trident}. A data analysis process is
  expressed as an explicit dataflow or control flow composition of components each serving a
  dedicated function. This has simplified the development and the maintenance of complex
  codes, both by allowing common component re-use and by highlighting
  the structure of the processing pipeline.
% Unsurprisingly, more
  % recent domains such as genomics have adopted such technologies from
  % the start; older domains still have substantial legacy codebases
  % with less transparent structuring} \katznote{previous sentence is
  % far too absolute to be correct.}

% The view of scientific applications as pipelines hides significant
% differences between (and within) domains, since the different stages
% of the pipeline receive different emphasis in different
% applications. Some applications must collect and store all the
% incoming data, with only minor pre-processing to clean the data ahead
% of the main processing. For others, a major constraint is to reduce
% the incoming data stream as quickly as possible, perhaps discarding a
% large fraction of the data before any processing occurs. Taking some
% of the examples from \S\ref{sec:scenarios}, medical imaging and
% astrophysics often fall into the first category, whilst particle
% physics (especially in the case of the Large Hadron Collider) often
% falls dramatically into the second -- to the extent that principled
% data reduction forms one of the major scientific challenges for the
% project.

% \note{One way to integrate legacy code -- or those codes too
%   tightly-coupled for a workflow structure -- is to build pipelines
%   from ``macro'' components. In this view the pipeline connects groups
%   of components that are not themselves pipelines. Conversely one
%   might have multiple pipelines running concurrently over a partially
%   shared set of components, driven by data dependencies and exchanging
%   data in more complex ways than are seen in simple pipelines.}

% There are also significant variations in where ``the pipeline'' begins
% and ends. Many systems are driven by data sets collected and shared
% widely, against which several different applications may be
% applied. Climate science is a good example of this approach. In this
% case the data is ``pre-collected'' in the sense of existing outside
% most of the applications, although its collection may have been a
% complex task. The raw data required for gene sequencing
% (\S\ref{bioSilvia}), for example, while being collected using a
% dedicated and very sophisticated computerized processing system, would
% generally be regarded as ``given'' and not part of the processing
% pipeline, since the pipeline would not affect the collection
% task. Another way to look at this is that data is collected
% \emph{independently} of its further processing.

% Alternatively the application may be driven directly
% from a live data stream, with processing happening on the
% fly. The front-end data collection sub-system is becoming increasingly
% flexible and capable, with the introduction of sensor networks to
% replace more passive approaches, and it may often make sense to
% include the sensor network into the application workflow pipeline. All
% these choices have implications for design and provisioning to which
% we return below.

% % Question: where does each application lie in this space?

% % Question: is it the case that each application can be placed at an
% % appropriate point \textit{a priori}, or do characteristics change?

% The choice of which sort of pipeline a particular application uses may
% well be fixed, or will change only slowly as the experimental
% environment evolves. In particular we see some systems which (for
% example) add simulation alongside experimental data collection to act
% as a verification and validation step.

% \note{Viewed from the perspective of the 3DPAS vectors
% (\S\ref{sec:vectors}), programming systems can be seen to exert
% comprehensive influence. In terms of \textit{communication and data},
% the key question is the complexity and extent of the system being
% discussed: does it extend to the collection of data, or to its
% presentation and interaction, or does it simply focus on the
% ``traditional'' processing steps? Does it provide a simple, coherent
% structure (like a pipeline), or does it require more complex,
% \textit{ad hoc} communications? Similarly, the \textit{coordination}
% vector is influenced both by how a computation adapts to on-going
% changes and to where control of this adaptation resides. The
% \textit{execution environment} vector is more centrally the topic of
% this section, and we explore these issues in more depth below,
% \S\ref{sec:issues} after examining some alternatives to pipelined
% architectures.}

%\subsection{Beyond Simple Pipelines}

% Question: what are the issues that face developers in using standard
% scientific workflow approaches?

The workflow approach is simple and attractive to compose sequential data processing pipelines and
offers the flexibility of modeling more complex constructs. However, more specialized programming
abstractions are becoming popular to address large scale and dynamic data processing.
% abstract-away many of the issues that are of immense relevance to
% efficient scientific computing. We will first 
We consider two alternative paradigms of bulk and stream data processing, before considering the
issues of adaptation and provisioning of these models.
%  and otherwise refining scientific
% workflows. Note that in many cases these alternative paradigms
% actually integrate very well into an overall pipeline, architecture,
% providing components that can fit into a general pipelined framework.

\subsection{MapReduce and Bulk Synchronous Parallel (BSP) Models}

% Often limited to pipeline \textit{versus} some stylised large-scale
% process such as map-reduce, within the large architectural
% pipeline. Often find scientific workflow systems like Taverna
% describing pipelines in terms of components, possibly with specialised
% high-performance constructions for some intensive
% components. Generally find this in the ``traditional'' space of codes
% running on workstations and servers; sensor networks themselves are
% immature in this respect and typically focus on robust data
% transmission without much in the way of in-network processing
% (although this is changing, slowly).

% Many criticisms of traditional programming languages begin from the
% challenges posed by the so-called \emph{von Neumann bottleneck}
% introduced by processing data in small chunks, for example within
% iterative loops of tail-recursive functions. In either case, the
% bottleneck comes from imposing linearity into the way the data is
% processed. Many high-performance approaches, perhaps beginning with
% various vectorizing Fortran compilers, have sought to sidestep the
% bottleneck by providing \emph{bulk} operators that express
% computations over collections of data, that can be more effectively
% parallelized.

The MapReduce paradigm~\cite{MapReduce} is the most prominent and recent example of
using \emph{bulk} operators that express computations over collections of data, that can be more
effectively parallelized.  MapReduce forces programmers to model computation in two
stages: a \emph{map} phase in which a transformation is applied to
every element of a collection independently of every other; and a
\emph{reduce} phase in which the transformed results are aggregated
through a binary function that is associative and commutative. These
properties serve to guarantee that the MapReduce computation can be
efficiently parallelized, by performing the map phase concurrently and
then reducing without the reduction order affecting the final
result. The result is to provide a large amount of ``excess''
parallelism that can utilize as many cores or processors as are
available.

% MapReduce arose in part from earlier work on ``skeletal parallelism''
% that generalize the idea of bulk operators to encompass any
% structure that can perform efficient parallel computation. The
% skeleton is essentially a higher-order function that captures the
% structure of the computation, which is then populated with
% application-specific functions to provide the actual calculation. The
% skeleton ensures that the overall structured computation can be
% computed efficiently, potentially making use of significant code
% generation and optimization. Flask~\cite{FlaskHaskell} is a modern example of this
% approach applied to sensor networks.

Whilst often not suitable as a complete system architecture,
MapReduce can be used in a workflow environment to provide scalability to
processor-intensive components, or for data-intensive applications
where it can address the relatively low scalability of I/O
capacity. It does not necessarily work effectively in a streaming
mode, and cannot be relied upon to adapt its structure according to
different data volumes: the structure is generally parameterized by
the processing resources available regardless of the data, rather than
making requests for appropriate resourcing based on the data
presented~\cite{DataIntensiveMapReduce}. Variants of the basic MapReduce model have been proposed,
such as iterative~\cite{twister} and hierarchical~\cite{hmr} MapReduce, to address some of these limitations. Framework implementations
such as Apache Hadoop~\cite{hadoop}, Microsoft Daytona~\cite{daytona}, and Twister~\cite{twister} are available, among others.

Bulk Synchronous Parallel (BSP) processing~\cite{valiant1990bsp} was developed in the 1990 as a bridging model for
parallel computing. This programming abstraction is garnering attention as a way to reduce the I/O complexity of
MapReduce and offer an easier message passing model than MPI. It is being used for data intensive
applications such as matrix and graph processing, with framework implementations by Apache Hama~\cite{hama},
Google Pregel~\cite{pregel} and others~\cite{goldenorb}. The key approach is to separate processing into supersteps where independent
processes operate concurrently on messages received from the previous superstep. This processing is
interleaved with message passing to processes in the next superstep. When a process completes its
execution of a superstep, it reaches a barrier that is synchronized with all other processes in the
superstep. Once synchronized, the next superstep starts.

\subsection{Stream and Event Processing}

% Stream processing is a variant on pipelines where the stream is
% processed locally and in real time, typically with only fairly small
% operations allowed at each stage (for example calculating time
% derivatives of a time series).

Stream processing is an approach modeled on signal processing, in
which a stream of data is transformed on-the-fly to produce one or
more further streams of values. As such it is a programming abstraction suitable for
continuous data feeds, or when results need to be pushed to users of interactive applications. Since
the data is not assumed to be available 
\textit{en masse}, it can be used in situations where storing the
entire data set is undesirable, unnecessary or impossible. Equally,
the approach is simple to understand and mirrors the composition of workflows. However, since
they operate continuously on a seemingly infinite stream of data, notions of data freshness,
processing latencies, resilience/checkpointing, and incremental completion become more
interesting. These introduce new data and composition abstractions
(e.g., Landmarks~\cite{chandrasekaran:telegraphcq2003}, Dynamic Stream
Rates~\cite{Simmhan:sciencecloud:2011}) and trade-offs during execution.

A simple example of stream processing occurs at the front end of a
pipeline, in data cleansing. A processor might perform simple sanity
checks such as outlier removal based on expert knowledge or on
properties calculated from the stream itself -- for example to remove
any data that lies more than two standard deviations from the running
mean. The analogy to signal processing is quite clear: the stream is
processed as a sequence of samples, used to extract stream-level
values such as moving average (mean), time derivatives and the
like. These values may be used in further processing, or may be output
in their own right alongside the raw data and used to feed other
processing steps.

A higher level of abstraction in stream processing is \emph{complex event processing}, where a
stream is processed to extract ``events'' that 
are notionally occurring at a higher semantic level than the raw
data. An example might be the detection of lights being left on in a room (ambient light sensor
stream) when no one is occupying it (infra-red sensor stream). The significant point is that the raw data
stream pre-processed, classified and packaged as a stream of events with timestamps, identifiers and properties on which
higher-order operations like standing-queries and pattern-matching can be performed. Events can even
be associated with semantic annotations~\cite{Teymourian:EDBT:2010, Zhou:debs:2011}.

There are many examples of stream
~\cite{chandrasekaran:telegraphcq2003, abadi:borealis2005, liew:streamgraph2010} and event
~\cite{ESPER, siddhi, MicrosoftInsight} processing frameworks, and recent works have combined
streaming with more traditional file-based workflows~\cite{ZinnCluster,Herath:CCGrid:2010}.

% Question: are there any significant application codes that use
% anything other than pipelines, streams and map-reduce?

% \note{\subsubsection{Specialized Algorithms} Inevitably many
%   scientific codes do not fit into either of these broad schemata and
%   so must be coded in a more \textit{ad hoc} manner. Pattern-matching
%   for gene sequencing (\S\ref{bioSilvia}) and particle physics
%   (\S\ref{WLCGSteve}) are good examples, where domain-specific
%   knowledge is used to optimize traditional pattern-matching.}

% \note{Programming systems such as MPI~\cite{MPI} provide a basis for distributed
% applications, although it is vital not to under-estimate the
% difficulties involved. While such systems allow developers to generate
% very complex patters of distribution and concurrency, they do so at
% the cost of a massive increase in coding complexity -- often beyond
% the programmers' capacities to manage efficiently. Significant
% performance figures are often elusive, and may be fragile in the sense
% that small changes in approach and algorithm destroy the
% gains. Systems that \emph{can} be phrased according to a structured
% algorithm or processing schema with known performance benefits
% therefore \emph{should} be.}

% \subsection{Issues}
% \label{sec:issues}

% \note{So far little we have said has addressed the essential aspects
%   of dynamic data, in particular the unpredictability and scaling
%   features that seem inherent in dealing with such data sets and
%   sources: the \textit{application adaptation} vector. However, it is
%   clear that dynamic data raises some important questions about the
%   management of scientific codes, many of which cross between the
%   various paradigms chosen for expressing the problems.}

% % Question: is there a problem with multiple applications running on the
% % same fabric and interfering? -- for example a classifier system
% % running alongside a collection system and generating storms of events?

% \note{We may divide the issues raised into two broad categories. The first
% pair of \textit{intrinsic} issues capture concerns such as
% provisioning and adaptation that relate to codes themselves in terms
% of their development, composition and management. The second set of
% \textit{extrinsic} issues are more concerned with software engineering
% factors such as the extent of the pipeline and the representation of a
% dataset's history and provenance.}

\subsection{Provisioning}

Chief among the management questions for dynamic data processing frameworks is the
issue of \emph{provisioning}: allocating sufficient and appropriate
resources to a component to allow it to perform most efficiently. This
must be set against the need to manage the interactions of components,
leading to a difficult resource allocation and scheduling problem.

Provisioning can have two different goals: one of optimally using
resources (maximizing resource usage), while the other to ensure
efficient application execution. While the former goal optimizes
provisioning from an infrastructure provider's perspective, the latter
goal is more driven by the application's quality of service (QoS) requirement -- something that
is necessary to be met for the results of the application to be
useful. The latter can be expressed as in specific terms (e.g. makespan, freshness, throughput), or may
simply be a best effort. Expressing requirements becomes difficult in
itself with dynamic applications that may produce continuous,
intermediary results.

Even if one can appropriately calculate the resource needs of a
component for a given data set -- which is in itself non-trivial --
dynamic data adds the need to deal with an unknown and possibly
changing optimal allocation. Changes to data rates, sizes, number of streams and processing
complexities can impact resource needs. Further, the programming
model itself can be inflexible when it comes to allowing dynamic
provisioning. Algorithms may not allow more nodes to be added midway
through a computation, and some data processing systems require all tasks to
be mapped to resources at the time the execution launches. All these
restrictions impede flexible adaptation.

% Commercial systems have addressed provisioning in part by using
% \emph{virtualization}, splitting a computation between a variable
% number of identical virtual machines that can be spun-up and -down on
% top of a smaller number of more capable servers. Scientific codes
% highlight many of the broader problems and benefits of
% virtualization. It can be hard to determine \textit{a priori} the
% computational needs of a component, which makes it hard to
% provision. Co-locating an unexpectedly large component with others may
% impact on the overall performance of the system. However, virtualizing
% the components as independent virtual machines potentially allows
% components to be frozen and re-located, allowing dynamic
% re-provisioning and optimization. This of course leads to further
% problems in the analysis and planning of such operations, but at least
% provides a technical basis for managing the performance of services
% across their execution lifetimes.

While most attention is given to
scaling-out computations (to handle more data), scaling down to
prevent waste of resources is equally problematic. This is particularly so as 
\emph{cloud computing} gains prominence as a cyber-infrastructure. Cloud platforms take
resource virtualization to another level, 
presenting computing as an elastic service than can be ramped up and down.
However, current cloud computing platforms are
highly targeted at scenarios of commercial interest such a
large volumes of independent transactions. While these may match some
scientific problems (such as some issues described in
 \S\ref{bioSilvia}  and \S\ref{WLCGSteve}), it is by no means suitable for
others. The multi-tenancy of clouds whereby the same resource is shared across customers also
introduces dynamism in the performance offered by the
infrastructure~\cite{iosup:ccgrid:2011}. 
Programming models such as MapReduce and the ease of getting access to resources have increased the
use of clouds for eScience, and cloud vendors such as Amazon have also responded to scientific
computing needs with HPC and GPGPU infrastructure. \ysnote{should we move the preceding line to ``D3 Infrastructure''?}
Significantly more research is needed to determine 
the gaps in existing cloud computing platforms, be it on the cloud provider's end or in developing
novel scientific programming models  to meet the needs of the broader scientific domains.

% \note{\subsubsection{Co-existing Services}}

% \note{Provisioning also extends to the co-existence of multiple services and
% components on the same substrate, where their provisioning needs may
% conflict or interact in peculiar ways. Indeed, much of the interest
% that computer scientists have in virtualization comes from these
% interactions. The variability of provisioning makes these challenges
% even more severe for scientific codes.}

% \note{At another level, service co-existence is strongly linked with
%   service-oriented architectures. Scientific codes that provide common
%   and widely-utilized services, for example in astronomy
%   (\S\ref{astroSOAAdam}), demonstrate the potential that these
%   architectures have. Even in systems for which there are fewer
%   outside users, SOA is a useful structuring mechanism within
%   pipelines and can be deployed generically by workflow engines. The
%   advantages in terms of the availability of re-usable components --
%   whether or not such re-use is foreseen -- are potentially highly
%   significant in fields (such as climate modeling, \S\ref{modisKeith})
%   where variations on analytic themes are commonly employed. SOA also
%   raises the issue that the program may be using existing shared or
%   external services that cannot be relocated because it is not in the
%   workflow author's control. It may or may not be possible to locate
%   the user-controlled service to co-locate with the external shared
%   service.}

\subsection{Adaptive Programming Systems}

% Question: would there be any need to push lots of processing to the
% front end, near the sensing? And if so, what would this do to the
% later stages?

As noted earlier, it is often not clear where a particular scientific
pipeline begins and ends. In particular we are seeing increasing use
of sensor networks and other intelligent front-end data collection
systems, as well as the integration of simulation and stream
processing. These components often have the feature of being
\emph{adaptive}, in that they can vary their exact behavior according
either to external control stimuli or from observations made of the
data they are themselves observing.

Adaptation may also be through users' interactively changing the
direction of the pipeline, either by changing parameters or by
adding or removing tasks in the pipeline. This allows decision logic
to be initiated by the pipeline itself, or using an external
control signal that may come from users or an instrument. This form of
``computational steering'' is especially useful in systems being used
to explore a data set.

What impact does this have on other stages? The advantage of including
such systems is that the pipeline then provides an end-to-end
description of the entire scientific process. This is valuable both
for documentation purposes, but also in pursuing issues of component
and data re-use.

% A lot of scientific processes are predicated on data being collected
% and then analysed off-line, which emphasises both query performance
% but also the collection and use of meta-data about the data set (for
% example the precision of sensing and the provenance effects of any
% pre-processing).

% \note{Adaptivity brings its own challenges, many of which are often captured
% under the notion of \emph{autonomic systems} which sense their own
% operation, decide on appropriate management actions to maintain their
% quality-of-service parameters, and then act to affect their own
% operation~\cite{KephartAutonomics,AC-Survey}. This is substantially
% different from the user-driven adaptations discussed above, focusing
% on managing a process rather than changing it on the fly. A
% substantial amount of research is already on-going in computer
% science, often centered on data centre management and provisioning (see
% above); an emerging strand of work is looking at autonomic techniques
% applied to scientific systems. The importance of these approaches is
% that they allow improvements in the service offered to workflows by
% adaptive components. However, this is accompanied by the danger that
% adaptation may not be scientifically benign, and may affect the
% results of the workflow itself. This leads on to issues of data
% provenance below.}

% \note{Many current techniques in autonomic systems approach the
%   problem very much from a systems perspective rather than from a
%   scientific or workflow perspective. This matters, as it makes it
%   harder for users to affect, control or understand the impact that
%   autonomic adaptation may have on their overall analysis. (An example
%   of this is the adaptive processing of a sensor data stream that --
%   if not recorded -- may affect the validity of later statistical
%   analyses. These are to some extent issues of provenance and are
%   discussed more below.)}

% \note{Including sensing into the pipeline (\S\ref{sensorSimon})
% allows analysis to inform collection. A good
% example is where a feature-recognition analysis detects a feature of
% interest, which can then be examined in more detail. Not all
% application areas are amenable to this approach, for example where
% data is collected \textit{en masse} and at a consistent granularity
% and regularity, perhaps independent of \emph{any} particular
% processing function. For many systems, however, the ability to
% feed-back analysis from later stages is of great
% significance. Understanding the stability and dynamics of such
% feedback loops remains a challenge in autonomic systems, especially
% when the feedback deviates from the simple closed-form specifications
% amenable to control theory: more dynamic approaches from game theory,
% physics and econometrics show promise.}

% \note{At the other end of the pipeline, interactive exploration and
%   computational steering (\S\ref{envJon}) involve the scientist and
%   data analyst in providing guidance as to the way in which
%   computation should proceed, affecting either current or future
%   runs. The challenge here is that the ``user'' at this stage is often
%   not the ``programmer'', and so may only be able to provide guidance
%   at a scientific level, rather than at a algorithmic or architectural
%   level. This having been said, such guidance is potentially extremely
%   valuable and allows insights directly from the level of the
%   ``mission'' of the scientific system. This suggests that data
%   visualization and interaction should be considered core to the
%   processing pipeline, and that attention be paid to how scientific
%   and data insights can be fed-back as autonomic control
%   parameters. This is not an area that has received any significant
%   attention from the autonomic systems community, and so offers a
%   major research avenue.}

% %\subsubsection{Collecting Provenance}

% % Question: is there a need to retain meta-data across the lifetime of a
% % system and across the complete architecture?

% \note{Finally we must consider the fact that data processing in a scientific
% setting is itself a scientific process, which introduces properties
% into the data that must be accounted for in a predictable and
% well-founded manner. This implies that the process is well-defined,
% describable and analyzable to explore its effects on the results
% obtained.}

% \note{A simple example of the effects of different choices comes from data
% cleansing. Deciding which elements are ``noise'' in a data set (or
% data stream) may exert a profound effect on the results of further
% processing. Discarding data that is \textit{not} noise will distort
% the results; discarding too many data points may damage the
% statistical properties of the data; while discarding too few may be
% distorting in other ways. The point is not that the data is
% transformed in ways that change its properties: rather, the issue is
% the \textit{hiding} of these transformations in ways that cannot be
% accounted for in future processing stages or interpretation.}

% \note{There is no way that a programming system or model can automate
% decisions of this kind, which remain essentially scientific. What a
% system \textit{can} do, however, is to ensure that the decisions made
% are recorded and propagated with the data and its results. This is the
% issue of the \emph{provenance} of the data and the results: recording
% the transformations that have been made to a data set or data stream
% and associating this \emph{meta-data} alongside any derived data. We
% can view provenance meta-data as a generalized kind of error bar on
% computations -- or, equally, as the information needed to produce
% standard error bars in a well-founded and traceable
% manner.}

% \note{There has been several years' work in developing a community standard,
% the Open Provenance Model~\cite{OPM}, which is leading to more formal
% standards such as the efforts of the W3C Provenance Working Group to
% provide frameworks for the capture and representation of
% provenance. Scientific workflow systems like Taverna are increasingly
% able to handle provenance attachments automatically and maintain them
% between processing stages~\cite{Taverna}, and there are dozens of stand-alone
% provenance systems as well. One may regard this as another example of
% moving analysis into the computer systems and away from the scientists
% to ensure faithful recording.}

% \subsection{Coverage}
% \label{sec:coverage}

% \begin{table}
%   \begin{scriptsize}
%     \begin{center}
%       \caption{Programming systems used by the applications\label{table:progSystems}}
%       \begin{tabular}{|p{6.3cm}|p{7.2cm}|}
%         \hline Application  & Current approaches \\
%         \hline
%         \hline NGS Analytics & pipelines that ship with
%         sequencers~\cite{illumina}; workflows~\cite{Taverna} \\
%         \hline ATLAS & custom algorithms; job submission tied to WLCG \\
%         \hline LSST & custom layer; scripted workflows; queries \\
%         \hline SOA Astronomy & web services \\
%         \hline Cosmic Microwave Background & \textit{ad hoc} \\
%         \hline Sensor Network Application & focused on collection \\
%         \hline Climate & \textit{ad hoc} \\
%         \hline Interactive Exploration of Environmental Data &
%         visualization tools \\
%         \hline Power Grids & stream processing; data mining \\
%         \hline Fusion (ITER) & scripted pipeline \\
%         \hline Industrial Incident Notification and Response &
%         agent-based  \\
%         \hline MODIS Data Processing & scripted pipeline \\
% %        \hline Floating Sensors & \textit{ad hoc} \\
%         \hline Distributed Network Intrusion Detection & monolithic
%         packages \\
%         \hline
%       \end{tabular}
%     \end{center}
%   \end{scriptsize}
% \end{table}

% We conclude this section by revisiting the application scenarios of
% \S\ref{sec:scenarios} in terms of their programming system and using
% this to pose two questions:

% \begin{enumerate}
% \item What programming systems and techniques are \emph{currently}
%   used in implementing the different systems?, and
% \item What elements identified in this section \emph{could} usefully
%   be deployed or further explored in each system?
% \end{enumerate}

% Table~\ref{table:progSystems} summarizes the programming systems
% currently dominant in each application case study. It is worth noting
% that -- in common with most long-lived systems -- many of the cases
% have accreted substantial code bases over time and so show little in
% the way of systematic architecture. (One reason for the dominance of
% pipelines may be that such architectures are quite suitable for adding
% extra processing steps to existing computations.)

% From this we can make several observations and recommendations. The
% use of scripted or \textit{ad hoc} pipelines can in most cases be
% easily replaced with a structured workflow engine. Doing so brings
% immediate benefits in terms of reproducibility, but also in terms of
% standardization of components and interfaces. Standardization is not
% simply a software issue: it has scientific significance in the sense
% that data produced using well-documented, well-used and well-debugged
% suites of components is often more trusted -- and trustworthy -- that
% data produced using locally-produced components that, while they may
% be more innovative and targeted, also run the risk of undetected
% assumptions and errors.

% A further implication of standardization is that component re-use is
% greatly improved. This is perhaps best illustrated by observing the
% need for collecting data provenance over processing runs. The
% recording of processing of processing steps and the like is completely
% generic and can be carried out by the workflow engine uniformly across
% all workflows. The use of standard component-based software techniques
% such as web services, coupled with web technologies for metadata
% representation, can further assist in re-use, re-purposing and
% integration.

% \note{One place in which \textit{ad hoc} approaches may improve over
% standardized engines is in adaptation: the \textit{coordination}
% vector. A targeted approach can make use of application-specific
% analysis and control, whereas a generic approach needs to be able to
% describe the adaptation strategy for \textit{any} application. This is
% a challenge at the current state of the art in autonomic systems, but
% again, if addressed, would make it easier to apply advanced solutions
% across domains and reduce the barriers to entry in creating adaptive
% codes.}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "3dpas_intro"
%%% End: